---
title: "Experiment 1: results"
output: word_document
date: "2025-09-22"
---

```{r global_options, include=FALSE, cache=F}
knitr::opts_chunk$set(fig.width=4, fig.height=3, fig.crop = F, 
                      fig.pos = "tb", fig.path='figs/',
                      echo=F, warning=F, cache=T, 
                      message=F, cache=T, sanitize = T)
```

```{r setup, include = FALSE, cache=F}

library('groundhog')
groundhog.library(
  c(
    'papaja',
    'tidyverse',
    'broom',
    'lme4',
    'lsr' # for cohen's d
  ), "2025-06-01"
)

r_refs("r-references.bib")

knitr::opts_chunk$set(warning=F,echo=F,message=F,cache=T)

```

```{r load and format, include = FALSE, cache=F}

# The label for catch trials does not work, so recoding here

fruit_list <-  c("Apples",
    "Bananas",
    "Cherry",
    "Grapes",
    "Oranges",
    "Pear",
    "Strawberries")

# Path to your txt file
file <- '/Users/sawakeishiro/Desktop/code/reconstruction/data/Experiment1/experiment1_data.txt'

# Read all lines
lines <- readLines(file)

# Find the indices where a new CSV starts
starts <- which(grepl('^"?\\s*success', lines))

# Add an extra index at the end to close the last chunk
starts <- c(starts, length(lines) + 1)

# Split into chunks
chunks <- map2(starts[-length(starts)], starts[-1] - 1,
               ~ lines[.x:.y])

# Parse each chunk as a CSV
dfs <- map(chunks, ~ read_csv(paste(.x, collapse = "\n")))

# Combine
raw_df <- bind_rows(dfs) %>%
  mutate(subj_id=PROLIFIC_PID,
         correct=correct=='true',
         is_catch_trial=is_catch_trial=='true',
         left_number = as.numeric(left_number),
         middle_number=as.numeric(middle_number),
         right_number=as.numeric(middle_number),
         cumulative_score = as.numeric(cumulative_score),
         is_consistent=as.logical(is_consistent),
         trial_index = as.numeric(trial_index),
         trial_index_e2=as.numeric(trial_index_e2),
         RT=as.numeric(rt),
         is_catch_trial = (left_word %in% fruit_list) | 
           (right_word %in% fruit_list) | 
           (middle_word %in% fruit_list)) %>%
  filter(!is.na(subj_id))

total_N <- raw_df$subj_id %>% unique() %>% length()


  
```

```{r exclusion and formatting, include = FALSE, cache=F}

phase_1_df <- raw_df %>%
  filter(phase=='part1') %>%
  dplyr::select(subj_id, trial_index,left_word,left_number,
                middle_word,middle_number,right_word,right_number,
                correct_choice_position,participant_response_position, 
                is_catch_trial,rule_condition,RT) %>%
  mutate(correct=correct_choice_position ==
           participant_response_position) %>% # just to make sure
  group_by(subj_id) %>%
  arrange(trial_index)%>%
  mutate(trial_index = 1:n(),
         chosen_side=ifelse(rule_condition==participant_response_position, 
                            'only_stones','other_side'),
         chose_gems_p1 = participant_response_position == correct_choice_position,
         chose_onlystones_p1 = participant_response_position == rule_condition)


participants_to_exclude_comprehension <- raw_df %>%
  group_by(subj_id) %>%
  summarise(failures_1 = mean(comprehension_failures_part1,na.rm=T),
            failures_2 = mean(comprehension_failures_part2,na.rm=T)) %>%
  rowwise() %>%
  mutate(max_failures = max(failures_1,failures_2)) %>%
  filter(max_failures > 1) %>%
  pull(subj_id)

participants_to_exclude_catch <- phase_1_df %>%
  filter(is_catch_trial) %>%
  group_by(subj_id) %>%
  summarise(acc_catch=mean(correct)) %>%
  filter(acc_catch<0.5) %>% #2 or 3 errors out of three
  pull(subj_id)

participants_to_exclude_RT <- raw_df %>%
  filter(!is.na(phase))%>%
  group_by(subj_id,phase) %>%
  summarise(RT_first_quartile=quantile(RT,0.25,na.rm=T)) %>%
  filter(RT_first_quartile<=100) %>%
  pull(subj_id)

participants_to_exclude <- unique(c(participants_to_exclude_RT,
                                    participants_to_exclude_catch, 
                                    participants_to_exclude_comprehension))

phase_1_df <- phase_1_df  %>%
  filter(!(subj_id %in% participants_to_exclude)) 

phase_2_df <- raw_df %>%
  filter(phase=='part2') %>%
  filter(!(subj_id %in% participants_to_exclude)) %>%
  dplyr::select(subj_id, trial_index, 
                participant_response_position_e2,
                rule_condition, is_consistent, RT, is_catch_trial) %>%
  mutate(reconstructed_action = participant_response_position_e2) %>%
  rename(RT2=RT) %>%
  group_by(subj_id) %>%
  arrange(trial_index)%>%
  mutate(trial_index = 1:n(),
         chosen_side_part2=ifelse(rule_condition==reconstructed_action, 
                                  'only_stones','other_side')) %>%
  inner_join(
    phase_1_df %>%
      mutate(correct_choice_position_p1 = correct_choice_position) %>%
      select(trial_index, subj_id, chose_gems_p1, correct_choice_position_p1, chose_onlystones_p1),
    by = c("subj_id", "trial_index")
  ) 

two_parts_df <- phase_1_df %>%
  merge(phase_2_df) %>%
  rowwise()%>%
  mutate(number1 = get(paste0(participant_response_position, "_number")),
         number2 = get(paste0(participant_response_position_e2, "_number")),
         chose_highest1 = number1 == pmax(left_number, middle_number, right_number),
         chose_highest2 = number2 == pmax(left_number, middle_number, right_number))


included_N <- two_parts_df$subj_id %>% unique() %>% length()

participants_by_condition <- two_parts_df %>%
  group_by(subj_id) %>%
  summarise(condition = rule_condition[1]) %>%
  group_by(condition) %>%
  summarise(n=n()) %>%
  spread(condition,n)

```

```{r explicit rule discoverers, include = FALSE, cache=F}

rule_check_df <- 
  raw_df %>%
  filter(grepl("rule_check", response)) %>% 
  dplyr::select(subj_id,response,rule_condition)

discovered_rule <- rule_check_df %>%
  filter(subj_id %in% c(
    '60fe7bf61c7567f5cca9ade6',
    '5ecf9b46617a9b3dd56b3fe3',
    '5ebc0abd2cc5f1054c82a0d7',
    '654bc148c95bca00d541ec33',
    '66dcc0e54f44ccecf2730c76',
    '631d197eba06364cbea7fab2',
    '66fd3f6c93c24ecc119be354',
    '66e9236f775eae67be04229d',
    '66b244a2f436437d3e109bbe',
    '63ee9abfa262bea95766dd9c',
    '66cfab152c56aa0472529001',
    '5bf368846edd7200017136c7',
    '66d9cd06e207a93377f7820e',
    '5962a5b29955ee00015bce98',
    '632d189d6b8f03d0d3db4c48',
    '63b3377ca1ccdfc7cc567503',
    '6658bba1a51a037d25bb677e',
    '662945728093113a43ab8077',
    '676058418c08c71e4bb7cbd1',
    '58954e2dc88c680001dd4b7e',
    '681b781064a5701a5ceee3bc',
    '682b3f63d0e2c4eff64fe664',
    '5d3695969749cc00165c222f',
    '6773dea609f77d473900997a',
    '60e3a25e8be835f754795d93') &
      !(subj_id %in% participants_to_exclude)) 

two_parts_df <- two_parts_df %>% 
  mutate(mentioned_rule = subj_id %in% discovered_rule$subj_id)

stoneposition<- raw_df %>%
  group_by(rule_condition) %>%
  summarise(n=length(unique(subj_id)))

```

`r total_N` participants took part in the experiment. `r participants_to_exclude_catch%>%length()` were excluded based on failing 2 or more catch trials, `r participants_to_exclude_comprehension%>%length()` based on requiring three or more repetitions of the instructions before passing the comprehension checks in part 1 or in part 2, and `r participants_to_exclude_RT%>%length()` based on responding too quickly (100 ms or below) on at least 25% of trials in at least one part of the experiment, leaving `r included_N` participants for the main analysis: `r participants_by_condition$left` for whom the left position never had gems, `r participants_by_condition$middle` for whom it was the middle one and `r participants_by_condition$right` for whom it was the right one.

When asked at the end of the experiment whether they discovered a rule, a subset of `r discovered_rule%>%nrow()` participants (hereafter, explicit learners) mentioned the location of the boxes -- either directly specifying the rule or mentioning that a specific position was more likely to have gems in it.

## phase 1

```{r accuracy,  cache=F, fig.width=8}

library(dplyr)
library(ggplot2)

# --- Define bins exactly as requested ---
# Define the breaks. This creates 13 breaks (0.5, 5.5, ..., 60.5)
# which define 12 distinct bins.
breaks <- seq(0.5, 60.5, by = 5)

# Create the corresponding 12 labels for the new bins
bin_labels <- c(
  "1–5", "6–10", "11–15", "16–20", "21–25", "26–30", 
  "31–35", "36–40", "41–45", "46–50", "51–55", "56–60"
)

# This code for vertical lines will work perfectly as-is.
# It automatically adjusts to the new length(bin_labels), which is 12.
# It will create lines at 1.5, 2.5, ..., 11.5
vline_x <- seq(1.5, length(bin_labels) - 0.5, by = 1)

# --- Clean + bin ---
df_binned <- two_parts_df %>%
  filter(!is_catch_trial) %>%     # drop catch trials if present
  mutate(
    bin = cut(trial_index, breaks = breaks, labels = bin_labels,
              include.lowest = TRUE, right = TRUE),
    bin = factor(bin, levels = bin_labels)
  )

# --- Per-subject accuracy within each bin & rule ---
by_subj <- df_binned %>%
  group_by(subj_id, mentioned_rule, bin) %>%
  summarise(acc = mean(correct, na.rm = TRUE), .groups = "drop")

# --- Group mean ± SE across participants (within each rule × bin) ---
group_summary <- by_subj %>%
  group_by(mentioned_rule, bin) %>%
  summarise(
    m  = mean(acc, na.rm = TRUE),
    se = sd(acc,  na.rm = TRUE) / sqrt(dplyr::n()),
    n  = dplyr::n(),    # number of participants contributing in that rule × bin
    .groups = "drop"
  )

# --- Plot: two colours (rules), points ± SE, vertical bin lines ---
pd <- position_dodge(width = 0.5)

ggplot(group_summary, aes(x = bin, y = m, color = mentioned_rule, group = mentioned_rule)) +
  geom_vline(xintercept = vline_x, linetype = "dashed", linewidth = 0.3) +
  geom_hline(yintercept = 1/2, linewidth = 0.3) +
  geom_hline(yintercept = 1/3, linewidth = 0.3) +# optional chance line
  geom_line(linewidth = 0.7,position=pd) +                     # connect group means
  geom_point(position = pd, size = 2) +
  geom_errorbar(aes(ymin = m - se, ymax = m + se), position = pd, width = 0.2) +
  labs(x = "Trial bin", y = "Accuracy", color = "mentioned position",
       title = "Binned Accuracy by rule discovery (group mean ± SE)") +
  theme_minimal()

phase_1_acc <- two_parts_df %>%
  group_by(subj_id,mentioned_rule) %>%
  filter(!is_catch_trial) %>%
  summarise(acc=mean(correct))

# acc_slopes <- phase_1_df %>%
#   group_by(subj_id) %>%
#   filter(!is_catch_trial) %>%
#   do(tidy(glm(correct ~ trial_index, family = binomial, data = .))) %>%
#   filter(term == "trial_index") %>%
#   select(subj_id, slope = estimate)

```

In part 1, mean proportion correct (in non-catch trials) was `r phase_1_acc %>% pull(acc) %>%mean()%>%printnum()` (median=`r phase_1_acc %>% pull(acc) %>%median()%>%printnum()`) and significantly higher than chance (1/3; `r apa_print(t.test(phase_1_acc %>% pull(acc),mu=1/3))$statistic`).

Among the explicit learners, accuracy was `r phase_1_acc %>% filter(mentioned_rule) %>% pull(acc) %>%mean()%>%printnum()`, and among the rest of participants it was `r phase_1_acc %>% filter(!mentioned_rule) %>% pull(acc) %>%mean()%>%printnum()` and still higher than chance (`r apa_print(phase_1_acc %>% filter(!mentioned_rule) %>% pull(acc) %>%t.test(mu=1/3))$statistic`).

```{r position avoidance, cache=F, fig.width=8}

# --- Per-subject accuracy within each bin & rule ---
by_subj1 <- df_binned %>%
  group_by(subj_id, mentioned_rule, bin) %>%
  summarise(bias = mean(chosen_side=='only_stones', na.rm = TRUE), .groups = "drop") %>%
  mutate(part=1)

by_subj2 <- df_binned %>%
  group_by(subj_id, mentioned_rule, bin) %>%
  summarise(bias = mean(chosen_side_part2=='only_stones', na.rm = TRUE), .groups = "drop") %>%
  mutate(part=2)

by_subj <- by_subj1 %>% rbind(by_subj2) %>%
  mutate(part=factor(part))

# --- Group mean ± SE across participants (within each rule × bin) ---
group_summary <- by_subj %>%
  group_by(mentioned_rule, part, bin) %>%
  summarise(
    m  = mean(bias, na.rm = TRUE),
    se = sd(bias,  na.rm = TRUE) / sqrt(dplyr::n()),
    n  = dplyr::n(),    # number of participants contributing in that rule × bin
    .groups = "drop"
  )

# --- Plot: two colours (rules), points ± SE, vertical bin lines ---
ggplot(group_summary, aes(x = bin, y = m, color = mentioned_rule,shape = part, linetype=part, group = interaction(mentioned_rule,part))) +
  geom_vline(xintercept = vline_x, linetype = "dashed", linewidth = 0.3) +
  geom_hline(yintercept = 1/3, linewidth = 0.3) +  # optional chance line
  geom_line(linewidth = 0.7,position=pd) +                     # connect group means
  geom_point(position = pd, size = 2) +
  geom_errorbar(aes(ymin = m - se, ymax = m + se), position = pd, width = 0.2, linetype='solid') +
  labs(x = "Trial bin", y = "P(stones only)", color = "mentioned position",
       title = "Binned Bias by rule discovery (group mean ± SE)") +
  theme_minimal()


phase_1_pstones <- two_parts_df %>%
  group_by(subj_id, mentioned_rule) %>%
  filter(!is_catch_trial) %>%
  summarise(pstones=mean(chosen_side=='only_stones'))

phase_2_pstones <- two_parts_df %>%
  group_by(subj_id, mentioned_rule) %>%
  filter(!is_catch_trial) %>%
  summarise(pstones = mean(chosen_side_part2=='only_stones'))
  

pstones_slopes <- two_parts_df %>%
  group_by(subj_id, mentioned_rule) %>%
  filter(!is_catch_trial) %>%
  mutate(chose_stone_side = chosen_side == 'only_stones') %>%
  do(tidy(glm(chose_stone_side ~ trial_index, family = binomial, data = .))) %>%
  filter(term == "trial_index") %>%
  select(subj_id, slope = estimate)

pstones_slopes_p2 <- two_parts_df %>%
  group_by(subj_id, mentioned_rule) %>%
  filter(!is_catch_trial) %>%
  mutate(chose_stone_side = chosen_side_part2 == 'only_stones') %>%
  do(tidy(glm(chose_stone_side ~ trial_index, family = binomial, data = .))) %>%
  filter(term == "trial_index") %>%
  select(subj_id, slope = estimate)

ggplot(pstones_slopes_p2, aes(x = slope)) +
  geom_histogram(bins = 30) +
  labs(
    x = "Slope",
    y = "Count"
  )

pstone_bytrial <- two_parts_df %>%
    group_by(subj_id, mentioned_rule) %>%
    filter(!is_catch_trial) %>%
    mutate(chose_stone_side = chosen_side_part2 == 'only_stones') %>% group_by(trial_index) %>% summarise(pstones = mean(chose_stone_side))

ggplot(pstone_bytrial, aes(x = trial_index, y = pstones)) +
  geom_line() +
  geom_point() +
  ylim(0, 1) +
  labs(
    x = "Trial",
    y = "P(only stones)"
  )

```

Hypothesis 1 (RULE LEARNING: AVOIDANCE): In part 1, participants chose the side that always had stones (in non-catch trials) on `r phase_1_pstones %>% pull(pstones) %>%mean()%>%printnum()` (median=`r phase_1_pstones %>% pull(pstones) %>%median()%>%printnum()`) of all trials, significantly below chance (1/3; `r apa_print(t.test(phase_1_pstones %>% pull(pstones),mu=1/3))$statistic`).

This figure was `r phase_1_pstones %>% filter(mentioned_rule) %>% pull(pstones) %>%mean()%>%printnum()` among explicit learners and `r phase_1_pstones %>% filter(!mentioned_rule) %>% pull(pstones) %>%mean()%>%printnum()`, still significantly below 0.33 (`r apa_print(phase_1_pstones %>% filter(!mentioned_rule) %>% pull(pstones) %>%t.test(mu=0.33))$statistic`), among the remaining participants.

Hypothesis 2 (RULE LEARNING: SLOPES): Participants improved as a function of trial number (a t-test on subject-specific learning slopes from a logistic regression against 0: `r apa_print(t.test(pstones_slopes %>% pull(slope)))$statistic`). This was marginally the case for those participants who were not explicit learners (`r apa_print(t.test(pstones_slopes %>% filter(!mentioned_rule)%>%pull(slope)))$statistic`))

Participants did avoid stones as a function of trial number (a t-test on subject-specific learning slopes from a logistic regression against 0): `r apa_print(t.test(pstones_slopes_p2 %>% pull(slope)))$statistic`). Explicit learners showed that they avoid the (`r apa_print(t.test(pstones_slopes_p2 %>% filter(mentioned_rule)%>%pull(slope)))$statistic`)
This was not the case for those participants who were not explicit learners (`r apa_print(t.test(pstones_slopes_p2 %>% filter(!mentioned_rule)%>%pull(slope)))$statistic`))

## phase 2

```{r phase 2 accuracy, include = FALSE, cache=F}


phase_2_acc <- two_parts_df %>%
  group_by(subj_id,mentioned_rule) %>%
  filter(!is_catch_trial) %>%
  summarise(acc=mean(is_consistent))


```

Hypothesis 3 (RECONSTRUCTION ACCURACY): In part 2, mean proportion correct (in non-catch trials) was `r phase_2_acc %>% pull(acc) %>%mean()%>%printnum()` (median=`r phase_2_acc %>% pull(acc) %>%median()%>%printnum()`) and significantly higher than chance (1/3; `r apa_print(t.test(phase_2_acc %>% pull(acc),mu=1/3))$statistic`).

This figure was `r phase_2_acc %>%filter(mentioned_rule)%>% pull(acc) %>%mean()%>%printnum()` and `r phase_2_acc %>%filter(!mentioned_rule)%>% pull(acc) %>%mean()%>%printnum()` among other participants.

In part 2, participants chose the side that always had stones (in non-catch trials) on `r phase_2_pstones %>% pull(pstones) %>%mean()%>%printnum()` (median=`r phase_2_pstones %>% pull(pstones) %>%median()%>%printnum()`) of all trials, significantly below chance (1/3; `r apa_print(t.test(phase_2_pstones %>% pull(pstones),mu=1/3))$statistic`).

This figure was `r phase_2_pstones %>% filter(mentioned_rule) %>% pull(pstones) %>%mean()%>%printnum()` among explicit learners and `r phase_2_pstones %>% filter(!mentioned_rule) %>% pull(pstones) %>%mean()%>%printnum()`, still significantly below 0.33 (`r apa_print(phase_2_pstones %>% filter(!mentioned_rule) %>% pull(pstones) %>%t.test(mu=0.33))$statistic`), among the remaining participants. 

```{r phase 2 accuracy by phase 1 accuracy, include = FALSE, cache=F}


phase_2_acc_by_phase_1_acc <- two_parts_df %>%
  group_by(subj_id,correct,mentioned_rule) %>%
  filter(!is_catch_trial) %>%
  summarise(acc=mean(is_consistent)) %>%
  spread(correct,acc,sep='')%>%
  mutate(diff=correctTRUE-correctFALSE)


```

Hypothesis 4 (RECONSTRUCTION ACCURACY BY GUESS ACCURACY): Participants' ability to reconstruct their past decisions was significantly higher when they were correct in part 1 (mean accuracy = `r phase_2_acc_by_phase_1_acc$correctTRUE%>%mean()%>%printnum()`) compared to incorrect (mean accuracy = `r phase_2_acc_by_phase_1_acc$correctFALSE%>%mean()%>%printnum()`; `r apa_print(phase_2_acc_by_phase_1_acc$diff%>%t.test())$statistic`).

This effect was similarly strong between explicit learners and other participants (`r apa_print(t.test(phase_2_acc_by_phase_1_acc%>%filter(!mentioned_rule)%>%pull(diff), phase_2_acc_by_phase_1_acc%>%filter(mentioned_rule)%>%pull(diff)))$statistic`).

```{r figure, include = FALSE, cache=F}

# hyp4_bins_design_like_example.R
# Pure R script - bin x accuracy with 4 lines (mentioned_rule × part1 correctness)
# Assumes two_parts_df exists and has subj_id, mentioned_rule, correct, is_consistent, trial_index, is_catch_trial

library(dplyr)
library(tidyr)
library(ggplot2)
library(scales)

# --- bins as you used before ---
breaks <- seq(0.5, 60.5, by = 5)
bin_labels <- c(
  "1–5", "6–10", "11–15", "16–20", "21–25", "26–30", 
  "31–35", "36–40", "41–45", "46–50", "51–55", "56–60"
)

# vertical line x positions for categorical bins:
vline_x <- seq(1.5, length(bin_labels) - 0.5, by = 1)

# position dodge to separate overlapping points/errorbars
pd <- position_dodge(width = 0.5)

# --- Per-subject accuracy within each bin & rule × part1 correctness ---
# p1_correct = TRUE when participant was correct in Part 1 on that trial
df_binned <- two_parts_df %>%
  filter(!is_catch_trial) %>%
  mutate(
    bin = cut(trial_index, breaks = breaks, labels = bin_labels,
              include.lowest = TRUE, right = TRUE),
    bin = factor(bin, levels = bin_labels),
    p1_correct = as.logical(correct)   # TRUE/FALSE
  )

# For plotting in the style of your example, compute per-subject means within each bin × condition
by_subj_correct <- df_binned %>%
  group_by(subj_id, mentioned_rule, bin) %>%
  filter(p1_correct == TRUE) %>%
  summarise(acc = mean(is_consistent, na.rm = TRUE), .groups = "drop") %>%
  mutate(part = "Part1: correct")

by_subj_incorrect <- df_binned %>%
  group_by(subj_id, mentioned_rule, bin) %>%
  filter(p1_correct == FALSE) %>%
  summarise(acc = mean(is_consistent, na.rm = TRUE), .groups = "drop") %>%
  mutate(part = "Part1: incorrect")

# bind into one dataframe, part as factor (so shapes/linetypes map to it)
by_subj <- bind_rows(by_subj_correct, by_subj_incorrect) %>%
  mutate(part = factor(part, levels = c("Part1: incorrect", "Part1: correct")))

# --- Group mean ± SE across participants (within each rule × part × bin) ---
group_summary <- by_subj %>%
  group_by(mentioned_rule, part, bin) %>%
  summarise(
    m  = mean(acc, na.rm = TRUE),
    se = sd(acc,  na.rm = TRUE) / sqrt(n()),
    n  = n(),
    .groups = "drop"
  ) %>%
  # friendly labels for facet/legend if you want
  mutate(mentioned_rule = factor(mentioned_rule, levels = c(FALSE, TRUE),
                                 labels = c("Did NOT mention rule", "Mentioned rule")))

# ensure bin order
group_summary$bin <- factor(group_summary$bin, levels = bin_labels)

# --- Plot: two colours (rules), shapes/linetype for part, vertical bin lines ---
p <- ggplot(group_summary, aes(x = bin, y = m, color = mentioned_rule,
                               shape = part, linetype = part,
                               group = interaction(mentioned_rule, part))) +
  # vertical separators between bins (dashed thin)
  geom_vline(xintercept = vline_x, linetype = "dashed", linewidth = 0.3, color = "grey80") +
  # horizontal chance line at 1/3
  geom_line(linewidth = 0.8, position = pd) +
  geom_point(position = pd, size = 2.4) +
  geom_errorbar(aes(ymin = m - se, ymax = m + se), position = pd, width = 0.18, size = 0.7) +
  scale_y_continuous(labels = percent_format(accuracy = 1), limits = c(0.4, 1), breaks = seq(0,1,0.2)) +
  labs(
    x = "Trial bin",
    y = "Accuracy (reconstruction)",
    color = "Rule report",
    shape = "Part 1 correctness",
    linetype = "Part 1 correctness",
    title = "Binned reconstruction accuracy by Part 1 correctness and rule report",
    subtitle = "Points = group mean; error bars = ±SE; colored = mentioned rule vs not; line type/shape = Part1 correct/incorrect"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "right",
    plot.title = element_text(face = "bold"),
    plot.subtitle = element_text(size = 10)
  )

# print and save
print(p)



by_subj_nonexp <- bind_rows(by_subj_correct, by_subj_incorrect) %>%
  mutate(part = factor(part, levels = c("Part1: incorrect", "Part1: correct")))

# --- Group mean ± SE across participants for this subgroup ---
group_summary_nonexp <- by_subj_nonexp %>%
  group_by(part, bin) %>%
  summarise(
    m  = mean(acc, na.rm = TRUE),
    se = sd(acc,  na.rm = TRUE) / sqrt(n()),
    n  = n(),
    .groups = "drop"
  ) %>%
  mutate(bin = factor(bin, levels = bin_labels))

# --- Plot ---
p_nonexp <- ggplot(group_summary_nonexp, aes(x = bin, y = m, shape = part, linetype = part, group = part)) +
  geom_vline(xintercept = vline_x, linetype = "dashed", linewidth = 0.3, color = "grey85") +
  geom_line(linewidth = 0.8, position = pd, color = "black") +
  geom_point(position = pd, size = 2.6, color = "black", fill = "white") +
  geom_errorbar(aes(ymin = m - se, ymax = m + se), position = pd, width = 0.18, size = 0.7, color = "black") +
  scale_y_continuous(labels = percent_format(accuracy = 1), limits = c(0.5, 1), breaks = seq(0,1,0.2)) +
  labs(
    x = "Trial bin",
    y = "Reconstruction accuracy (Part 2)",
    shape = "Part 1 correctness",
    linetype = "Part 1 correctness",
    title = "Binned reconstruction accuracy (Non-explicit learners)",
    subtitle = "Only participants who did NOT report the rule; line type/shape = Part1 correct vs incorrect"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "right",
    plot.title = element_text(face = "bold"),
    plot.subtitle = element_text(size = 10)
  )

# Print and save
print(p_nonexp)






```

```{r hyp4_plot, fig.width=8}
# hyp4_plot.R
# Pure R script to plot Hypothesis 4 (reconstruction accuracy by Part 1 correctness)
# Assumes `two_parts_df` exists in the environment.

# Packages
library(dplyr)
library(tidyr)
library(ggplot2)
library(scales)

# ---------------------------
# Prepare data
# ---------------------------

# Compute per-subject reconstruction accuracy split by whether Part 1 was correct
phase2_by_p1acc <- two_parts_df %>%
  filter(!is_catch_trial) %>%
  mutate(p1_correct = correct) %>%                    # TRUE/FALSE
  group_by(subj_id, mentioned_rule, p1_correct) %>%
  summarise(recon_acc = mean(is_consistent, na.rm = TRUE), .groups = "drop")

# Make wide table for paired test: only subjects with both p1_TRUE and p1_FALSE
wide_for_test <- phase2_by_p1acc %>%
  pivot_wider(names_from = p1_correct, names_prefix = "p1_", values_from = recon_acc) %>%
  filter(!is.na(p1_TRUE) & !is.na(p1_FALSE))

# Paired t-test (only on subjects that have both cell types)
if (nrow(wide_for_test) >= 2) {
  paired_test <- t.test(wide_for_test$p1_TRUE, wide_for_test$p1_FALSE, paired = TRUE)
  print(paired_test)
  pval <- paired_test$p.value
  pval_txt <- if (pval < 0.001) "p < .001" else paste0("p = ", signif(pval, 3))
} else {
  paired_test <- NULL
  pval_txt <- "Not enough subjects with both trial types for paired test"
  message(pval_txt)
}

# Group summary for plotting
group_summary <- phase2_by_p1acc %>%
  group_by(mentioned_rule, p1_correct) %>%
  summarise(
    m = mean(recon_acc, na.rm = TRUE),
    se = sd(recon_acc, na.rm = TRUE) / sqrt(n()),
    n = n(),
    .groups = "drop"
  ) %>%
  mutate(p1_label = ifelse(p1_correct, "Part1: correct", "Part1: incorrect"),
         p1_label = factor(p1_label, levels = c("Part1: incorrect","Part1: correct")))

# Participant-level for paired lines
plot_pairs <- phase2_by_p1acc %>%
  mutate(p1_label = ifelse(p1_correct, "Part1: correct", "Part1: incorrect")) %>%
  mutate(p1_label = factor(p1_label, levels = c("Part1: incorrect","Part1: correct")))

# ---------------------------
# Plot
# ---------------------------

p <- ggplot() +
  # lines connecting each participant's paired points
  geom_line(data = plot_pairs, aes(x = p1_label, y = recon_acc, group = subj_id),
            color = "grey70", alpha = 0.35, size = 0.4) +
  # individual subject points
  geom_point(data = plot_pairs, aes(x = p1_label, y = recon_acc),
             color = "grey70", alpha = 0.4, position = position_dodge(width = 0.4), size = 1.7) +
  # group mean points
  geom_point(data = group_summary, aes(x = p1_label, y = m, color = mentioned_rule),
             position = position_dodge(width = 0.4), size = 3) +
  # error bars for mean ± SE
  geom_errorbar(data = group_summary, aes(x = p1_label, ymin = m - se, ymax = m + se, color = mentioned_rule),
                width = 0.12, position = position_dodge(width = 0.4), size = 0.7) +
  facet_wrap(~ mentioned_rule, nrow = 1,
             labeller = as_labeller(c("FALSE" = "Did NOT mention position", "TRUE" = "Mentioned position"))) +
  scale_y_continuous(limits = c(0.25,1), breaks = seq(0,1,0.25), labels = percent_format(accuracy = 1)) +
  labs(x = NULL,
       y = "Reconstruction accuracy (Part 2)",
       x = "Part 1 guess",
       title = "Reconstruction Accuracy in Part 2 by Part 1 Guess",
       color = "Mentioned rule") +
  theme_minimal(base_size = 12) +
  theme(legend.position = "none",
        strip.text = element_text(face = "bold"),
        plot.title = element_text(size = 12, face = "bold"),
        plot.subtitle = element_text(size = 10))

# Print to screen
print(p)


```

```{r remembered versus actual accuracy, include = FALSE, cache=F}

remembered_versus_actual_accuracy <- two_parts_df %>%
  group_by(subj_id, mentioned_rule) %>%
  filter(!is_catch_trial) %>%
  summarise(remembered_accuracy=mean(participant_response_position_e2==correct_choice_position),
            actual_accuracy = mean(participant_response_position==correct_choice_position),
            diff=remembered_accuracy-actual_accuracy)

```

Hypothesis 5 (REMEMBERED ACCURACY VERSUS TRUE ACCURACY): Participants remembered their guesses as aligned with the location of the gems on `r remembered_versus_actual_accuracy$remembered_accuracy%>%mean()%>%printnum()`, a frequency significantly higher than their actual accuracy in part 1, which was `r remembered_versus_actual_accuracy$actual_accuracy%>%mean()%>%printnum()` (`r apa_print(remembered_versus_actual_accuracy$diff%>%t.test())$statistic`).

This effect was not statistically different between explicit learners and other participants (`r apa_print(t.test(remembered_versus_actual_accuracy%>%filter(!mentioned_rule)%>%pull(diff), remembered_versus_actual_accuracy%>%filter(mentioned_rule)%>%pull(diff)))$statistic`).

```{r phase 2 position bias, include = FALSE, cache=F}

phase_2_position_bias <- two_parts_df %>%
  group_by(subj_id, mentioned_rule) %>%
  filter(!is_catch_trial) %>%
  summarise(bias=mean(chosen_side_part2=='only_stones')) %>%
  merge(phase_1_pstones) %>%
  mutate(diff = bias-pstones)

```

In phase 2, participants chose the stone-only option on `r phase_2_position_bias$bias%>%mean()%>%printnum()` of all trials, significantly less than 1/3 (`r apa_print(phase_2_position_bias$bias%>%t.test(mu=1/3))$statistic`), and significantly less than the proportion in phase 1 (`r apa_print(phase_2_position_bias$diff%>%t.test())$statistic`).

```{r phase 2 position bias incorrect only, include = FALSE, cache=F}

phase_2_position_bias_incorrect_only <- two_parts_df %>%
  group_by(subj_id, mentioned_rule) %>%
  filter(!is_catch_trial & !is_consistent & correct) %>%
  summarise(bias=mean(chosen_side_part2=='only_stones'))

dat <- two_parts_df %>%
  group_by(subj_id) %>%
  filter(!is_catch_trial & !is_consistent & correct) %>%
  mutate(decision=as.integer(chosen_side_part2=='only_stones'))

m <- glmer(decision ~ 1 + (1 | subj_id),
           data = dat, family = binomial,
           control = glmerControl(optimizer = "bobyqa"))

```





```{r, r_plots, fig.width=8}

never_stone_ids <- two_parts_df %>%
  group_by(subj_id) %>%
  summarise(ever_stone = any(chosen_side_part2 == 'only_stones', na.rm = TRUE)) %>%
  filter(!ever_stone) %>%
  pull(subj_id)
  

h6_df <- two_parts_df %>%
  filter(!is_catch_trial,
         !is.na(correct),
         correct == TRUE,       
         is_consistent == FALSE,
         !(subj_id %in% never_stone_ids)
         ) %>%  
  mutate(
    chose_onlystones = (chosen_side_part2 == "only_stones"),
    # keep trial_index and mentioned_rule as-is
    trial_index = as.integer(trial_index)
  )

control_opts <- glmerControl(optimizer = "bobyqa")

# Overall model
h6_model <- glmer(chose_onlystones ~ 1 + (1 | subj_id),
                  data = h6_df,
                  family = binomial,
                  control = control_opts)

h6_model_mentioned_vs_notmentioned <- glmer(chose_onlystones ~ mentioned_rule + (1 | subj_id),
                  data = h6_df,
                  family = binomial,
                  control = glmerControl(optimizer = "bobyqa"))

# Model for participants who mentioned the rule
r_h6_model <- glmer(chose_onlystones ~ 1 + (1 | subj_id),
                            data = filter(h6_df, mentioned_rule == TRUE),
                            family = binomial,
                            control = control_opts)

# Model for participants who mentioned the rule
h6_model_mentioned <- glmer(chose_onlystones ~ 1 + (1 | subj_id),
                            data = filter(h6_df, mentioned_rule == TRUE),
                            family = binomial,
                            control = control_opts)


# Model for participants who did NOT mention the rule
nr_h6_model <- glmer(chose_onlystones ~ 1 + (1 | subj_id),
                               data = filter(h6_df, mentioned_rule == FALSE),
                               family = binomial,
                               control = control_opts)

h6_model_notmentioned <- glmer(chose_onlystones ~ 1 + (1 | subj_id),
                               data = filter(h6_df, mentioned_rule == FALSE),
                               family = binomial,
                               control = control_opts)

h6_model_trial <- glmer(chose_onlystones ~ trial_index + (1 + trial_index | subj_id),
                  data = h6_df,
                  family = binomial,
                  control = control_opts)

r_h6_model_trial <- glmer(chose_onlystones ~ trial_index + (1 + trial_index | subj_id),
                  data = filter(h6_df, mentioned_rule==TRUE),
                  family = binomial,
                  control = control_opts)

nr_h6_model_trial <- glmer(chose_onlystones ~ trial_index + (1 + trial_index | subj_id),
                  data = filter(h6_df, mentioned_rule==FALSE),
                  family = binomial,
                  control = control_opts)


inv_logit <- function(x) 1 / (1 + exp(-x))

get_prob_and_p <- function(model) {
  cf <- fixef(model)
  se  <- sqrt(diag(vcov(model)))
  intercept <- cf[1]
  se_intercept <- se[1]
  z <- intercept / se_intercept
  p <- 2 * (1 - pnorm(abs(z)))
  prob <- inv_logit(intercept)
  list(intercept = intercept, se = se_intercept, z = z, p = p, prob = prob)
}

overall <- get_prob_and_p(h6_model)
mentioned <- get_prob_and_p(h6_model_mentioned)
notmentioned <- get_prob_and_p(h6_model_notmentioned)

cat("\nEstimated probability (overall) of choosing stones-only (on these error trials): ",
    round(overall$prob, 3), " (p = ", signif(overall$p, 3), ")\n", sep = "")

cat("Estimated probability (mentioned rule): ",
    if (!is.null(mentioned$prob)) round(mentioned$prob, 3) else NA,
    " (p = ", if (!is.null(mentioned$p)) signif(mentioned$p, 3) else NA, ")\n", sep = "")

cat("Estimated probability (not mentioned rule): ",
    if (!is.null(notmentioned$prob)) round(notmentioned$prob, 3) else NA,
    " (p = ", if (!is.null(notmentioned$p)) signif(notmentioned$p, 3) else NA, ")\n", sep = "")



# Define bins of 10 trials (change breaks/labels if you prefer different binning)
breaks <- seq(0, 60, by = 10)  # 0-10,10-20,...50-60
bin_labels <- c("1–10", "11–20", "21–30", "31–40", "41–50", "51–60")

# h6_viz_data <- h6_df %>%
#   
#   mutate(
#     trial_bin = cut(trial_index,
#                     breaks = breaks,
#                     include.lowest = TRUE,
#                     labels = bin_labels,
#                     right = TRUE)
#   ) %>%
#   
#   group_by(subj_id, mentioned_rule, trial_bin) %>%
#   summarise(
#     n = n(),
#     prob_stones = mean(chose_onlystones, na.rm = TRUE),
#     se = sqrt(prob_stones * (1 - prob_stones) / pmax(n, 1)), # handle n=0 safely
#     .groups = "drop"
#   ) %>%
#   mutate(
#     ci_lower = pmax(0, prob_stones - 1.96 * se),
#     ci_upper = pmin(1, prob_stones + 1.96 * se),
#     Group = ifelse(mentioned_rule, "Position mentioned", "Position NOT mentioned"),
#     trial_bin = factor(trial_bin, levels = bin_labels)
#   )

# Per-bin (keeps trial_bin)
h6_viz_data <- h6_df %>%
  mutate(
    trial_bin = cut(trial_index,
                    breaks = breaks,
                    include.lowest = TRUE,
                    labels = bin_labels,
                    right = TRUE)
  ) %>%
  group_by(subj_id, mentioned_rule, trial_bin) %>%
  summarise(
    n = n(),
    prob_stones = mean(chose_onlystones, na.rm = TRUE),
    se = ifelse(n > 0,
                sqrt(prob_stones * (1 - prob_stones) / n), 
                NA_real_),  # avoid misleading SE when n == 0
    .groups = "drop"
  ) %>%
  mutate(
    ci_lower = pmax(0, prob_stones - 1.96 * se),
    ci_upper = pmin(1, prob_stones + 1.96 * se),
    Group = ifelse(mentioned_rule, "Position mentioned", "Position NOT mentioned"),
    trial_bin = factor(trial_bin, levels = bin_labels)
  )

trial_bin1 <- h6_viz_data %>%
  filter(trial_bin == '1-10') %>%
  summarise(mean_prob_stones = mean(prob_stones))

trial_bin2 <- h6_viz_data %>%
  filter(trial_bin == '11-20') %>%
  summarise(mean_prob_stones = mean(prob_stones))

# t.test(trial_bin1$mean_prob_stones, trial_bin2$mean_prob_stones, paired.data=TRUE)

subject_bin_data <- h6_viz_data %>%
  filter(trial_bin %in% c('1–10', '11–20'), mentioned_rule) %>%      
  select(subj_id, trial_bin, prob_stones) %>%
  pivot_wider(names_from = trial_bin, values_from = prob_stones) %>%
  drop_na(`1–10`, `11–20`)                           



print(h6_viz_data %>% select(trial_bin, Group, n))

# Plot
pd <- position_dodge(width = 0.4)
p <- ggplot(h6_viz_data, aes(x = trial_bin, y = prob_stones, group = Group, color = Group)) +
  geom_hline(yintercept = 0.5, linetype = "dashed", color = "red", linewidth = 0.6) +
  geom_errorbar(aes(ymin = ci_lower, ymax = ci_upper), width = 0.12, position = pd, alpha = 0.9) +
  geom_line(size = 1.0, alpha = 0.9, position = pd) +
  geom_point(size = 3, position = pd) +
  labs(
    title = "Hypothesis 6: Bias in Reconstruction Errors (Exp. 1)",
    subtitle = "Trials where original choice revealed gems, but reconstruction was an error",
    x = "Trial Bin",
    y = "Probability of Choosing Stones-Only Location",
    color = "Participant Group"
  ) +
  theme_minimal() +
  theme(
    legend.position = "bottom",
    plot.title = element_text(face = "bold")
  )

print(p)

library(dplyr)
library(ggplot2)
library(scales)

# ---- Prepare data frames ----
# h6_viz_data must be subject-level per bin (one row per subj_id x mentioned_rule x trial_bin)
# If your h6_viz_data has NA prob_stones for empty bins, filter them out for plotting points/stats.

h6_subject_bin <- h6_viz_data %>%
  # keep only real subject rows (drop rows where prob_stones is NA because n == 0)
  filter(!is.na(prob_stones)) %>%
  select(subj_id, mentioned_rule, trial_bin, prob_stones, Group)

# summary_df: mean ± SE **across subjects** per trial_bin × Group
summary_df <- h6_subject_bin %>%
  group_by(trial_bin, Group) %>%
  summarise(
    mean_prob = mean(prob_stones, na.rm = TRUE),
    n_subj   = sum(!is.na(prob_stones)),
    se       = ifelse(n_subj > 1,
                      sd(prob_stones, na.rm = TRUE) / sqrt(n_subj),
                      NA_real_),   # se undefined for n=1, set to NA
    .groups = "drop"
  )

# ---- Plot ----
pd <- position_dodge(width = 0.75)

ggplot(h6_subject_bin, aes(x = trial_bin, y = prob_stones, fill = Group, group = Group)) +

  # jittered subject-level points (helps see spread)
  geom_jitter(aes(color = Group),
              position = position_jitterdodge(jitter.width = 0.12, dodge.width = 0.75),
              size = 1.6, alpha = 0.6, show.legend = FALSE) +

  # overlay mean ± SE from summary_df (black errorbars and points)
  geom_errorbar(data = summary_df,
                aes(x = trial_bin, y = mean_prob, ymin = mean_prob - se, ymax = mean_prob + se, group = Group),
                width = 0.18,
                position = pd,
                color = "black",
                linewidth = 0.8,
                na.rm = TRUE) +

  # connect group means with colored lines
  geom_line(data = summary_df,
            aes(x = trial_bin, y = mean_prob, color = Group, group = Group),
            position = pd,
            linewidth = 1.0,
            alpha = 0.8,
            na.rm = TRUE) +

  # plot mean points
  geom_point(data = summary_df,
             aes(x = trial_bin, y = mean_prob, color = Group, group = Group),
             position = pd,
             size = 3,
             shape = 21,
             fill = "white",
             stroke = 1,
             na.rm = TRUE) +

  # reference line at 0.5
  geom_hline(yintercept = 0.5, linetype = "dashed", color = "black", linewidth = 0.7) + 

  labs(
    title = "Exp. 1: Subject-level probability of choosing stones by trial bin",
    x = "Trial bin",
    y = "Probability of choosing stones (per subject)",
    fill = "Group",
    color = "Group"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    legend.position = "right",
    panel.grid.minor = element_blank()
  )


h6_subj_summary <- h6_df %>%
  group_by(subj_id, mentioned_rule) %>%
  summarise(
    mean_trial_index = mean(trial_index, na.rm = TRUE),    # x coordinate
    prob_stones = mean(chose_onlystones, na.rm = TRUE),    # y coordinate
    n_trials = n(),
    .groups = "drop"
  ) %>%
  mutate(Group = ifelse(mentioned_rule, "Position mentioned", "Position NOT mentioned"))

# Scatter: one point per participant
ggplot(h6_subj_summary, aes(x = n_trials, y = prob_stones, color = Group)) +
  geom_point(size = 3, alpha = 0.85) +
  geom_smooth(method = "lm", se = FALSE, linetype = "dashed", linewidth = 0.7, aes(group = Group)) +
  geom_hline(yintercept = 0.5, linetype = "dashed", color = "red", linewidth = 0.7) +
  scale_y_continuous(labels = percent_format(accuracy = 1), limits = c(0, 1)) +
  scale_x_continuous(breaks = pretty(h6_subj_summary$n_trials)) +
  labs(
    title = "Per-participant probability of choosing stones-only (one point per participant)",
    x = "Number of trials per participant",
    y = "Probability of choosing stones-only",
    color = "Group"
  ) +
  theme_minimal(base_size = 13)

```


Hypothesis 6 (POSITION BIAS IN RECONSTRUCTION ERRORS): We tested the probability of choosing the stones-only position on trials in which participants' guesses in part 2 were inaccurate relative to their decision in part 1, which revealed gems. This deconfounds the bias against the stones-only side from a bias to be correct, or from a memory for originally biased actions.

We fitted a mixed-effects logistic regression with a random intercept for participant. The estimated probability of selecting the stones-only position was 
`r printnum(round(plogis(fixef(h6_model)[1]), 2))`, 
corresponding to a log-odds estimate of 
β = `r printnum(round(fixef(h6_model)[1], 2))`,  (
SE = `r printnum(round(summary(h6_model)$coefficients[1, "Std. Error"], 2))`, 
*z* = `r printnum(round(summary(h6_model)$coefficients[1, "z value"], 2))`, 
*p* `r printp(summary(h6_model)$coefficients[1, "Pr(>|z|)"])`)


For participants who mentioned the rule the estimated probability of selecting the stones-only position was `r printnum(round(unname(1 / (1 + exp(-fixef(r_h6_model)))), 2))` (β = `r printnum(round(fixef(r_h6_model)[1], 2))`,  (
SE = `r printnum(round(summary(r_h6_model)$coefficients[1, "Std. Error"], 2))`, 
*z* = `r printnum(round(summary(r_h6_model)$coefficients[1, "z value"], 2))`, 
*p* `r printp(summary(r_h6_model)$coefficients[1, "Pr(>|z|)"])`)

For participants who did not mention the rule, the estimated probability of selecting the stones-only position was `r printnum(round(unname(1 / (1 + exp(-fixef(nr_h6_model)))), 2))` (β = `r printnum(round(fixef(nr_h6_model)[1], 2))`,  (
SE = `r printnum(round(summary(nr_h6_model)$coefficients[1, "Std. Error"], 2))`, 
*z* = `r printnum(round(summary(nr_h6_model)$coefficients[1, "z value"], 2))`, 
*p* `r printp(summary(nr_h6_model)$coefficients[1, "Pr(>|z|)"])`)

The difference between the two groups (those who mentioned and did not mention the rule) is significant, with *mentioned_rule* included as a fixed effect in the model. Participants who mentioned the rule were less likely to choose the stones-only option than those who did not mention the rule (β =`r printnum(round(summary(h6_model_mentioned_vs_notmentioned)$coefficients["mentioned_ruleTRUE", "Estimate"],2))`, SE = `r printnum(round(summary(h6_model_mentioned_vs_notmentioned)$coefficients["mentioned_ruleTRUE", "Std. Error"],2))`, z = `r printnum(round(summary(h6_model_mentioned_vs_notmentioned)$coefficients["mentioned_ruleTRUE", "z value"],2))`, p `r printp(summary(h6_model_mentioned_vs_notmentioned)$coefficients["mentioned_ruleTRUE","Pr(>|z|)"][1])`)




### Reaction times

```{r reaction times in part 2, include = FALSE, cache=F}

RT2_by_gems <- two_parts_df %>%
  mutate(correct2 = participant_response_position_e2 == correct_choice_position) %>%
  group_by(subj_id,mentioned_rule,correct2) %>%
  summarise(RT=median(RT2)) %>%
  spread(correct2,RT,sep='') %>%
  mutate(diff=correct2TRUE-correct2FALSE)

RT2_by_gems_consistent_only <- two_parts_df %>%
  filter(is_consistent) %>%
  mutate(correct2 = participant_response_position_e2 == correct_choice_position) %>%
  group_by(subj_id,mentioned_rule,correct2) %>%
  summarise(RT=median(RT2)) %>%
  spread(correct2,RT,sep='') %>%
  mutate(diff=correct2TRUE-correct2FALSE)

RT2_by_consistency <- two_parts_df %>%
  group_by(subj_id,mentioned_rule,is_consistent) %>%
  summarise(RT=median(RT2)) %>%
  spread(is_consistent,RT,sep='') %>%
  mutate(diff=is_consistentTRUE-is_consistentFALSE)


two_parts_df %>%
  mutate(correct2 = participant_response_position_e2 == correct_choice_position) %>%
  group_by(subj_id,mentioned_rule,correct, correct2) %>%
  summarise(RT=median(RT2)) %>%
  mutate(condition = factor(ifelse(correct & correct2, 'hit then hit',
                            ifelse(correct & !correct2, 'hit then miss',
                                   ifelse(!correct & !correct2, 'miss then miss',
                                          'miss then hit'))),
                            levels=c('hit then hit','miss then miss',
                                     'miss then hit','hit then miss'))) %>%
  ggplot(aes(x=condition,y=RT)) +
  geom_boxplot()+
  geom_jitter(width=0.15,alpha=0.5)
```

```{r, median rt plot, include = FALSE, cache=F}

plot_data_rt_raw <- two_parts_df %>%
  filter(is_consistent, !is_catch_trial) %>%
  mutate(
    # Create 'choice_type' based on the Phase 2 response columns
    choice_type = case_when(
      participant_response_position_e2 == correct_choice_position ~ "Chose Gems",
      participant_response_position_e2 == rule_condition        ~ "Chose Only Stones",
      TRUE                                                      ~ "Chose Other" 
    ),
    choice_type = factor(choice_type, levels = c("Chose Gems", "Chose Other", "Chose Only Stones"))
  ) %>%
  filter(!is.na(choice_type))

rt_boxplot <- ggplot(plot_data_rt_raw, 
                     aes(x = choice_type, y = RT2, fill = choice_type)) +
  geom_boxplot(notch = TRUE, show.legend = FALSE, alpha = 0.8) +
  
  scale_fill_manual(
    values = c(
      "Chose Gems" = "forestgreen",
      "Chose Only Stones" = "firebrick",
      "Chose Other" = "royalblue"
    )
  ) +
  
  scale_y_continuous(
    labels = scales::label_comma(), 
    limits = c(0, 8000),        
    breaks = seq(0, 8000, by = 1000) 
  ) +
  
  labs(
    title = "Median RT by Choice (Consistent Trials)",
    subtitle = "(Notches represent ~95% Confidence Interval around the median)",
    x = "Participant's Choice",
    y = "Reaction Time (ms)" 
  ) +
  theme_minimal(base_size = 12) +
  theme(plot.title = element_text(face = "bold"))


print(rt_boxplot)
```

```{r rt plot alternative, fig.width=8}
library(dplyr)
library(ggplot2)

plot_data <- two_parts_df %>%
  filter(is_consistent, !is_catch_trial) %>%
  mutate(
    choice_type = case_when(
      participant_response_position_e2 == correct_choice_position ~ "Chose Gems",
      participant_response_position_e2 == rule_condition        ~ "Chose Only Stones",
      TRUE                                                     ~ "Chose Other"
    ),
    choice_type = factor(choice_type, levels = c("Chose Gems", "Chose Other", "Chose Only Stones"))
  ) %>%
  filter(!is.na(choice_type)) %>%
  group_by(subj_id, choice_type) %>%
  summarise(median_rt = median(RT2, na.rm = TRUE), .groups = "drop")

# summary for mean +/- SEM
summary_data <- plot_data %>%
  group_by(choice_type) %>%
  summarise(
    mean_rt = mean(median_rt, na.rm = TRUE),
    sem = sd(median_rt, na.rm = TRUE) / sqrt(n()),
    .groups = "drop"
  )

choice_colors <- c(
  "Chose Gems" = "forestgreen",
  "Chose Other" = "royalblue",
  "Chose Only Stones" = "firebrick"
)

# plot
ggplot(plot_data, aes(x = choice_type, y = median_rt, fill = choice_type)) +
  geom_boxplot(width = 0.6, alpha = 0.8, color = "gray30", outlier.shape = NA) +
  geom_errorbar(
    data = summary_data,
    mapping = aes(x = choice_type, ymin = mean_rt - sem, ymax = mean_rt + sem),
    width = 0.15, size = 0.6, inherit.aes = FALSE
  ) +
  geom_point(
    data = summary_data,
    mapping = aes(x = choice_type, y = mean_rt),
    shape = 21, size = 3.5, fill = "white", color = "black", stroke = 1,
    inherit.aes = FALSE
  ) +
  scale_fill_manual(values = choice_colors) +
  labs(
    title = "Median Reaction Times by Choice Type",
    x = "Participant's choice",
    y = "Median Reaction Time (ms)"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold"),
    legend.position = "none"
  )
```

In part 2, participants were slower by `r RT2_by_gems$diff%>%median()%>%abs()%>%round()` ms to reconstruct their actions when they reported choosing the box that did not have gems in it compared to when reporting choosing the box that did have gems in it (`r apa_print(RT2_by_gems$diff%>%t.test())$statistic`). When focusing on trials in which participants correctly reconstructed their action, this difference was `r RT2_by_gems_consistent_only$diff%>%median()%>%abs()%>%round()` ms (`r apa_print(RT2_by_gems_consistent_only$diff%>%t.test())$statistic`).

Participants were slower by `r RT2_by_consistency$diff%>%median()%>%abs()%>%round()` ms when incorrectly reconstructing their actions compared to when their reconstructions were accurate (`r apa_print(RT2_by_consistency$diff%>%t.test())$statistic`).

```{r phase 2 slopes, include = TRUE, cache=F}

phase_2_slopes <- two_parts_df %>%
  group_by(subj_id) %>%
  filter(!is_catch_trial) %>%
  mutate(chose_stone_side = chosen_side_part2 == 'only_stones') %>%
  do(tidy(glm(chose_stone_side ~ trial_index, family = binomial, data = .))) %>%
  filter(term == "trial_index") %>%
  select(subj_id, slope = estimate)

learning_effect <- two_parts_df %>%
  mutate(half=ifelse(trial_index<31,'first','second')) %>%
  group_by(subj_id,half) %>%
  filter(!is_catch_trial) %>%
  mutate(chose_stone_side = chosen_side_part2 == 'only_stones') %>%
  summarise(p=mean(chose_stone_side)) %>%
  spread(half,p) %>%
  mutate(diff=first-second) %>%
  pull(diff) %>% 
  t.test()

part1_choices <- phase_1_df %>%
  select(subj_id, trial_index, part1_chose_gems = correct)

phase_2_merged_df <- two_parts_df %>%
  left_join(part1_choices, by = c("subj_id", "trial_index"))

data_for_test <- phase_2_merged_df %>%
  filter(!is_catch_trial, !is.na(part1_chose_gems)) %>%
  group_by(subj_id)
  

data_for_test_chosegems <- data_for_test %>%
  filter(part1_chose_gems)

data_for_test_chosenongems <- data_for_test %>%
  filter(!part1_chose_gems)

# t.test(x=data_for_test_chosegems$is_consistent, y=data_for_test_chosenongems$is_consistent, paired=TRUE)
# 
# cohen.d(as.numeric(is_consistent) ~ part1_chose_gems, data = data_for_test)



consistency_summary <- two_parts_df %>%
  filter(!is_catch_trial) %>%
  group_by(subj_id, correct) %>% # 'correct' is a boolean for Part 1 accuracy
  summarise(mean_consistency = mean(is_consistent, na.rm = TRUE), .groups = "drop") %>%
  pivot_wider(names_from = correct,
              names_prefix = "consistency_when_p1_",
              values_from = mean_consistency)

# t.test(x = consistency_summary$consistency_when_p1_TRUE,
#        y = consistency_summary$consistency_when_p1_FALSE,
#        paired = TRUE)


# 
# cohen.d(as.numeric(is_consistent) ~ correct,
#         data = two_parts_df %>% filter(!is_catch_trial),
#         pooled = TRUE, paired = FALSE)


first_trial_data <- two_parts_df %>%
  filter(trial_index == 1)

correct_t1_ids <- first_trial_data %>%
  filter(correct == TRUE) %>%
  pull(subj_id)

incorrect_t1_ids <- first_trial_data %>%
  filter(correct == FALSE) %>%
  pull(subj_id)
  
t1_correct <- two_parts_df %>%
  filter(subj_id %in% correct_t1_ids) %>%
  filter(trial_index != 1) %>%
  group_by(subj_id) %>%
  summarise(mean_acc = mean(correct))
  
t1_incorrect  <- two_parts_df %>%
  filter(subj_id %in% incorrect_t1_ids) %>%
  filter(trial_index != 1) %>%
  group_by(subj_id) %>%
  summarise(mean_acc = mean(correct))

t.test(t1_correct$mean_acc, t1_incorrect$mean_acc)


```

Like in part 1, the probability of selecting the only-stones option did not decrease as a function of trial number (`r apa_print(phase_2_slopes$slope%>%t.test())$statistic`).

```{r selected_box_number, include = FALSE, cache=F}

two_parts_df %>%
  group_by(trial_index) %>%
  filter(!is_catch_trial) %>%
  summarise(n1 = mean(number1),
         n2 = mean(number2),
         se1 = se(number1),
         se2 = se(number2)) %>%
  pivot_longer(
    cols = c(n1, n2, se1, se2),
    names_to = c(".value", "part"),
    names_pattern = "([a-z]+)([0-9]+)"
  ) %>%
  ggplot(aes(x=trial_index,y=n,color=part,fill=part))+
  geom_line()+
  geom_ribbon(aes(ymin=n-se,ymax=n+se),alpha=0.5)

number_df <- two_parts_df %>% 
  group_by(subj_id) %>%
  summarise(n1=mean(number1), 
            n2=mean(number2)) %>%
  mutate(diff=n1-n2) # no difference
  

  # do(tidy(glm(chose_stone_side ~ trial_index, family = binomial, data = .))) %>%
  # filter(term == "trial_index") %>%
  # select(subj_id, slope = estimate)

learning_effect <- two_parts_df %>%
  mutate(half=ifelse(trial_index<31,'first','second')) %>%
  group_by(subj_id,half) %>%
  filter(!is_catch_trial) %>%
  mutate(chose_stone_side = chosen_side_part2 == 'only_stones') %>%
  summarise(p=mean(chose_stone_side)) %>%
  spread(half,p) %>%
  mutate(diff=first-second) %>%
  pull(diff) %>% 
  t.test()


conditions <- phase_1_df %>%
  
  # This is the fix: It removes any pre-existing grouping
  ungroup() %>% 
  
  # This gets one row per participant (which you had correct)
  distinct(subj_id, rule_condition) %>% 
  
  # This will now count all participants together
  count(rule_condition)

print(conditions)

combined_prolific_path <- "/Users/sawakeishiro/Desktop/code/reconstruction/analysis/exp1/prolific_demographic_export_68d4087eea848a9d64b857a5.csv"

tryCatch({
  # 2. Read the combined prolific data
  prolific_data <- read_csv(combined_prolific_path)
  
  # 3. IMPORTANT: Load your raw_df dataframe
  # This script assumes 'raw_df' is already loaded in your R session.
  # If not, you must load it here, for example:
  # raw_df <- read_csv("/path/to/your/raw_data.csv")
  
  # Check if raw_df exists
  if (!exists("raw_df")) {
    stop("The dataframe 'raw_df' was not found. Please load it before running this merge.")
  }
  
  # 4. Merge the dataframes
  # We use inner_join() which keeps only the rows where the ID exists
  # in *both* dataframes, as you requested.
  merged_data <- inner_join(
    prolific_data, 
    two_parts_df, 
    by = c("Participant id" = "subj_id")
  )
  
  # 5. Save the new merged dataframe
  write_csv(merged_data, merged_output_path)
  
  # 6. Print a success message
  print(paste("Successfully merged data and saved to:", merged_output_path))
  print(paste("Original prolific rows:", nrow(prolific_data)))
  print(paste("Original raw_df rows:", nrow(raw_df)))
  print(paste("Total merged rows:", nrow(merged_data)))

}, error = function(e) {
  # Handle potential errors
  print(paste("An error occurred:", e$message))
})




# --- 1. DEFINE PARTICIPANT-LEVEL DATA ---

# Accuracy
participant_accuracy_1 <- merged_data %>%
  filter(!is_catch_trial) %>%
  group_by(`Participant id`, Age) %>%
  summarise(participant_acc = mean(correct), .groups = 'drop')

participant_accuracy_2 <- merged_data %>%
  filter(!is_catch_trial) %>%
  group_by(`Participant id`, Age) %>%
  # FIX: The .groups = 'drop' goes on summarise(), not mean()
  summarise(participant_acc = mean(is_consistent), .groups = 'drop')

# Avoidance
age_avoid_1 <- merged_data %>%
  filter(!is_catch_trial) %>%
  group_by(`Participant id`, Age) %>%
  summarise(participant_avoid = mean(chosen_side == 'only_stones'), .groups = 'drop')

age_avoid_2 <- merged_data %>%
  filter(!is_catch_trial) %>%
  group_by(`Participant id`, Age) %>%
  summarise(participant_avoid = mean(chosen_side_part2 == 'only_stones'), .groups = 'drop')

# --- 2. DEFINE AGE BINS ---
age_breaks <- c(17, 22, 27, 32, 37, 42, 47, 52, 57, 62, 67, 72, 77, Inf)
age_labels <- c("18-22", "23-27", "28-32", "33-37", "38-42", "43-47",
                "48-52", "53-57", "58-62", "63-67", "68-72", "73-77", "78+")

# --- 3. PROCESS AND PLOT ACCURACY ---

# Combine participant accuracy data
combined_participant_data <- bind_rows(
  participant_accuracy_1 %>% mutate(Condition = "Part 1"),
  participant_accuracy_2 %>% mutate(Condition = "Part 2")
)

# Bin and summarise accuracy data
combined_age_summary <- combined_participant_data %>%
  mutate(
    Age = as.numeric(as.character(Age)),
    AgeBin = cut(Age, breaks = age_breaks, labels = age_labels, right = TRUE, ordered_result = TRUE)
  ) %>%
  filter(!is.na(AgeBin)) %>%
  group_by(AgeBin, Condition) %>% 
  summarise(
    mean_accuracy = mean(participant_acc),
    se = sd(participant_acc) / sqrt(n()),
    .groups = 'drop'
  )

# Plot accuracy (with CORRECTED scale_color_manual)
ggplot(combined_age_summary, aes(x = AgeBin, 
                                 y = mean_accuracy, 
                                 group = Condition, 
                                 color = Condition)) +
  geom_line(linewidth = 1.2, position = position_dodge(width = 0.3)) +
  geom_point(size = 3, position = position_dodge(width = 0.3)) +
  geom_errorbar(
    aes(ymin = mean_accuracy - se, ymax = mean_accuracy + se),
    width = 0.25, linewidth = 0.8, position = position_dodge(width = 0.3)
  ) +
  
  # --- THE FIX ---
  # Names here MUST match the names in your 'Condition' column
  scale_color_manual(values = c("Part 1" = "blue", "Part 2" = "red")) +
  
  labs(
    title = "Mean Participant Accuracy by Age Bin",
    x = "Age Bin",
    y = "Mean Accuracy",
    color = "Phase" # Renamed legend title
  ) +
  coord_cartesian(ylim = c(0, 1)) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))


# --- 4. PROCESS AND PLOT AVOIDANCE ---
# (I am re-using the "combine-first" method here, as it's cleaner)

# Combine participant AVOIDANCE data
combined_avoid_data <- bind_rows(
  age_avoid_1 %>% mutate(Condition = "Part 1"),
  age_avoid_2 %>% mutate(Condition = "Part 2")
)

# Bin and summarise AVOIDANCE data
combined_age_summary_avoid <- combined_avoid_data %>%
  mutate(
    Age = as.numeric(as.character(Age)),
    AgeBin = cut(Age, breaks = age_breaks, labels = age_labels, right = TRUE, ordered_result = TRUE)
  ) %>%
  filter(!is.na(AgeBin)) %>%
  group_by(AgeBin, Condition) %>%
  summarise(
    # FIX: Use a non-confusing name like mean_avoid
    mean_avoid = mean(participant_avoid), 
    se = sd(participant_avoid) / sqrt(n()),
    .groups = 'drop'
  )

# Plot avoidance (with CORRECTED variable names and color scale)
ggplot(combined_age_summary_avoid, aes(x = AgeBin, 
                                       # FIX: Plot the new mean_avoid column
                                       y = mean_avoid, 
                                       group = Condition, 
                                       color = Condition)) +
  geom_line(linewidth = 1.2, position = position_dodge(width = 0.3)) +
  geom_point(size = 3, position = position_dodge(width = 0.3)) +
  geom_errorbar(
    # FIX: Use mean_avoid for error bars
    aes(ymin = mean_avoid - se, ymax = mean_avoid + se), 
    width = 0.25, linewidth = 0.8, position = position_dodge(width = 0.3)
  ) +
  
  # --- THE FIX ---
  # Names here MUST match the names from combined_avoid_data
  scale_color_manual(values = c("Part 1" = "blue", "Part 2" = "red")) +
  
  labs(
    title = "Mean Participant Avoidance by Age Bin",
    x = "Age Bin",
    y = "Mean Avoidance", # FIX: Correct y-axis label
    color = "Phase"
  ) +
  coord_cartesian(ylim = c(0, 1)) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

participant_age_data <- merged_data %>%
  mutate(Age_numeric = as.numeric(as.character(Age))) %>%
  filter(!is.na(Age_numeric)) %>%
  distinct(`Participant id`, Age_numeric)


ggplot(participant_age_data, aes(x = Age_numeric)) +
  # binwidth = 5 will group ages (e.g., 18-22, 23-27...)
  geom_histogram(binwidth = 5, fill = "cornflowerblue", color = "white", alpha = 0.9, boundary = 0) +
  labs(
    title = "Distribution of Participant Age",
    x = "Age",
    y = "Frequency (Count of Participants)" # Updated label
  ) +
  theme_minimal()


```




```{r gems, cache=F}

library(lme4)

gems_data <- phase_2_df %>%
  filter(!is_catch_trial) %>%
  filter(!chose_gems_p1, !is_consistent) %>%
  mutate(chose_gems_p2 = participant_response_position_e2 == correct_choice_position_p1)

model <- glmer(
  chose_gems_p2 ~ 1 + (1 | subj_id),
  data = gems_data,
  family = binomial,
  control = glmerControl(
    optimizer = "bobyqa",)
)

summary(model)

plogis(fixef(model)[1])


```


We ran a mixed-effects logistic regression model, for the probability that participants chose gems in part 2 given that they did not choose gems in part 1 and incorrectly reconstruct in part 2. Participants choose the option with gems with a probability of `r printnum(round(plogis(fixef(model)[1]), 2))`, significantly higher than chance (β = `r printnum(round(fixef(model)[1], 2))`,  
SE = `r printnum(round(summary(model)$coefficients[1, "Std. Error"], 2))`, 
*z* = `r printnum(round(summary(model)$coefficients[1, "z value"], 2))`, 
*p* `r printp(summary(model)$coefficients[1, "Pr(>|z|)"])`)


```{r prepare_data, cache=F}

chose_onlystones_df_exp1 <- phase_2_df %>%
  filter(!is_catch_trial) %>%
  mutate(
    chose_gems_p2 = participant_response_position_e2 == correct_choice_position_p1,
    chose_onlystones_p2 = rule_condition == participant_response_position_e2,
    trial_num = trial_index
         ) %>%
  select(chose_onlystones_p1, chose_onlystones_p2, chose_gems_p2, subj_id, trial_num)
  

write.csv(
  chose_onlystones_df_exp1,
  "chose_onlystones_df_exp1.csv",
  row.names = FALSE
)




```

```{r transfer_data, cache=F}

discovered_rule <- rule_check_df %>%
  filter(subj_id %in% c(
    '60fe7bf61c7567f5cca9ade6',
    '5ecf9b46617a9b3dd56b3fe3',
    '5ebc0abd2cc5f1054c82a0d7',
    '654bc148c95bca00d541ec33',
    '66dcc0e54f44ccecf2730c76',
    '631d197eba06364cbea7fab2',
    '66fd3f6c93c24ecc119be354',
    '66e9236f775eae67be04229d',
    '66b244a2f436437d3e109bbe',
    '63ee9abfa262bea95766dd9c',
    '66cfab152c56aa0472529001',
    '5bf368846edd7200017136c7',
    '66d9cd06e207a93377f7820e',
    '5962a5b29955ee00015bce98',
    '632d189d6b8f03d0d3db4c48',
    '63b3377ca1ccdfc7cc567503',
    '6658bba1a51a037d25bb677e',
    '662945728093113a43ab8077',
    '676058418c08c71e4bb7cbd1',
    '58954e2dc88c680001dd4b7e',
    '681b781064a5701a5ceee3bc',
    '682b3f63d0e2c4eff64fe664',
    '5d3695969749cc00165c222f',
    '6773dea609f77d473900997a',
    '60e3a25e8be835f754795d93') &
      !(subj_id %in% participants_to_exclude)) 

keys <- c("subj_id", "trial_num")

names1 <- names(phase_1_df)
names2 <- names(phase_2_df)


common_names <- intersect(names1, names2)
common_names
common_nonkey <- setdiff(common_names, keys)
common_nonkey

phase_2_df %>% inner_join(raw_df %>% filter(phase=="part2")%>% select(rt, trial_index, subj_id), by = c("trial_index", "subj_id")) %>%
      mutate(rt_p2 = rt) 

two_parts_df <- phase_1_df %>%
  inner_join(
    phase_2_df %>%
      
    select(-c(
      is_catch_trial,
      rule_condition,
      chose_gems_p1, 
      chose_onlystones_p1
    )),
    by = c("subj_id", "trial_index")) %>%
  mutate(
    trial_num = trial_index,
    chose_gems = correct_choice_position == participant_response_position,
    chose_onlystones = rule_condition == participant_response_position,
    chose_other = !chose_gems & !chose_onlystones,
    response_position = participant_response_position,
    stoneposition = rule_condition,
    chose_gems_p2 = correct_choice_position_p1 == participant_response_position_e2,
    chose_onlystones_p2 = rule_condition == participant_response_position_e2,
    chose_other_p2 = !chose_gems_p2 & !chose_onlystones_p2,
    response_position_p2 = participant_response_position_e2,
    correct_choice_position_p2 = correct_choice_position,
    
    
  ) %>%
  mutate(mentioned_rule = subj_id %in% discovered_rule$subj_id) %>%
  select(-c(
    trial_index,
    participant_response_position,
    rule_condition,
    RT,
    correct,
    chosen_side,
    chose_gems_p1,
    chose_onlystones_p1,
    participant_response_position,
    RT2,
    reconstructed_action,
    chosen_side_part2,
    correct_choice_position_p1
  ))

write.csv(
  two_parts_df,
  "two_parts_df_exp1.csv",
  row.names = FALSE
)

```

