---
title: "Experiment 2: results"
output: word_document
date: "i forgot, think it was 2025"
---

```{r setup, include = FALSE}
library("papaja")
r_refs("r-references.bib")

library('groundhog')
groundhog.library(
  c(
    'papaja',
    'tidyverse'
  ), "2025-06-01"
)

r_refs("r-references.bib")

knitr::opts_chunk$set(warning=F,echo=F,message=F,cache=T)
```

```{r analysis-preferences}
# Seed for random number generation
set.seed(42)
knitr::opts_chunk$set(cache.extra = knitr::rand_seed)
```


```{r loadformat, include = FALSE, cache=F}
library(tidyverse)
library(stringr)
library(effsize)


file_list <- '/Users/sawakeishiro/Desktop/code/reconstruction/data/Experiment2/experimen2_data.txt'

raw_df <- map_dfr(file_list, function(file) {

  lines <- readLines(file)
  starts <- which(grepl('^"?\\s*success', lines))
  starts <- c(starts, length(lines) + 1)
  chunks <- map2(starts[-length(starts)], starts[-1] - 1, ~ lines[.x:.y])
  map_dfr(chunks, ~ read_csv(
      paste(.x, collapse = "\n"),
      show_col_types = FALSE,
      col_types = cols(.default = "c") 
    ))  %>%
  # map_dfr(chunks, ~ read_csv(
  #   paste(.x, collapse = "\n"),
  #   show_col_types = FALSE,
  #   col_types = cols(.default = "c")   
    mutate(
      subj_id = PROLIFIC_PID,
      is_catch_trial = as.logical(is_catch_trial),
      is_show_trial = as.logical(is_show_trial),
      response = as.numeric(response),
      chose_gems = as.logical(chose_gems),
      chose_onlystones = as.logical(chose_onlystones),
      chose_other = as.logical(chose_other),
      left_number = as.numeric(left_number),
      middle_number = as.numeric(middle_number),
      right_number = as.numeric(right_number),
      cumulative_score = as.numeric(cumulative_score),
      is_consistent = as.logical(is_consistent),
      trial_num = as.numeric(trial_num),
      rt = as.numeric(rt),
      confidence_rt = as.numeric(confidence_rt),
      comprehension_failures_part1 = as.numeric(comprehension_failures_part1),
      comprehension_failures_part2_1 = as.numeric(comprehension_failures_part2_1),
      comprehension_failures_part2_2 = as.numeric(comprehension_failures_part2_2),
      comprehension_failures_part2_3 = as.numeric(comprehension_failures_part2_3),
      comprehension_failures_part2_4 = as.numeric(comprehension_failures_part2_4),
      confidence = as.numeric(confidence),
      rule_check = as.character(rule_check),
      worker_comments = as.character(worker_comments),
      number_correct_part1 = as.numeric(number_correct_part1_slider),
      number_correct_part2 = as.numeric(number_correct_part2_slider)
    )
})

comments_sparse_df <- raw_df %>%
  dplyr::select(subj_id, rule_check, worker_comments, number_correct_part1, number_correct_part2) %>%
  dplyr::filter(!is.na(worker_comments) | !is.na(rule_check) | !is.na(number_correct_part1) | !is.na(number_correct_part2))

comments_clean_df <- comments_sparse_df %>%
  group_by(subj_id) %>%
  summarise(
    # For each column, find the first non-NA value for that subject
    rule_check = first(na.omit(rule_check)),
    worker_comments = first(na.omit(worker_comments)),
    number_correct_part1 = first(na.omit(number_correct_part1)),
    number_correct_part2 = first(na.omit(number_correct_part2)),
    .groups = 'drop' 
  )

print(head(comments_clean_df))

write.csv(comments_clean_df, 'comments_exp2_clean.csv', row.names = FALSE)
  
  


# Process data for Phase 1
phase_1_df <- raw_df %>%
  filter(phase == 'part1') %>%
  dplyr::select(
    subj_id,
    trial_num,
    left_word,
    left_number,
    middle_word, 
    middle_number,
    right_word, 
    right_number, 
    chose_gems,
    chose_onlystones,
    chose_other,
    response_position,
    correct_choice_position,
    response_position,
    is_catch_trial, 
    is_show_trial,
    stoneposition,
    emeraldposition,
    rt,
    comprehension_failures_part1,
    number_correct_part1) %>%
  group_by(subj_id) %>%
  arrange(trial_num) 
  




write.csv(raw_df, 'raw_df.csv')
phase_2_df <- raw_df %>%
  filter(phase == 'part2') %>%
  dplyr::select(
    subj_id, 
    trial_num,
    response_position,
    is_consistent,
    correct_choice_position,
    confidence,
    chose_gems,
    chose_onlystones,
    chose_other,
    rt,
    confidence_rt,
    is_catch_trial,
    is_show_trial,
    stoneposition,
    emeraldposition,
    left_word,
    left_number,
    middle_word, 
    middle_number,
    right_word, 
    right_number, 
    comprehension_failures_part2_1,
    comprehension_failures_part2_2,
    comprehension_failures_part2_3,
    comprehension_failures_part2_4,
    number_correct_part2
  ) %>%
  group_by(subj_id) %>%
  arrange(trial_num)  %>%
  filter(!is.na(is_consistent))



#excluding participants

total_participants_before <- length(unique(phase_1_df$subj_id))
cat("Total number of participants before exclusion:", total_participants_before)

exclude_catch <- phase_1_df %>%
  group_by(subj_id) %>%
  summarise(
    n_failed_catch = sum(!chose_gems[is_catch_trial == TRUE], na.rm = TRUE),)%>%
  filter(n_failed_catch >= 2) %>%
  pull(subj_id)

exclude_part1_rt <- phase_1_df %>%
  group_by(subj_id) %>%
  summarise(
  first_quartile_rt = quantile(rt, probs = 0.25, na.rm = TRUE))%>%
  filter(first_quartile_rt < 100 ) %>%
  pull(subj_id)

exclude_part2_rt <- phase_2_df %>%
  group_by(subj_id) %>%
  summarise(
    first_quartile_rt = quantile(rt, probs = 0.25, na.rm = TRUE)) %>%
  filter( first_quartile_rt < 100 ) %>%
  pull(subj_id)

exclude_part2_confidence_rt <- phase_2_df %>%
  group_by(subj_id) %>%
  summarise(
    first_quantile_rt = quantile(confidence_rt, probs = 0.25, na.rm = TRUE)
  ) %>%
  filter(first_quantile_rt < 100) %>%
  pull(subj_id)
  

exclude_comprehension <- raw_df %>%
  filter(comprehension_failures_part1 >= 2 | comprehension_failures_part2_1 >= 2 |  comprehension_failures_part2_2 >= 2| comprehension_failures_part2_3 >= 2| comprehension_failures_part2_4 >= 2) %>%
  distinct(subj_id) %>%
  pull(subj_id)


participants_to_exclude <- unique(c(exclude_catch,
                                    exclude_part1_rt, 
                                    exclude_part2_rt, 
                                    exclude_part2_confidence_rt,
                                    exclude_comprehension))


cat("Number of participants to exclude:", length(participants_to_exclude), "\n")
if (length(participants_to_exclude) > 0) {
  cat("Excluding participants:", paste(participants_to_exclude, collapse=", "), "\n")
}



phase_1_df <- phase_1_df %>%
  filter(!(subj_id %in% participants_to_exclude))

phase_2_df <- phase_2_df %>%
  filter(!(subj_id %in% participants_to_exclude))

total_participants_after <- length(unique(phase_1_df$subj_id))
cat("Total number of participants after exclusion:", total_participants_after, "\n")


two_parts_df <- phase_1_df %>%
  inner_join(
    phase_2_df,
    by = c("subj_id", "trial_num"),
    suffix = c("_part1", "_part2")
  )

# rank participants based on score:
top_ids <- raw_df %>%
  dplyr::select(
    subj_id, 
    cumulative_score
) %>%
  filter(!subj_id %in% participants_to_exclude) %>%
  group_by(subj_id) %>%
  summarise(max_cumulative_score = max(cumulative_score, na.rm = TRUE)) %>%
  arrange(desc(max_cumulative_score)) %>%
  slice_head(n = 30) %>% # 20*0.3 = 6
  pull(subj_id)

print(top_ids)




```

```{r demographic, include = FALSE, cache=F}

library(readr)
library(dplyr)

file1_path <- "/Users/sawakeishiro/Desktop/code/reconstruction/data/Experiment2/prolific_demographic_export_68f7a58a11e006757d9bc651.csv"
file2_path <- "/Users/sawakeishiro/Desktop/code/reconstruction/data/Experiment2/prolific_demographic_export_690240fa59fb33d0715ea0dd.csv"
output_dir <- "/Users/sawakeishiro/Desktop/code/reconstruction/data/Experiment2/"
output_file_path <- file.path(output_dir, "prolific_demographic_combined.csv")

tryCatch({
  data1 <- read_csv(file1_path)
  data2 <- read_csv(file2_path)

  combined_data <- bind_rows(data1, data2)

  write_csv(combined_data, output_file_path)


}, error = function(e) {
  print(paste("An error occurred:", e$message))
})

library(readr)
library(dplyr)

combined_prolific_path <- "/Users/sawakeishiro/Desktop/code/reconstruction/data/Experiment2/prolific_demographic_combined.csv"
output_dir <- "/Users/sawakeishiro/Desktop/code/reconstruction/data/Experiment2/"
merged_output_path <- file.path(output_dir, "merged_prolific_and_raw_data.csv")

tryCatch({
  prolific_data <- read_csv(combined_prolific_path)
  if (!exists("raw_df")) {
    stop("The dataframe 'raw_df' was not found. Please load it before running this merge.")
  }

  merged_data <- inner_join(
    prolific_data, 
    raw_df, 
    by = c("Participant id" = "subj_id")
  )
  
  write_csv(merged_data, merged_output_path)
  
  print(paste("Successfully merged data and saved to:", merged_output_path))
  print(paste("Original prolific rows:", nrow(prolific_data)))
  print(paste("Original raw_df rows:", nrow(raw_df)))
  print(paste("Total merged rows:", nrow(merged_data)))

}, error = function(e) {
  print(paste("An error occurred:", e$message))
})


```

```{r explicit rule discoverers, include = FALSE, cache=F}
### RULE DISCOVERY 

rule_list <- c(
  "546e8bc6fdf99b2bc7ebd029",
  "56d56e550b5634000bdc5703",
  "572bcdbb34b25a0010ddd977",
  "59dd90f6e75b450001a68dac",
  "5ca6717bb54fe60001e22441",
  "5cafc17ace1fa80017a20d6b",
  "5ecf5a78ba30501d2ca5e34b",
  "5ee0edcb962eba464f486b40",
  "5f035c7ce9d86254e87992e4",
  "5f08e3eda7b89d16266b0cdf",
  "6151556ef73ba765c7ff55f6",
  "629f6b8c65fcae219e245284",
  "62fccb9647eb07d62e9b6baa",
  "63af557b3d4f219c3226b7d6",
  "63f77a5dde1ea03305e119e0",
  "63f7ac13d942e94c18c92d4b",
  "64a9858be9b52e570ffb3c1e",
  "661149a91b7fad3a3d2caf0c",
  "6658e2cb395c02efada6a7a1",
  "66661b1242b581f199e49f4c",
  "6693a6ffc58de3bb9c5341de",
  "669eb709554e753adddffe01",
  "66cbc2627559d53c0e3e4ac4",
  "6730f2e20f5915ff4ad2c07d",
  "673e2923ed2dd1f79b63d19d",
  "6762d29bf8f9712a46c73b62",
  "67803b5508f3d995ef13696b",
  "67e069199f5b036960b0cc5f",
  "681b781064a5701a5ceee3bc"
)

two_parts_df <- two_parts_df %>%
  mutate(mentioned_rule = subj_id %in% rule_list)

phase_1_df <- phase_1_df %>%
  mutate(mentioned_rule = subj_id %in% rule_list)

phase_2_df <- phase_2_df %>%
  mutate(mentioned_rule = subj_id %in% rule_list)

stoneposition<- raw_df %>%
  group_by(stoneposition) %>%
  summarise(n=length(unique(subj_id)))
  




```

### phase 1 accuracy

```{r phase_1_and_2_accuracy,  cache=F, fig.width=8}
# --- Define new bins for 50 trials ---
# Bins are now 1-5, 6-10, 11-15, 16-20, 21-25, 26-30, 31-35, 36-40, 41-45, 46-50
# --- Load necessary libraries (if not already loaded) ---
# library(ggplot2)
# library(dplyr)

# --- Define Position Dodge ---
# This was 'pd' in your target code. It nudges overlapping elements apart.
pd <- position_dodge(width = 0.2)
bin_labels <- c("1–5", "6–10", "11–15", "16–20", "21–25", 
                "26–30", "31–35", "36–40", "41–45", "46–50")

# Breaks are set at 0.5, 5.5, 10.5, etc. to correctly group the integer trial numbers
breaks <- c(0.5, 5.5, 10.5, 15.5, 20.5, 25.5, 30.5, 35.5, 40.5, 45.5, 50.5)

# This sequence creates vertical lines between the bins (at 1.5, 2.5, etc.)
vline_x <- seq(1.5, length(bin_labels) - 0.5, by = 1)

# --- Create binned dataframe ---
# Uses raw_df and rule_list which are available from previous chunks
df_binned <- raw_df %>%
  filter(!is_catch_trial, phase %in% c("part1", "part2")) %>%
  mutate(
    mentioned_rule = subj_id %in% rule_list,
    bin = cut(trial_num, breaks = breaks, labels = bin_labels,
              include.lowest = TRUE, right = TRUE),
    bin = factor(bin, levels = bin_labels)
  )

# --- Summarise by subj, phase, bin, and mentioned_rule ---
by_subj_bias <- df_binned %>%
  group_by(subj_id, phase, bin, mentioned_rule) %>%
  summarise(bias = mean(chose_onlystones, na.rm = TRUE), .groups = "drop")

# --- Subgroup summary (mentioned vs not) ---
# THIS CREATES THE MISSING 'group_summary_sub' OBJECT
group_summary_sub <- by_subj_bias %>%
  group_by(phase, bin, mentioned_rule) %>%
  summarise(
    m  = mean(bias, na.rm = TRUE),
    se = sd(bias,  na.rm = TRUE) / sqrt(n()),
    .groups = "drop"
  )
# --- Rewritten Plot ---
ggplot(group_summary_sub, 
       aes(x = bin, y = m, 
           color = mentioned_rule,  # Swapped from 'phase'
           shape = phase,           # Swapped from 'mentioned_rule'
           linetype = phase,        # Swapped from 'mentioned_rule'
           group = interaction(mentioned_rule, phase))) +
  
  # vertical and horizontal lines (from your original)
  geom_vline(xintercept = vline_x, linetype = "dashed", linewidth = 0.3, color = "grey") +
  geom_hline(yintercept = 1/3, linewidth = 0.3) +
  
  # Geoms now inherit aesthetics and use position_dodge
  geom_line(linewidth = 0.7, position = pd) +
  geom_point(position = pd, size = 2) +
  geom_errorbar(aes(ymin = m - se, ymax = m + se), 
                position = pd, width = 0.2, linetype = 'solid') +
  
  # Updated scales to match new aesthetics
  
  scale_shape_discrete(name = "Part") +
  scale_linetype_discrete(name = "Part") +
  
  scale_color_discrete(
    name = "Mentioned position",
    labels = c("TRUE" = "Mentioned", "FALSE" = "NOT mentioned")
  ) +
 

  # Updated labs
  labs(x = "Trial Bin", y = "P(stones only)",
       title = "Binned bias by rule discovery (group mean ± SE)") +
  theme_minimal()



phase_1_acc <- phase_1_df %>%
  group_by(subj_id) %>%
  filter(!is_catch_trial) %>%
  summarise(acc=mean(chose_gems))

r_phase_1_acc <- phase_1_df %>%
  group_by(subj_id) %>%
  filter(!is_catch_trial, mentioned_rule) %>%
  summarise(acc=mean(chose_gems))

nr_phase_1_acc <- phase_1_df %>%
  group_by(subj_id) %>%
  filter(!is_catch_trial, !mentioned_rule) %>%
  summarise(acc=mean(chose_gems))

phase_1_acc_slopes <- phase_1_df %>%
  group_by(subj_id) %>%
  filter(!is_catch_trial) %>%
  summarise(r=cor(chose_gems,trial_num))

r_phase_1_acc_slopes <- phase_1_df %>%
  group_by(subj_id) %>%
  filter(!is_catch_trial, mentioned_rule) %>%
  summarise(r=cor(chose_gems,trial_num))

nr_phase_1_acc_slopes <- phase_1_df %>%
  group_by(subj_id) %>%
  filter(!is_catch_trial, !mentioned_rule) %>%
  summarise(r=cor(chose_gems,trial_num))

phase_2_acc <- phase_2_df %>%
  group_by(subj_id) %>%
  filter(!is_catch_trial) %>%
  summarise(acc=mean(is_consistent, na.rm =TRUE))

r_phase_2_acc <- phase_2_df %>%
  group_by(subj_id) %>%
  filter(!is_catch_trial, mentioned_rule) %>%
  summarise(acc=mean(is_consistent, na.rm =TRUE))

nr_phase_2_acc <- phase_2_df %>%
  group_by(subj_id) %>%
  filter(!is_catch_trial, !mentioned_rule) %>%
  summarise(acc=mean(is_consistent, na.rm =TRUE))

phase_2_acc_slopes <- phase_2_df %>%
  group_by(subj_id) %>%
  filter(!is_catch_trial) %>%
  summarise(r=cor(is_consistent,trial_num, use = "complete.obs"))

r_phase_2_acc_slopes <- phase_2_df %>%
  group_by(subj_id) %>%
  filter(!is_catch_trial, mentioned_rule) %>%
  summarise(r=cor(is_consistent,trial_num, use = "complete.obs"))

nr_phase_2_acc_slopes <- phase_2_df %>%
  group_by(subj_id) %>%
  filter(!is_catch_trial, !mentioned_rule) %>%
  summarise(r=cor(is_consistent,trial_num, use = "complete.obs"))

phase_2_pstones <- phase_2_df %>%
  group_by(subj_id) %>%
  filter(!is_catch_trial) %>%
  summarise(pstones=mean(chose_onlystones,na.rm = TRUE))

r_phase_2_pstones <- phase_2_df %>%
  group_by(subj_id) %>%
  filter(!is_catch_trial, mentioned_rule) %>%
  summarise(pstones=mean(chose_onlystones,na.rm = TRUE))

nr_phase_2_pstones <- phase_2_df %>%
  group_by(subj_id) %>%
  filter(!is_catch_trial, !mentioned_rule) %>%
  summarise(pstones=mean(chose_onlystones,na.rm = TRUE))

```


In part 1, mean proportion correct (in non-catch trials) was `r phase_1_acc %>% pull(acc) %>%mean()%>%printnum()` (median=`r phase_1_acc %>% pull(acc) %>%median()%>%printnum()`) and was significantly higher than chance (1/3; `r apa_print(t.test(phase_1_acc %>% pull(acc),mu=1/3))$statistic`). Participants significantly improved as a function of trial number (a t-test on subject-specific learning slopes against 0: `r apa_print(t.test(phase_1_acc_slopes %>% pull(r)))$statistic`). 

The mean accuracy for participants who mentioned the rule was `r r_phase_1_acc %>% pull(acc) %>%mean()%>%printnum()` (median=`r r_phase_1_acc %>% pull(acc) %>%median()%>%printnum()`) and was significantly higher than chance (1/3; `r apa_print(t.test(r_phase_1_acc %>% pull(acc),mu=1/3))$statistic`). Participants significantly improved as a function of trial number (a t-test on subject-specific learning slopes against 0: `r apa_print(t.test(r_phase_1_acc_slopes %>% pull(r)))$statistic`). 

The mean accuracy for participants who did not mention the rule was r `r nr_phase_1_acc %>% pull(acc) %>%mean()%>%printnum()` (median=`r nr_phase_1_acc %>% pull(acc) %>%median()%>%printnum()`) and was not significantly higher than chance (1/3; `r apa_print(t.test(nr_phase_1_acc %>% pull(acc),mu=1/3))$statistic`). Participants did not improve as a function of trial number (a t-test on subject-specific learning slopes against 0: `r apa_print(t.test(nr_phase_1_acc_slopes %>% pull(r)))$statistic`). 



### Hypothesis 1 (RULE LEARNING AVOIDANCE)

```{r phase1avoidance, include = FALSE, cache=F}
phase_1_pstones <- phase_1_df %>%
  group_by(subj_id) %>%
  filter(!is_catch_trial) %>%
  summarise(pstones=mean(chose_onlystones,na.rm = TRUE))

r_phase_1_pstones <- phase_1_df %>%
  group_by(subj_id) %>%
  filter(!is_catch_trial, mentioned_rule) %>%
  summarise(pstones=mean(chose_onlystones,na.rm = TRUE))

nr_phase_1_pstones <- phase_1_df %>%
  group_by(subj_id) %>%
  filter(!is_catch_trial, !mentioned_rule) %>%
  summarise(pstones=mean(chose_onlystones,na.rm = TRUE))


```

In part 1, participants chose the side that always had stones (in non-catch trials)  on `r phase_1_pstones %>% pull(pstones) %>%mean()%>%printnum()` (median=`r phase_1_pstones %>% pull(pstones) %>%median()%>%printnum()`) of all trials, significantly below chance (1/3; `r apa_print(t.test(phase_1_pstones %>% pull(pstones),mu=1/3))$statistic`).

Participants who mentioned the rule chose the side that always had stones on `r r_phase_1_pstones %>% pull(pstones) %>%mean()%>%printnum()` (median=`r r_phase_1_pstones %>% pull(pstones) %>%median()%>%printnum()`) of all trials, significantly below chance (1/3; `r apa_print(t.test(r_phase_1_pstones %>% pull(pstones),mu=1/3))$statistic`). This is in contrast to participants who did not mention the rule who chose the side that always had stones on `r nr_phase_1_pstones %>% pull(pstones) %>%mean()%>%printnum()` (median=`r nr_phase_1_pstones %>% pull(pstones) %>%median()%>%printnum()`) of all trials, significantly below chance (1/3; `r apa_print(t.test(nr_phase_1_pstones %>% pull(pstones),mu=1/3))$statistic`). The difference in means between the two groups are significant (`r apa_print(t.test(r_phase_1_pstones %>% pull(pstones),nr_phase_1_pstones %>% pull(pstones)))$statistic`)

### Hypothesis 2 (RULE LEARNING: SLOPES)

```{r phase1slopes, include = FALSE, cache=F}
library(broom)
pstones_slopes <- two_parts_df %>%
  group_by(subj_id) %>%
  filter(!is_catch_trial_part1) %>% 
  do(tidy(glm(
    chose_onlystones_part1 ~ trial_num, 
    family = binomial, 
    data = .
  ))) %>%
  filter(term == "trial_num") %>%
  select(subj_id, slope = estimate)

r_pstones_slopes <- two_parts_df %>%
  group_by(subj_id) %>%
  filter(!is_catch_trial_part1, mentioned_rule) %>% 
  do(tidy(glm(
    chose_onlystones_part1 ~ trial_num, 
    family = binomial, 
    data = .
  ))) %>%
  filter(term == "trial_num") %>%
  select(subj_id, slope = estimate)

nr_pstones_slopes <- two_parts_df %>%
  group_by(subj_id) %>%
  filter(!is_catch_trial_part1, !mentioned_rule) %>% 
  do(tidy(glm(
    chose_onlystones_part1 ~ trial_num, 
    family = binomial, 
    data = .
  ))) %>%
  filter(term == "trial_num") %>%
  select(subj_id, slope = estimate)


pstones_slopes_p2 <- phase_2_df %>%
  group_by(subj_id) %>%
  filter(!is_catch_trial) %>%
  do(tidy(glm(
    chose_onlystones ~ trial_num,
    family = binomial,
    data = .
  )))%>%
  filter(term == "trial_num") %>%
  select(subj_id, slope=estimate)

r_pstones_slopes_p2 <- phase_2_df %>%
  group_by(subj_id) %>%
  filter(!is_catch_trial, mentioned_rule) %>%
  do(tidy(glm(
    chose_onlystones ~ trial_num,
    family = binomial,
    data = .
  )))%>%
  filter(term == "trial_num") %>%
  select(subj_id, slope=estimate)

nr_pstones_slopes_p2 <- phase_2_df %>%
  group_by(subj_id) %>%
  filter(!is_catch_trial, !mentioned_rule) %>%
  do(tidy(glm(
    chose_onlystones ~ trial_num,
    family = binomial,
    data = .
  )))%>%
  filter(term == "trial_num") %>%
  select(subj_id, slope=estimate)


```

Participants improved as a function of trial number (a t-test on subject-specific learning slopes from a logistic regression against 0: `r apa_print(t.test(pstones_slopes %>% pull(slope)))$statistic`).This was true for participants who explicitly learnt the rule (`r apa_print(t.test(r_pstones_slopes %>% pull(slope)))$statistic`) and true for participants who did not learn the rule (`r apa_print(t.test(nr_pstones_slopes %>% pull(slope)))$statistic`). 

In part 2, participants did not avoid stones only as a function of trial number (a t-test on subject-specific learning slopes from a logistic regression against 0: `r apa_print(t.test(pstones_slopes_p2 %>% pull(slope)))$statistic`).This was true for participants who explicitly learnt the rule (`r apa_print(t.test(r_pstones_slopes_p2 %>% pull(slope)))$statistic`) but not true for participants who did not learn the rule (`r apa_print(t.test(nr_pstones_slopes_p2 %>% pull(slope)))$statistic`). 

### Hypothesis 3 (RECONSTRUCTION ACCURACY)

In part 2, mean proportion correct (in non-catch trials) was `r phase_2_acc %>% pull(acc) %>%mean()%>%printnum()` (median=`r phase_2_acc %>% pull(acc) %>%median()%>%printnum()`) and significantly higher than chance (1/3; `r apa_print(t.test(phase_2_acc %>% pull(acc),mu=1/3))$statistic`). 

Mean accuracy for participants who mentioned rule was `r r_phase_2_acc %>% pull(acc) %>%mean()%>%printnum()` (median=`r r_phase_2_acc %>% pull(acc) %>%median()%>%printnum()`) and significantly higher than chance (1/3; `r apa_print(t.test(r_phase_2_acc %>% pull(acc),mu=1/3))$statistic`). 

Mean accuracy for participants who did not mention the rule was `r nr_phase_2_acc %>% pull(acc) %>%mean()%>%printnum()` (median=`r nr_phase_2_acc %>% pull(acc) %>%median()%>%printnum()`) and significantly higher than chance (1/3; `r apa_print(t.test(nr_phase_2_acc %>% pull(acc),mu=1/3))$statistic`). 

The difference in mean between the two groups are not significant (`r apa_print(t.test(r_phase_2_acc %>% pull(acc),nr_phase_2_acc %>% pull(acc)))$statistic`). 

The probability that participants chose the only stones option in part 2 was `r phase_2_pstones %>% pull(pstones) %>% mean() %>% printnum()` (median = `r phase_2_pstones %>% pull(pstones) %>% median() %>% printnum()`). Mean probability for participants  who mentioned the rule was `r r_phase_2_pstones %>% pull(pstones) %>% mean() %>% printnum()` (median = `r r_phase_2_pstones %>% pull(pstones) %>% median() %>% printnum()`) and for participants who did not mention the rule was `r nr_phase_2_pstones %>% pull(pstones) %>% mean() %>% printnum()` (median = `r nr_phase_2_pstones %>% pull(pstones) %>% median() %>% printnum`)

### Hytpoehsis 4 (RECONSTRUCTION ACCURACY BY GUESS ACCURACY): 

```{r hypothesis_4, include = FALSE, cache=F}

part1_choices <- phase_1_df %>%
  select(subj_id, trial_num, part1_chose_gems = chose_gems)

paired_data <- phase_2_df %>%
  left_join(part1_choices, by = c("subj_id", "trial_num")) %>%
  filter(!is_catch_trial, !is.na(part1_chose_gems)) %>%
  group_by(subj_id, part1_chose_gems) %>%
  summarise(mean_accuracy = mean(is_consistent, na.rm = TRUE), .groups = 'drop') %>%
  pivot_wider(names_from = part1_chose_gems, values_from = mean_accuracy)

r_paired_data <- phase_2_df %>%
  left_join(part1_choices, by = c("subj_id", "trial_num")) %>%
  filter(!is_catch_trial, !is.na(part1_chose_gems), mentioned_rule) %>%
  group_by(subj_id, part1_chose_gems) %>%
  summarise(mean_accuracy = mean(is_consistent, na.rm = TRUE), .groups = 'drop') %>%
  pivot_wider(names_from = part1_chose_gems, values_from = mean_accuracy) %>%
  mutate(
    "diff" = `TRUE` - `FALSE`
  )

nr_paired_data <- phase_2_df %>%
  left_join(part1_choices, by = c("subj_id", "trial_num")) %>%
  filter(!is_catch_trial, !is.na(part1_chose_gems), !mentioned_rule) %>%
  group_by(subj_id, part1_chose_gems) %>%
  summarise(mean_accuracy = mean(is_consistent, na.rm = TRUE), .groups = 'drop') %>%
  pivot_wider(names_from = part1_chose_gems, values_from = mean_accuracy) %>%
  mutate(
    "diff" = `TRUE` - `FALSE`
  )

hyp4_test <- t.test(x = paired_data$`TRUE`, 
                    y = paired_data$`FALSE`, 
                    paired = TRUE,
                    alternative = "two.sided")


hyp4_d <- cohen.d(paired_data$`TRUE`, 
                  paired_data$`FALSE`, 
                  paired = TRUE)

r_hyp4_test <- t.test(x = r_paired_data$`TRUE`, 
                    y = r_paired_data$`FALSE`, 
                    paired = TRUE,
                    alternative = "two.sided")


r_hyp4_d <- cohen.d(r_paired_data$`TRUE`, 
                  r_paired_data$`FALSE`, 
                  paired = TRUE)

nr_hyp4_test <- t.test(x = nr_paired_data$`TRUE`, 
                    y = nr_paired_data$`FALSE`, 
                    paired = TRUE,
                    alternative = "two.sided")


nr_hyp4_d <- cohen.d(nr_paired_data$`TRUE`, 
                  nr_paired_data$`FALSE`, 
                  paired = TRUE)

mean_acc_true <- mean(paired_data[["TRUE"]], na.rm = TRUE)
r_mean_acc_true <- mean(r_paired_data[["TRUE"]], na.rm = TRUE)
nr_mean_acc_true <- mean(nr_paired_data[["TRUE"]], na.rm = TRUE)

mean_acc_false <- mean(paired_data[["FALSE"]], na.rm = TRUE)
r_mean_acc_false <- mean(r_paired_data[["FALSE"]], na.rm = TRUE)
nr_mean_acc_false <- mean(nr_paired_data[["FALSE"]], na.rm = TRUE)


```

Participants' ability to reconstruct their past decisions was significantly **higher** when they were correct in part 1 (mean accuracy = `r printnum(mean_acc_true)`) compared to incorrect (mean accuracy = `r printnum(mean_acc_false)`; `r apa_print(hyp4_test)$statistic`). 

For participants who mentioned the rule, their ability to reconstruct their past decisions was significantly higher when they were correct in part 1 (mean accuracy = `r printnum(r_mean_acc_true)`) compared to incorrect (mean accuracy = `r printnum(r_mean_acc_false)`; `r apa_print(r_hyp4_test)$statistic`). 

For participants who did not mention the rule, their ability to reconstruct their past decisions was higher when they were correct in part 1 (mean accuracy = `r printnum(nr_mean_acc_true)`) compared to incorrect (mean accuracy = `r printnum(nr_mean_acc_false)`. However, this result was not significant: `r apa_print(nr_hyp4_test)$statistic`). 

```{r accuracy,  cache=F, fig.width=8}



part1_choices <- phase_1_df %>%
  select(subj_id, trial_num, part1_chose_gems = chose_gems)

phase_2_merged_df <- phase_2_df %>%
  left_join(part1_choices, by = c("subj_id", "trial_num"))

data_for_test <- phase_2_merged_df %>%
  filter(!is_catch_trial, !is.na(part1_chose_gems))

long_paired_data_all <- data_for_test %>%
  group_by(subj_id, part1_chose_gems) %>%
  summarise(mean_accuracy = mean(is_consistent, na.rm = TRUE), .groups = "drop") %>%
  mutate(
    phase_1_choice = factor(part1_chose_gems, labels = c("Did Not Choose Gems", "Chose Gems")),
    group = "all"
  )

r_long_paired_data <- data_for_test %>%
  filter(mentioned_rule) %>%
  group_by(subj_id, part1_chose_gems) %>%
  summarise(mean_accuracy = mean(is_consistent, na.rm = TRUE), .groups = "drop") %>%
  mutate(
    phase_1_choice = factor(part1_chose_gems, labels = c("Did Not Choose Gems", "Chose Gems")),
    group = "r_mentioned"
  )

nr_long_paired_data <- data_for_test %>%
  filter(!mentioned_rule) %>%
  group_by(subj_id, part1_chose_gems) %>%
  summarise(mean_accuracy = mean(is_consistent, na.rm = TRUE), .groups = "drop") %>%
  mutate(
    phase_1_choice = factor(part1_chose_gems, labels = c("Did Not Choose Gems", "Chose Gems")),
    group = "nr_not_mentioned"
  )

plot_data_summary_all <- long_paired_data_all %>%
  group_by(group, phase_1_choice) %>%
  summarise(
    grand_mean_accuracy = mean(mean_accuracy, na.rm = TRUE),
    n = n(),
    sem = sd(mean_accuracy, na.rm = TRUE) / sqrt(n),
    .groups = "drop"
  )

r_plot_data_summary <- r_long_paired_data %>%
  group_by(group, phase_1_choice) %>%
  summarise(
    grand_mean_accuracy = mean(mean_accuracy, na.rm = TRUE),
    n = n(),
    sem = sd(mean_accuracy, na.rm = TRUE) / sqrt(n),
    .groups = "drop"
  )

nr_plot_data_summary <- nr_long_paired_data %>%
  group_by(group, phase_1_choice) %>%
  summarise(
    grand_mean_accuracy = mean(mean_accuracy, na.rm = TRUE),
    n = n(),
    sem = sd(mean_accuracy, na.rm = TRUE) / sqrt(n),
    .groups = "drop"
  )

plot_data_summary <- bind_rows(plot_data_summary_all,
                               r_plot_data_summary,
                               nr_plot_data_summary) %>%
  mutate(group = factor(group, levels = c("all", "r_mentioned", "nr_not_mentioned")))

long_subject_all <- long_paired_data_all
long_subject_r   <- r_long_paired_data
long_subject_nr  <- nr_long_paired_data

y_limits <- c(0, 1)

build_group_plot <- function(summary_df, subject_df, plot_title, bar_color) {
  ggplot(summary_df, aes(x = phase_1_choice, y = grand_mean_accuracy)) +
    geom_line(data = subject_df,
              aes(x = phase_1_choice, y = mean_accuracy, group = subj_id),
              color = "gray70", alpha = 0.4, inherit.aes = FALSE) +
    geom_point(data = subject_df,
               aes(x = phase_1_choice, y = mean_accuracy, group = subj_id),
               color = "gray70", alpha = 0.4, inherit.aes = FALSE) +
    # bars
    geom_bar(stat = "identity", width = 0.6, fill = bar_color) +
    # error bars
    geom_errorbar(aes(ymin = grand_mean_accuracy - sem, ymax = grand_mean_accuracy + sem),
                  width = 0.2, color = "black") +
    labs(title = plot_title,
         x = "Choice in Phase 1",
         y = "Mean Phase 2 Accuracy") +
    scale_y_continuous(limits = y_limits, expand = c(0, 0)) +
    theme_minimal() +
    theme(legend.position = "none",
          plot.title = element_text(size = 11, face = "bold"))
}

# --- create each plot ---
plot_all <- build_group_plot(plot_data_summary_all, long_subject_all,
                             "All participants", "#4C72B0")

plot_r <- build_group_plot(r_plot_data_summary, long_subject_r,
                           "Position mentioned", "#55A868")

plot_nr <- build_group_plot(nr_plot_data_summary, long_subject_nr,
                            "Position NOT mentioned", "#C44E52")

print(plot_all)
print(plot_r)
print(plot_nr)


plot_data_paired <- data_for_test %>%
  group_by(subj_id, mentioned_rule, part1_chose_gems) %>%
  summarise(recon_acc = mean(is_consistent, na.rm = TRUE), .groups = "drop")

group_summary_new <- plot_data_paired %>%
  group_by(mentioned_rule, part1_chose_gems) %>%
  summarise(
    m = mean(recon_acc, na.rm = TRUE),
    se = sd(recon_acc, na.rm = TRUE) / sqrt(n()),
    n = n(),
    .groups = "drop"
  ) %>%
  mutate(p1_label = ifelse(part1_chose_gems, "Part1: correct", "Part1: incorrect"),
         p1_label = factor(p1_label, levels = c("Part1: incorrect","Part1: correct")))


plot_pairs_new <- plot_data_paired %>%
  mutate(p1_label = ifelse(part1_chose_gems, "Part1: correct", "Part1: incorrect")) %>%
  mutate(p1_label = factor(p1_label, levels = c("Part1: incorrect","Part1: correct")))

# --- Load scales library for plot labels (if not already loaded via tidyverse) ---
library(scales)

p <- ggplot() +
  # lines connecting each participant's paired points
  geom_line(data = plot_pairs_new, aes(x = p1_label, y = recon_acc, group = subj_id),
            color = "grey70", alpha = 0.35, size = 0.4) +
  # individual subject points
  geom_point(data = plot_pairs_new, aes(x = p1_label, y = recon_acc),
             color = "grey40", alpha = 0.6, position = position_dodge(width = 0.4), size = 1.7) +
  # group mean points
  geom_point(data = group_summary_new, aes(x = p1_label, y = m, color = mentioned_rule),
             position = position_dodge(width = 0.4), size = 3) +
  # error bars for mean ± SE
  geom_errorbar(data = group_summary_new, aes(x = p1_label, ymin = m - se, ymax = m + se, color = mentioned_rule),
                width = 0.12, position = position_dodge(width = 0.4), size = 0.7) +
  facet_wrap(~ mentioned_rule, nrow = 1,
             labeller = as_labeller(c("FALSE" = "Did NOT mention position", "TRUE" = "Mentioned position"))) +
  scale_y_continuous(limits = c(0.25,1), breaks = seq(0,1,0.25), labels = percent_format(accuracy = 1)) +
  labs(x = NULL,
       y = "Reconstruction accuracy (Part 2)",
       title = "Reconstruction accuracy in Part 2 by Part 1 guess accuracy",
       color = "Mentioned rule") +
  theme_minimal(base_size = 12) +
  theme(legend.position = "none",
        strip.text = element_text(face = "bold"),
        plot.title = element_text(size = 12, face = "bold"),
        plot.subtitle = element_text(size = 10))

print(p)



```


### Hypothesis 5 (REMEMBERED ACCURACY VERSUS TRUE ACCURACY)

```{r hypothesis_5, include = FALSE, cache=F}

# all participants
accuracy_df <- phase_1_df %>%
  filter(!is_catch_trial) %>%
  group_by(subj_id) %>%
  summarise(actual_accuracy = mean(chose_gems), .groups = "drop") %>%
  left_join(
    phase_2_df %>%
      filter(!is_catch_trial) %>%
      group_by(subj_id) %>%
      summarise(remembered_accuracy = mean(chose_gems), .groups = "drop"),
    by = "subj_id"
  )


# r_mentioned
r_accuracy_df <- phase_1_df %>%
  filter(!is_catch_trial, mentioned_rule) %>%
  group_by(subj_id) %>%
  summarise(actual_accuracy = mean(chose_gems), .groups = "drop") %>%
  left_join(
    phase_2_df %>%
      filter(!is_catch_trial, mentioned_rule) %>%
      group_by(subj_id) %>%
      summarise(remembered_accuracy = mean(chose_gems), .groups = "drop"),
    by = "subj_id"
  )



# nr_not_mentioned
nr_accuracy_df <- phase_1_df %>%
  filter(!is_catch_trial, !mentioned_rule) %>%
  group_by(subj_id) %>%
  summarise(actual_accuracy = mean(chose_gems), .groups = "drop") %>%
  left_join(
    phase_2_df %>%
      filter(!is_catch_trial, !mentioned_rule) %>%
      group_by(subj_id) %>%
      summarise(remembered_accuracy = mean(chose_gems), .groups = "drop"),
    by = "subj_id"
  )




```

Participants remembered their guesses as aligned with the location of the gems on `r accuracy_df$remembered_accuracy%>%mean()%>%printnum()`, a frequency significantly higher than their actual accuracy in part 1, which was `r accuracy_df$actual_accuracy%>%mean()%>%printnum()` (`r apa_print(t.test(accuracy_df$remembered_accuracy, accuracy_df$actual_accuracy, paired = TRUE))$statistic`).

Participants who mentioned the rule remembered their guesses as aligned with the location of the gems on `r r_accuracy_df$remembered_accuracy%>%mean()%>%printnum()`, a frequency not significantly higher than their actual accuracy in part 1, which was `r r_accuracy_df$actual_accuracy%>%mean()%>%printnum()` (`r apa_print(t.test(r_accuracy_df$remembered_accuracy, r_accuracy_df$actual_accuracy, paired = TRUE))$statistic`).

Participants remembered their guesses as aligned with the location of the gems on `r nr_accuracy_df$remembered_accuracy%>%mean()%>%printnum()`, a frequency not significantly higher than their actual accuracy in part 1, which was `r nr_accuracy_df$actual_accuracy%>%mean()%>%printnum()` (`r apa_print(t.test(nr_accuracy_df$remembered_accuracy, nr_accuracy_df$actual_accuracy, paired = TRUE))$statistic`).


### Hypothesis 6 (POSITION BIAS IN RECONSTRUCTION ERRORS

We will test the null hypothesis that participants’remembered decisions are not biased for or against the stones-only location. For this, we will focus on trials in which the original decision revealed gems, but they made an error in reconstruction. We will compare the probability of choosing the stones-only position against 50%. Since different participants will have different number of trials for this analysis, this will be tested using a mixed-effects logistic regression (chose_stones_only∼1+(1|subj_id), run with lme4’s glmer, assuming a binomial variable and using a bobyqa optimizer).

```{r hypothesis_6, include = TRUE, cache=F}

part1_choices <- phase_1_df %>%
  mutate(
    part1_choice = case_when(
      chose_gems ~ "gem",
      chose_onlystones ~ "stone",
      chose_other ~ "other"
    ),
    stone_number = case_when(
      stoneposition == 'right' ~ right_number,
      stoneposition == "middle" ~ middle_number,
      stoneposition == "left" ~ left_number
    ),
    other_number = case_when(
      correct_choice_position == "right" & stoneposition == "middle" ~ left_number,
      correct_choice_position == "right" & stoneposition == "left" ~ middle_number,
      correct_choice_position == "middle" & stoneposition == "right" ~ left_number,
      correct_choice_position == "middle" & stoneposition == "left" ~ right_number, 
      correct_choice_position == "left" & stoneposition == "right" ~ middle_number,
      correct_choice_position == "left" & stoneposition == "middle" ~ right_number
    ),
    stone_higher_than_other = case_when(
      stone_number > other_number ~ TRUE,
      stone_number <= other_number ~ FALSE,
    )
  ) %>%
  select(subj_id, trial_num, part1_choice, is_catch_trial, mentioned_rule, stone_number, other_number, stone_higher_than_other)

part2_choices <- phase_2_df %>%
  mutate(
    part2_choice = case_when(
      chose_gems ~ "gem",
      chose_onlystones ~ "stone",
      chose_other ~ "other"
    )
  ) %>%
  select(subj_id, trial_num, part2_choice, is_consistent, is_catch_trial, mentioned_rule)

choices_merged <- part2_choices %>%
  left_join(part1_choices, by = c("subj_id", "trial_num", "is_catch_trial", "mentioned_rule")) %>%
  mutate(
    chose_onlystones_part2 = (part2_choice == 'stone')
  )

never_stone_ids <- choices_merged %>%
  group_by(subj_id) %>%
  summarise(ever_stone = any(part2_choice == "stone", na.rm = TRUE)) %>%
  filter(!ever_stone) %>%
  pull(subj_id)

choices_merged <- choices_merged %>%
  filter(!subj_id %in% never_stone_ids)

is_not_consistent_chosegems_part1 <- choices_merged %>%
  filter(!is_catch_trial,
         part1_choice == 'gem',
         !is_consistent)

r_is_not_consistent_chosegems_part1 <- choices_merged %>%
  filter(!is_catch_trial,
         part1_choice == 'gem',
         !is_consistent,
         mentioned_rule)

nr_is_not_consistent_chosegems_part1 <- choices_merged %>%
  filter(!is_catch_trial,
         part1_choice == 'gem',
         !is_consistent,
         !mentioned_rule)



library(lme4)


h6_model <- glmer(chose_onlystones_part2  ~ 1 + (1 | subj_id),
                  data = is_not_consistent_chosegems_part1,
                  family = binomial,
                  control = glmerControl(optimizer = "bobyqa"))

r_h6_model <- glmer(chose_onlystones_part2 ~ 1 + (1 | subj_id),
                  data = r_is_not_consistent_chosegems_part1,
                  family = binomial,
                  control = glmerControl(optimizer = "bobyqa"))

nr_h6_model <- glmer(chose_onlystones_part2 ~ 1 + (1 | subj_id),
                  data = nr_is_not_consistent_chosegems_part1,
                  family = binomial,
                  control = glmerControl(optimizer = "bobyqa"))

h6_model_trial <-glmer(chose_onlystones_part2 ~ trial_num + (1 + trial_num | subj_id),
                  data = is_not_consistent_chosegems_part1,
                  family = binomial,
                  control = glmerControl(optimizer = "bobyqa"))

r_h6_model_trial <- glmer(chose_onlystones_part2 ~ trial_num + (1 + trial_num | subj_id),
                  data = r_is_not_consistent_chosegems_part1,
                  family = binomial,
                  control = glmerControl(optimizer = "bobyqa"))

nr_h6_model_trial <- glmer(chose_onlystones_part2 ~ trial_num + (1 + trial_num | subj_id),
                  data = nr_is_not_consistent_chosegems_part1,
                  family = binomial,
                  control = glmerControl(optimizer = "bobyqa"))

h6_model_mentioned_vs_notmentioned <- glmer(chose_onlystones_part2 ~ mentioned_rule + (1 | subj_id),
                  data = is_not_consistent_chosegems_part1,
                  family = binomial,
                  control = glmerControl(optimizer = "bobyqa"))


```

In Part 2 trials in which the original decision revealed gems but participants made an error in reconstruction, we fitted a mixed-effects logistic regression with a random intercept for participant. The estimated probability of selecting the stones-only position was 
`r printnum(round(plogis(fixef(h6_model)[1]), 2))`, 
corresponding to a log-odds estimate of 
β = `r printnum(round(fixef(h6_model)[1], 2))`, 
SE = `r printnum(round(summary(h6_model)$coefficients[1, "Std. Error"], 2))`, 
*z* = `r printnum(round(summary(h6_model)$coefficients[1, "z value"], 2))`, 
*p* `r printp(summary(h6_model)$coefficients[1, "Pr(>|z|)"])`, 



For participants who mentioned the rule the estimated probability of selecting the stones-only position was `r printnum(round(unname(1 / (1 + exp(-fixef(r_h6_model)))), 2))` (β = `r printnum(round(fixef(r_h6_model)[1], 2))`, 
SE = `r printnum(round(summary(r_h6_model)$coefficients[1, "Std. Error"], 2))`, 
*z* = `r printnum(round(summary(r_h6_model)$coefficients[1, "z value"], 2))`, 
*p* `r printp(summary(r_h6_model)$coefficients[1, "Pr(>|z|)"])`, ).

For participants who did not mention the rule, the estimated probability of selecting the stones-only position was `r printnum(round(unname(1 / (1 + exp(-fixef(nr_h6_model)))), 2))` (β = `r printnum(round(fixef(nr_h6_model)[1], 2))`, 
SE = `r printnum(round(summary(nr_h6_model)$coefficients[1, "Std. Error"], 2))`, 
*z* = `r printnum(round(summary(nr_h6_model)$coefficients[1, "z value"], 2))`, 
*p* `r printp(summary(nr_h6_model)$coefficients[1, "Pr(>|z|)"])`, ).

The difference between the two groups (those who mentioned and did not mention the rule) is significant, with *mentioned_rule* included as a fixed effect in the model. Participants who mentioned the rule were less likely to choose the stones-only option than those who did not mention the rule (β =`r printnum(round(summary(h6_model_mentioned_vs_notmentioned)$coefficients["mentioned_ruleTRUE", "Estimate"],2))`, SE = `r printnum(round(summary(h6_model_mentioned_vs_notmentioned)$coefficients["mentioned_ruleTRUE", "Std. Error"],2))`, z = `r printnum(round(summary(h6_model_mentioned_vs_notmentioned)$coefficients["mentioned_ruleTRUE", "z value"],2))`, p `r printp(summary(h6_model_mentioned_vs_notmentioned)$coefficients["mentioned_ruleTRUE","Pr(>|z|)"][1])`)

```{r reconstruction plots,  cache=F, fig.width=8}


h6_viz_data <- is_not_consistent_chosegems_part1 %>%
  # Create trial bins
  mutate(
    trial_bin = cut(
      trial_num,
      breaks = c(0, 10, 20, 30, 40, 50),
      labels = c("1-10", "11-20", "21-30", "31-40", "41-50")
    )
  ) %>%
  
  # Group and summarize as before
  group_by(trial_bin, mentioned_rule, subj_id) %>%
  summarise(
    n = n(),
    prob_stones = mean(chose_onlystones_part2, na.rm = TRUE),
    se = sqrt(prob_stones * (1 - prob_stones) / n)
  ) %>%
  mutate(
    ci_lower = prob_stones - 1.96 * se,
    ci_upper = prob_stones + 1.96 * se
  ) %>%
  ungroup() %>%
  mutate(
    Group = ifelse(mentioned_rule, "Position mentioned", "Position NOT mentioned")
  ) 


# 2. Create the plot
ggplot(h6_viz_data, aes(x = trial_bin, y = prob_stones, group = Group, color = Group)) +
  
  # Add the 50% null hypothesis line
  geom_hline(
    yintercept = 0.5,
    linetype = "dashed",
    color = "red",
    linewidth = 1
  ) +
  
  # Add error bars
  geom_errorbar(
    aes(ymin = ci_lower, ymax = ci_upper),
    width = 0.2,
    alpha = 0.7
  ) +
  
  # Add lines and points
  geom_line(linewidth = 1.2, alpha = 0.7) +
  geom_point(size = 3) +
  
  # Clean up labels and appearance
  scale_y_continuous(
    labels = scales::percent_format(),
    limits = c(0, 1) # Set y-axis from 0% to 100%
  ) +
  labs(
    title = "Hypothesis 6: Bias in Reconstruction Errors (Exp. 2)",
    subtitle = "Trials where original choice revealed gems, but reconstruction was an error.",
    x = "Trial Bin",
    y = "Probability of Choosing Stones-Only Location",
    color = "Participant Group" # Legend title
  ) +
  theme_minimal() +
  theme(legend.position = "bottom",
        plot.title = element_text(face = "bold"))

library(dplyr)
library(ggplot2)

h6_subject_bin <- is_not_consistent_chosegems_part1 %>%
  mutate(
    trial_bin = cut(
      trial_num,
      breaks = c(0, 10, 20, 30, 40, 50),
      labels = c("1-10", "11-20", "21-30", "31-40", "41-50"),
      right = TRUE
    )
  ) %>%
  group_by(subj_id, trial_bin, mentioned_rule) %>%
  summarise(
    prob_stones = mean(chose_onlystones_part2, na.rm = TRUE),
    n_trials = n(),
    .groups = "drop"
  ) %>%
  mutate(
    Group = ifelse(mentioned_rule, "Position mentioned", "Position NOT mentioned"),
    trial_bin = factor(trial_bin, levels = c("1-10","11-20","21-30","31-40","41-50"))
  )

summary_df <- h6_subject_bin %>%
  group_by(trial_bin, Group) %>%
  summarise(
    mean_prob = mean(prob_stones, na.rm = TRUE),
    se = sd(prob_stones, na.rm = TRUE) / sqrt(n()),
    n_subj = n(),
    .groups = "drop"
  )


pd <- position_dodge(width = 0.75)

ggplot(h6_subject_bin, aes(x = trial_bin, y = prob_stones, fill = Group, group = Group)) +
  geom_jitter(aes(color = Group), 
              position = position_jitterdodge(jitter.width = 0.12, dodge.width = 0.75), 
              size = 1.6, alpha = 0.6, show.legend = FALSE) +
  geom_errorbar(data = summary_df,
                aes(x = trial_bin, y = mean_prob, ymin = mean_prob - se, ymax = mean_prob + se, group = Group),
                width = 0.18,
                position = pd,
                color = "black",
                linewidth = 0.8) +
  geom_line(data = summary_df,
            aes(x = trial_bin, y = mean_prob, color = Group, group = Group),
            position = pd,
            linewidth = 0.9,
            alpha = 0.8) +
  geom_point(data = summary_df,
             aes(x = trial_bin, y = mean_prob, group = Group),
             position = pd,
             shape = 23, size = 3, fill = "white", color = "black") +
  geom_hline(yintercept = 0.5, linetype = "dashed", color = "black", linewidth = 0.7) +
  labs(
    title = "Exp. 2: Subject-level probability of choosing stones by trial bin",
    x = "Trial bin",
    y = "Probability of choosing stones (per subject)",
    fill = "Group"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    legend.position = "right",
    panel.grid.minor = element_blank()
  )


h6_subject_bin_stonehigher <- is_not_consistent_chosegems_part1 %>%
  mutate(
    trial_bin = cut(
      trial_num,
      breaks = c(0, 10, 20, 30, 40, 50),
      labels = c("1-10", "11-20", "21-30", "31-40", "41-50"),
      right = TRUE
    )
  ) %>%
  group_by(subj_id, trial_bin, mentioned_rule) %>%
  filter(stone_higher_than_other) %>%
  summarise(
    prob_stones = mean(chose_onlystones_part2, na.rm = TRUE),
    n_trials = n(),
    .groups = "drop"
  ) %>%
  mutate(
    Group = ifelse(mentioned_rule, "Position mentioned", "Position NOT mentioned"),
    trial_bin = factor(trial_bin, levels = c("1-10","11-20","21-30","31-40","41-50"))
  )

summary_df_stonehigher <- h6_subject_bin_stonehigher %>%
  group_by(trial_bin, Group) %>%
  summarise(
    mean_prob = mean(prob_stones, na.rm = TRUE),
    se = sd(prob_stones, na.rm = TRUE) / sqrt(n()),
    n_subj = n(),
    .groups = "drop"
  )

ggplot(h6_subject_bin_stonehigher, aes(x = trial_bin, y = prob_stones, fill = Group, group = Group)) +
  geom_jitter(aes(color = Group), 
              position = position_jitterdodge(jitter.width = 0.12, dodge.width = 0.75), 
              size = 1.6, alpha = 0.6, show.legend = FALSE) +
  geom_errorbar(data = summary_df_stonehigher,
                aes(x = trial_bin, y = mean_prob, ymin = mean_prob - se, ymax = mean_prob + se, group = Group),
                width = 0.18,
                position = pd,
                color = "black",
                linewidth = 0.8) +
  geom_line(data = summary_df_stonehigher,
            aes(x = trial_bin, y = mean_prob, color = Group, group = Group),
            position = pd,
            linewidth = 0.9,
            alpha = 0.8) +
  geom_point(data = summary_df_stonehigher,
             aes(x = trial_bin, y = mean_prob, group = Group),
             position = pd,
             shape = 23, size = 3, fill = "white", color = "black") +
  geom_hline(yintercept = 0.5, linetype = "dashed", color = "black", linewidth = 0.7) +
  labs(
    title = "Exp. 2: Subject-level probability of choosing stones by trial bin (stone higher)",
    x = "Trial bin",
    y = "Probability of choosing stones (per subject)",
    fill = "Group"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    legend.position = "right",
    panel.grid.minor = element_blank()
  )

library(ggplot2)
library(dplyr)

h6_viz_data <- is_not_consistent_chosegems_part1 %>%
  
  
  # Now, group by the bins AND by the rule-mentioning group
  group_by(subj_id, mentioned_rule) %>%
  
  # Calculate the probability, count (for CI), and standard error
  summarise(
    n = n(),
    prob_stones = mean(chose_onlystones_part2, na.rm = TRUE),
    se = sqrt(prob_stones * (1 - prob_stones) / n) # Std. Error of a proportion
  ) %>%
  
  # Calculate 95% Confidence Intervals
  mutate(
    ci_lower = prob_stones - 1.96 * se,
    ci_upper = prob_stones + 1.96 * se
  ) %>%
  

  
  # Ungroup and also recode 'mentioned_rule' for a clearer plot legend
  ungroup() %>%
  mutate(
    Group = ifelse(mentioned_rule, "Position mentioned", "Position NOT mentioned")
  )

h6_alltrial <- is_not_consistent_chosegems_part1 %>%
  group_by(subj_id) %>%
  summarise(
    mean_prob_stones = mean(chose_onlystones_part2, na.rm = TRUE),
    mentioned_rule  = unique(mentioned_rule)
  ) 


ggplot(h6_alltrial, aes(x = mentioned_rule, y = mean_prob_stones, fill = mentioned_rule)) +
  geom_boxplot(alpha = 0.5, width = 0.6, outlier.shape = NA, color = "gray40") +
  
  # Add mean points
  stat_summary(fun = mean, geom = "point", 
               shape = 23, size = 4, fill = "white", color = "black") +
  
  # Add mean ± SE error bars
  stat_summary(fun.data = mean_se, geom = "errorbar", 
               width = 0.2, color = "black", linewidth = 0.8) +
  
  # Optionally overlay individual points for transparency
  geom_jitter(width = 0.1, alpha = 0.5, color = "gray50") +
  
  labs(
    title = "Probability of Choosing Stones by Rule Mention",
    x = "",
    y = "Probability of Choosing Stones",
    fill = "Group"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    legend.position = "none",
    panel.grid.minor = element_blank()
  )



ggplot(h6_viz_data, aes(x = n, y = prob_stones, color = Group)) +
  geom_point(size = 3, alpha = 0.7) +
  geom_smooth(method = "lm", se = FALSE, linetype = "dashed", linewidth = 0.7, aes(group = Group)) +
  theme_minimal() +
  labs(
    title = "Probability of Stones by Group",
    x = "Trial Number",
    y = "Probability of only stones"
  )



#scatter plopt, x-axis number of trials, the probability of choosing stones y-axis 
```

 
### Hypothesis 7 (CONFIDENCE: CORRECR VERSUS INCORRECT)

```{r hypothesis_7, include = FALSE, cache=F}

conf_mean_correct <- phase_2_df %>%
  group_by(subj_id) %>%
  filter(!is_catch_trial, is_consistent) %>%
  summarise(mean_conf = mean(confidence, na.rm = TRUE),
            median_conf = median(confidence, na.rm=TRUE))

r_conf_mean_correct <- phase_2_df %>%
  group_by(subj_id) %>%
  filter(!is_catch_trial, is_consistent, mentioned_rule) %>%
  summarise(mean_conf = mean(confidence, na.rm = TRUE),
            median_conf = median(confidence, na.rm=TRUE))

nr_conf_mean_correct <- phase_2_df %>%
  group_by(subj_id) %>%
  filter(!is_catch_trial, is_consistent, !mentioned_rule) %>%
  summarise(mean_conf = mean(confidence, na.rm = TRUE),
            median_conf = median(confidence, na.rm=TRUE))


conf_mean_incorrect <- phase_2_df %>%
  group_by(subj_id) %>%
  filter(!is_catch_trial, !is_consistent) %>%
  summarise(mean_conf = mean(confidence, na.rm=TRUE),
            median_conf = median(confidence, na.rm=TRUE))

r_conf_mean_incorrect <- phase_2_df %>%
  group_by(subj_id) %>%
  filter(!is_catch_trial, !is_consistent, mentioned_rule) %>%
  summarise(mean_conf = mean(confidence, na.rm=TRUE),
            median_conf = median(confidence, na.rm=TRUE))

nr_conf_mean_incorrect <- phase_2_df %>%
  group_by(subj_id) %>%
  filter(!is_catch_trial, !is_consistent,!mentioned_rule) %>%
  summarise(mean_conf = mean(confidence, na.rm=TRUE),
            median_conf = median(confidence, na.rm=TRUE))


```

In part 2, mean confidence for correct decisions was `r conf_mean_correct$mean_conf%>%mean()%>%printnum()` (median=`r conf_mean_correct$mean_conf%>%median()%>%printnum()`). The mean confidence for incorrect decisions was `r conf_mean_incorrect$mean_conf%>%mean()%>%printnum()` (median=`r conf_mean_incorrect$mean_conf%>%median()%>%printnum()`). The difference in confidence was significant `r apa_print(t.test(x=conf_mean_correct$mean_conf, y=conf_mean_incorrect$mean_conf, alternative="two.sided",paired=TRUE))$statistic`

For participants who mentioned the rule, mean confidence for correct decisions was `r r_conf_mean_correct$mean_conf%>%mean()%>%printnum()` (median=`r r_conf_mean_correct$mean_conf%>%median()%>%printnum()`). The mean confidence for incorrect decisions was `r r_conf_mean_incorrect$mean_conf%>%mean()%>%printnum()` (median=`r r_conf_mean_incorrect$mean_conf%>%median()%>%printnum()`). The difference in confidence was significant `r apa_print(t.test(x=r_conf_mean_correct$mean_conf, y=r_conf_mean_incorrect$mean_conf, alternative="two.sided",paired=TRUE))$statistic`

For participants who did not mention the rule, mean confidence for correct decisions was `r nr_conf_mean_correct$mean_conf%>%mean()%>%printnum()` (median=`r nr_conf_mean_correct$mean_conf%>%median()%>%printnum()`). The mean confidence for incorrect decisions was `r nr_conf_mean_incorrect$mean_conf%>%mean()%>%printnum()` (median=`r nr_conf_mean_incorrect$mean_conf%>%median()%>%printnum()`). The difference in confidence was significant `r apa_print(t.test(x=nr_conf_mean_correct$mean_conf, y=nr_conf_mean_incorrect$mean_conf, alternative="two.sided",paired=TRUE))$statistic`


### Hypothesis 8 (CONFIDENCE: GEMS VERSUS STONES)


```{r hypothesis_8, include = FALSE, cache=F}
gem_conf <- phase_2_df %>%
  group_by(subj_id) %>%
  filter(!is_catch_trial, !is_consistent, chose_gems) %>%
  summarise(mean_conf=mean(confidence),
            median = median(confidence))

r_gem_conf <- phase_2_df %>%
  group_by(subj_id) %>%
  filter(!is_catch_trial, !is_consistent, chose_gems, mentioned_rule) %>%
  summarise(mean_conf=mean(confidence),
            median = median(confidence))

nr_gem_conf <- phase_2_df %>%
  group_by(subj_id) %>%
  filter(!is_catch_trial, !is_consistent, chose_gems, !mentioned_rule) %>%
  summarise(mean_conf=mean(confidence),
            median = median(confidence))

stone_conf <- phase_2_df %>% 
  group_by(subj_id) %>%
  filter(!is_catch_trial, is_consistent, !chose_gems) %>%
  summarise(mean_conf=mean(confidence),
            median = median(confidence))

r_stone_conf <- phase_2_df %>% 
  group_by(subj_id) %>%
  filter(!is_catch_trial, !is_consistent, !chose_gems, mentioned_rule) %>%
  summarise(mean_conf=mean(confidence),
            median = median(confidence))

nr_stone_conf <- phase_2_df %>% 
  group_by(subj_id) %>%
  filter(!is_catch_trial, !is_consistent, !chose_gems, !mentioned_rule) %>%
  summarise(mean_conf=mean(confidence),
            median = median(confidence))

conf_df <- gem_conf %>%
  rename(mean_conf_gem = mean_conf, median_gem = median) %>%
  inner_join(stone_conf %>% rename(mean_conf_stone = mean_conf, median_stone = median),
             by = "subj_id")

r_conf_df <- r_gem_conf %>%
  rename(mean_conf_gem = mean_conf, median_gem = median) %>%
  inner_join(r_stone_conf %>% rename(mean_conf_stone = mean_conf, median_stone = median),
             by = "subj_id")

nr_conf_df <- nr_gem_conf %>%
  rename(mean_conf_gem = mean_conf, median_gem = median) %>%
  inner_join(nr_stone_conf %>% rename(mean_conf_stone = mean_conf, median_stone = median),
             by = "subj_id")


```

In part 2, mean confidence when touching gems on trials in which participants were correct was `r gem_conf$mean_conf %>%mean()%>%printnum()` (median=`r gem_conf$mean_conf%>%median()%>%printnum() `). In trials which they chose stones, it was `r stone_conf$mean_conf %>%mean()%>%printnum()` (median=`r  stone_conf$mean_conf%>%median()%>%printnum() `). This difference was significant (`r apa_print(t.test(conf_df$mean_conf_gem, conf_df$mean_conf_stone, paired=TRUE))$statistic`). 

For participants who mentioned the rule, mean confidence when touching gems on trials in which participants were correct was `r r_gem_conf$mean_conf %>%mean()%>%printnum()` (median=`r r_gem_conf$mean_conf%>%median()%>%printnum() `). In trials which they chose stones, it was `r r_stone_conf$mean_conf %>%mean()%>%printnum()` (median=`r  r_stone_conf$mean_conf%>%median()%>%printnum() `). This difference was significant (`r apa_print(t.test(r_conf_df$mean_conf_gem, r_conf_df$mean_conf_stone, paired=TRUE))$statistic`). 

For participants who did not mention the rule, , mean confidence when touching gems on trials in which participants were correct was `r nr_gem_conf$mean_conf %>%mean()%>%printnum()` (median=`r nr_gem_conf$mean_conf%>%median()%>%printnum() `). In trials which they chose stones, it was `r nr_stone_conf$mean_conf %>%mean()%>%printnum()` (median=`r  nr_stone_conf$mean_conf%>%median()%>%printnum() `). This difference was significant (`r apa_print(t.test(nr_conf_df$mean_conf_gem, nr_conf_df$mean_conf_stone, paired=TRUE))$statistic`). 

```{r hypothesis8_plot, include = TRUE, cache=F}
library(dplyr)
library(ggplot2)

compute_plot_data <- function(df) {
  df %>%
    filter(!is_consistent, !is_catch_trial) %>%   # <--- use !is_consistent
    mutate(
      choice_type = case_when(
        chose_gems       ~ "Chose Gems",
        chose_other      ~ "Chose Other",
        chose_onlystones ~ "Chose Only Stones",
        TRUE             ~ NA_character_
      )
    ) %>%
    filter(!is.na(choice_type)) %>%
    group_by(subj_id, choice_type) %>%
    summarise(mean_conf_per_subj = mean(confidence, na.rm = TRUE), .groups = "drop") %>%
    group_by(choice_type) %>%
    summarise(
      mean_confidence = mean(mean_conf_per_subj, na.rm = TRUE),
      n = n(),
      sem = sd(mean_conf_per_subj, na.rm = TRUE) / sqrt(n),
      .groups = "drop"
    )
}


plot_data_all <- compute_plot_data(phase_2_df)
plot_data_r   <- compute_plot_data(filter(phase_2_df, mentioned_rule))
plot_data_nr  <- compute_plot_data(filter(phase_2_df, !mentioned_rule))

all_max <- max(
  c(plot_data_all$mean_confidence + plot_data_all$sem,
    plot_data_r$mean_confidence + plot_data_r$sem,
    plot_data_nr$mean_confidence + plot_data_nr$sem),
  na.rm = TRUE
)
y_limits <- c(0, ifelse(is.finite(all_max), all_max * 1.05, 1))

choice_colors <- c(
  "Chose Gems" = "forestgreen",
  "Chose Other" = "royalblue",
  "Chose Only Stones" = "firebrick"
)

build_conf_plot <- function(summary_df, plot_title) {
  ggplot(summary_df, aes(x = choice_type, y = mean_confidence, fill = choice_type)) +
    geom_bar(stat = "identity", show.legend = FALSE, alpha=0.8) +
    geom_errorbar(aes(ymin = mean_confidence - sem, ymax = mean_confidence + sem), width = 0.2) +
    scale_fill_manual(values = choice_colors) +
    labs(
      title = plot_title,
      subtitle = "Confidence scores from Part 2 (consistent trials only)",
      x = "Participant's Choice",
      y = "Mean Confidence"
    ) +
    scale_y_continuous(limits = y_limits, expand = c(0, 0)) +
    theme_minimal() +
    theme(plot.title = element_text(size = 12, face = "bold"))
}

plot_all <- build_conf_plot(plot_data_all, "All Participants")
plot_r   <- build_conf_plot(plot_data_r, "Participants Who Mentioned the Rule")
plot_nr  <- build_conf_plot(plot_data_nr, "Participants Who Did NOT Mention the Rule")

print(plot_all)
print(plot_r)
print(plot_nr)


```


```{r conf_significance, fig.width=8, cache=F}
library(dplyr)
library(tidyr)
library(broom)
library(ggpubr)
library(ggplot2)


compute_subject_means <- function(df) {
  df %>%
    filter(!is_consistent, !is_catch_trial) %>%
    mutate(
      choice_type = case_when(
        chose_gems       ~ "Chose Gems",
        chose_other      ~ "Chose Other",
        chose_onlystones ~ "Chose Only Stones",
        TRUE             ~ NA_character_
      )
    ) %>%
    filter(!is.na(choice_type)) %>%
    group_by(subj_id, choice_type) %>%
    summarise(mean_conf_per_subj = mean(confidence, na.rm = TRUE), .groups = "drop")
}

subject_means_all <- compute_subject_means(phase_2_df)
subject_means_r   <- compute_subject_means(filter(phase_2_df, mentioned_rule))
subject_means_nr  <- compute_subject_means(filter(phase_2_df, !mentioned_rule))

make_summary <- function(subject_means_df) {
  subject_means_df %>%
    group_by(choice_type) %>%
    summarise(
      mean_confidence = mean(mean_conf_per_subj, na.rm = TRUE),
      n = n(),
      sem = sd(mean_conf_per_subj, na.rm = TRUE)/sqrt(n),
      .groups = "drop"
    )
}

plot_data_all <- make_summary(subject_means_all)
plot_data_r   <- make_summary(subject_means_r)
plot_data_nr  <- make_summary(subject_means_nr)

choice_colors <- c(
  "Chose Gems" = "forestgreen",
  "Chose Other" = "royalblue",
  "Chose Only Stones" = "firebrick"
)

all_max <- max(
  c(plot_data_all$mean_confidence + plot_data_all$sem,
    plot_data_r$mean_confidence + plot_data_r$sem,
    plot_data_nr$mean_confidence + plot_data_nr$sem),
  na.rm = TRUE
)
y_limits <- c(0, ifelse(is.finite(all_max), all_max * 1.20, 1))

build_conf_plot <- function(summary_df, plot_title) {
  ggplot(summary_df, aes(x = choice_type, y = mean_confidence, fill = choice_type)) +
    geom_bar(stat = "identity", show.legend = FALSE, alpha = 0.8) +
    geom_errorbar(aes(ymin = mean_confidence - sem, ymax = mean_confidence + sem), width = 0.2) +
    scale_fill_manual(values = choice_colors) +
    labs(
      title = plot_title,
      subtitle = "Confidence scores from Part 2 (consistent trials only)",
      x = "Participant's Choice",
      y = "Mean Confidence"
    ) +
    scale_y_continuous(limits = y_limits, expand = c(0, 0)) +
    theme_minimal() +
    theme(plot.title = element_text(size = 12, face = "bold"))
}

plot_all <- build_conf_plot(plot_data_all, "All Participants")
plot_r   <- build_conf_plot(plot_data_r, "Participants Who Mentioned the Rule")
plot_nr  <- build_conf_plot(plot_data_nr, "Participants Who Did NOT Mention the Rule")

compute_pairwise_safe_with_stars <- function(subject_means_df,
                                             measure = "mean_conf_per_subj",
                                             group = "choice_type",
                                             subj = "subj_id",
                                             test = c("t.test", "wilcox"),
                                             paired = TRUE,
                                             p_adjust = "bonferroni",
                                             summary_df = NULL,
                                             pair_order = NULL,
                                             y_offset = 0.02) {
  test <- match.arg(test)
  subject_means_df[[group]] <- factor(subject_means_df[[group]])
  levels_vec <- levels(subject_means_df[[group]])

  if (!is.null(pair_order)) {
    if (is.data.frame(pair_order) || is.matrix(pair_order)) {
      if (ncol(pair_order) != 2) stop("pair_order must have two columns (group1, group2) in each row.")
      pairs <- matrix(as.character(as.matrix(pair_order)), ncol = 2, byrow = TRUE)
    } else if (is.list(pair_order)) {
      pairs <- do.call(rbind, lapply(pair_order, function(x) {
        if (length(x) != 2) stop("each element in pair_order list must be length 2")
        matrix(as.character(x), nrow = 1)
      }))
    } else if (is.character(pair_order) && length(pair_order) == 2) {
      pairs <- matrix(as.character(pair_order), nrow = 1)
    } else {
      stop("pair_order must be a 2-column matrix/data.frame, a list of 2-char vectors, or a length-2 char vector.")
    }
  } else {
    if (length(levels_vec) < 2) stop("Need at least 2 levels in the group to compare.")
    pairs <- t(combn(levels_vec, 2))
  }

  res <- vector("list", nrow(pairs))
  for (i in seq_len(nrow(pairs))) {
    a <- as.character(pairs[i, 1]); b <- as.character(pairs[i, 2])
    sub_df <- subject_means_df %>% filter(.data[[group]] %in% c(a, b))

    wide <- sub_df %>%
      select(.data[[subj]], .data[[group]], .data[[measure]]) %>%
      pivot_wider(names_from = .data[[group]], values_from = .data[[measure]])

    if (paired) {
      wide_matched <- wide %>% tidyr::drop_na(.data[[a]], .data[[b]])
      x <- wide_matched[[a]]
      y <- wide_matched[[b]]
    } else {
      x <- sub_df %>% filter(.data[[group]] == a) %>% pull(.data[[measure]])
      y <- sub_df %>% filter(.data[[group]] == b) %>% pull(.data[[measure]])
    }

    if (length(x) < 2 || length(y) < 2) {
      tidy_res <- tibble::tibble(
        group1 = a, group2 = b,
        statistic = NA_real_, p.value = NA_real_, method = NA_character_, n1 = length(x), n2 = length(y)
      )
    } else {
      if (test == "t.test") {
        tt <- t.test(x, y, paired = paired)
        tidy_res <- broom::tidy(tt) %>%
          transmute(statistic = statistic, p.value = p.value, method = method, estimate = estimate)
      } else {
        wt <- wilcox.test(x, y, paired = paired)
        tidy_res <- broom::tidy(wt) %>%
          transmute(statistic = statistic, p.value = p.value, method = method, estimate = estimate)
      }
      tidy_res <- tidy_res %>% mutate(n1 = length(x), n2 = length(y), group1 = a, group2 = b)
    }
    res[[i]] <- tidy_res
  }

  pw <- bind_rows(res)

  pw <- pw %>% mutate(p.adj = p.adjust(p.value, method = p_adjust))

  if (is.null(summary_df)) stop("Please supply 'summary_df' (the bar summary) to compute y positions.")

  top_y <- max(summary_df$mean_confidence + summary_df$sem, na.rm = TRUE)
  bottom_y <- min(summary_df$mean_confidence - summary_df$sem, na.rm = TRUE)
  step <- (top_y - bottom_y) * 0.12
  if (!is.finite(step) || step == 0) step <- 0.02

  pw <- pw %>%
    mutate(
      group1 = as.character(group1),
      group2 = as.character(group2),
      xmin = as.numeric(factor(group1, levels = levels_vec)),
      xmax = as.numeric(factor(group2, levels = levels_vec))
    ) %>%
    mutate(idx = seq_len(n()), y.position = top_y + (idx - 1) * step + y_offset) %>%
    mutate(p.signif = case_when(
      is.na(p.adj) ~ "n.s.",
      p.adj < 0.001 ~ "***",
      p.adj < 0.01  ~ "**",
      p.adj < 0.05  ~ "*",
      TRUE ~ "n.s."
    ))

  pw
}

pair_order <- matrix(
  c(
    "Chose Gems",        "Chose Other",
    "Chose Only Stones", "Chose Other",
    "Chose Gems",        "Chose Only Stones"
  ),
  ncol = 2,
  byrow = TRUE
)

pw_all_stars <- compute_pairwise_safe_with_stars(
  subject_means_all,
  measure = "mean_conf_per_subj",
  group = "choice_type",
  subj = "subj_id",
  test = "t.test",
  paired = TRUE,
  p_adjust = "bonferroni",
  summary_df = plot_data_all,
  pair_order = pair_order
)

plot_all_sigstars <- plot_all +
  ggpubr::stat_pvalue_manual(
    pw_all_stars,
    label = "p.signif",
    xmin = "xmin", xmax = "xmax",
    y.position = "y.position",
    tip.length = 0.02,
    bracket.size = 0.4
  )

print(plot_all_sigstars)

pw_r_stars  <- compute_pairwise_safe_with_stars(subject_means_r,  summary_df = plot_data_r,  pair_order = pair_order, paired = TRUE)
pw_nr_stars <- compute_pairwise_safe_with_stars(subject_means_nr, summary_df = plot_data_nr, pair_order = pair_order, paired = TRUE)

plot_r_sigstars  <- plot_r  + ggpubr::stat_pvalue_manual(pw_r_stars,  label = "p.signif", xmin="xmin", xmax="xmax", y.position="y.position", tip.length=0.02)
plot_nr_sigstars <- plot_nr + ggpubr::stat_pvalue_manual(pw_nr_stars, label = "p.signif", xmin="xmin", xmax="xmax", y.position="y.position", tip.length=0.02)

print(plot_r_sigstars); print(plot_nr_sigstars)

```

### Hypothesis 10 (RT: CORRECT VERSUS INCORRECT (all trials))

```{r hypothesis_9, include = FALSE, cache=F}


rt_pooled_correct <-phase_2_df %>%
  ungroup() %>%  
  filter(!is_catch_trial, is_consistent) %>%
  summarise(pooled_median_rt = median(rt, na.rm = TRUE)) %>%
  pull(pooled_median_rt)

r_rt_pooled_correct<- phase_2_df %>%
  ungroup() %>%  
  filter(!is_catch_trial, is_consistent, mentioned_rule) %>%
  summarise(pooled_median_rt = median(rt, na.rm = TRUE)) %>%
  pull(pooled_median_rt)

nr_rt_pooled_correct <-phase_2_df %>%
  ungroup() %>%  
  filter(!is_catch_trial,!mentioned_rule, is_consistent) %>%
  summarise(pooled_median_rt = median(rt, na.rm = TRUE)) %>%
  pull(pooled_median_rt)

rt_pooled_incorrect <- phase_2_df %>%
  ungroup() %>%  
  filter(!is_catch_trial, !is_consistent) %>%
  summarise(pooled_median_rt = median(rt, na.rm = TRUE)) %>%
  pull(pooled_median_rt)

r_rt_pooled_incorrect <- phase_2_df %>%
  ungroup() %>%  
  filter(!is_catch_trial, mentioned_rule, !is_consistent) %>%
  summarise(pooled_median_rt = median(rt, na.rm = TRUE)) %>%
  pull(pooled_median_rt)

nr_rt_pooled_incorrect <- phase_2_df %>%
  ungroup() %>%  
  filter(!is_catch_trial, !mentioned_rule, !is_consistent) %>%
  summarise(pooled_median_rt = median(rt, na.rm = TRUE)) %>%
  pull(pooled_median_rt)

rt_correct <- phase_2_df %>%
  group_by(subj_id) %>%
  filter(!is_catch_trial, is_consistent) %>%
  summarise(median_rt = median(rt, na.rm=TRUE))

r_rt_correct <- phase_2_df %>%
  group_by(subj_id) %>%
  filter(!is_catch_trial, is_consistent, mentioned_rule) %>%
  summarise(median_rt = median(rt, na.rm=TRUE))

nr_rt_correct <- phase_2_df %>%
  group_by(subj_id) %>%
  filter(!is_catch_trial, is_consistent, !mentioned_rule) %>%
  summarise(median_rt = median(rt, na.rm=TRUE))

rt_incorrect <- phase_2_df %>%
  group_by(subj_id) %>%
  filter(!is_catch_trial, !is_consistent) %>%
  summarise(median_rt = median(rt, na.rm=TRUE))

r_rt_incorrect <- phase_2_df %>%
  group_by(subj_id) %>%
  filter(!is_catch_trial, !is_consistent, mentioned_rule) %>%
  summarise(median_rt = median(rt, na.rm=TRUE))

nr_rt_incorrect <- phase_2_df %>%
  group_by(subj_id) %>%
  filter(!is_catch_trial, !is_consistent, !mentioned_rule) %>%
  summarise(median_rt = median(rt, na.rm=TRUE))





```

In part 2, the median reaction time for consistent choices was `r rt_pooled_correct%>% printnum()`and for inconsistent choices was `r rt_pooled_incorrect%>%printnum()`. A paired t-test reveals that there is a significant difference between median reaction times (`r apa_print(wilcox.test(x =rt_correct$median_rt,y = rt_incorrect$median_rt,alternative = "two.sided",paired = TRUE))$statistic`). 

For participants who mentioned rule, the median reaction time for consistent choices was `r r_rt_pooled_correct%>% printnum()`and for inconsistent choices was `r r_rt_pooled_incorrect%>%printnum()`. A paired t-test reveals that there is a significant difference between median reaction times (`r apa_print(wilcox.test(x =r_rt_correct$median_rt,y = r_rt_incorrect$median_rt,alternative = "two.sided",paired = TRUE))$statistic`).

For participants who did not mention the rule, the median reaction time for consistent choices was `r nr_rt_pooled_correct%>% printnum()`and for inconsistent choices was `r nr_rt_pooled_incorrect%>%printnum()`. A paired t-test reveals that there is a significant difference between median reaction times (`r apa_print(wilcox.test(x =nr_rt_correct$median_rt,y = nr_rt_incorrect$median_rt,alternative = "two.sided",paired = TRUE))$statistic`).

### Hypothesis 11 (RT: GEMS VERSUS STONES (for correct trials))

```{r hypothesis_10, include = FALSE, cache=F}

rt_pooled_gems <- phase_2_df %>%
  ungroup() %>%  
  filter(!is_catch_trial, is_consistent, chose_gems) %>%
  summarise(pooled_median_rt = median(rt, na.rm = TRUE)) %>%
  pull(pooled_median_rt)

r_rt_pooled_gems <- phase_2_df %>%
  ungroup() %>%  
  filter(!is_catch_trial, is_consistent, chose_gems, mentioned_rule) %>%
  summarise(pooled_median_rt = median(rt, na.rm = TRUE)) %>%
  pull(pooled_median_rt)

nr_rt_pooled_gems <- phase_2_df %>%
  ungroup() %>%  
  filter(!is_catch_trial, !mentioned_rule, is_consistent, chose_gems) %>%
  summarise(pooled_median_rt = median(rt, na.rm = TRUE)) %>%
  pull(pooled_median_rt)

rt_pooled_nongems <- phase_2_df %>%
  ungroup() %>%  
  filter(!is_catch_trial, is_consistent, !chose_gems) %>%
  summarise(pooled_median_rt = median(rt, na.rm = TRUE)) %>%
  pull(pooled_median_rt)

r_rt_pooled_nongems <- phase_2_df %>%
  ungroup() %>%  
  filter(!is_catch_trial, mentioned_rule,  is_consistent, !chose_gems) %>%
  summarise(pooled_median_rt = median(rt, na.rm = TRUE)) %>%
  pull(pooled_median_rt)

nr_rt_pooled_nongems <- phase_2_df %>%
  ungroup() %>%  
  filter(!is_catch_trial, !mentioned_rule, is_consistent, !chose_gems) %>%
  summarise(pooled_median_rt = median(rt, na.rm = TRUE)) %>%
  pull(pooled_median_rt)

rt_gems <- phase_2_df %>%
  group_by(subj_id) %>%
  filter(!is_catch_trial, is_consistent, chose_gems) %>%
  summarise(median_rt = median(rt, na.rm=TRUE))

r_rt_gems <- phase_2_df %>%
  group_by(subj_id) %>%
  filter(!is_catch_trial, is_consistent, chose_gems, mentioned_rule) %>%
  summarise(median_rt = median(rt, na.rm=TRUE))

nr_rt_gems <- phase_2_df %>%
  group_by(subj_id) %>%
  filter(!is_catch_trial, is_consistent, chose_gems, !mentioned_rule) %>%
  summarise(median_rt = median(rt, na.rm=TRUE))

rt_nongems <- phase_2_df %>%
  group_by(subj_id) %>%
  filter(!is_catch_trial, is_consistent, !chose_gems) %>%
  summarise(median_rt = median(rt, na.rm=TRUE))

r_rt_nongems <- phase_2_df %>%
  group_by(subj_id) %>%
  filter(!is_catch_trial, is_consistent, !chose_gems, mentioned_rule) %>%
  summarise(median_rt = median(rt, na.rm=TRUE))


nr_rt_nongems <- phase_2_df %>%
  group_by(subj_id) %>%
  filter(!is_catch_trial, is_consistent, !chose_gems, !mentioned_rule) %>%
  summarise(median_rt = median(rt, na.rm=TRUE))

rt_gem_diff <- rt_gems %>%
  rename(median_rt_gems = median_rt) %>%
  inner_join(rt_nongems %>%
               rename(median_rt_nongems = median_rt),
             by = "subj_id")

r_rt_gem_diff <- r_rt_gems %>%
  rename(median_rt_gems = median_rt) %>%
  inner_join(r_rt_nongems %>%
               rename(median_rt_nongems = median_rt),
             by = "subj_id")

nr_rt_gem_diff <-nr_rt_gems %>%
  rename(median_rt_gems = median_rt) %>%
  inner_join(nr_rt_nongems %>%
               rename(median_rt_nongems = median_rt),
             by = "subj_id")





```

In part 2, the mean median reaction time when choosing gems in correct responses was `r printnum(mean(rt_gems$median_rt))`and for choosing stones was `r printnum(mean(rt_nongems$median_rt))`. A paired t-test reveals that there is significant difference between median reaction times (`r apa_print(wilcox.test(rt_gem_diff$median_rt_gems,rt_gem_diff$median_rt_nongems,paired = TRUE))$statistic`). 

For participants that mentioned the rule, the median reaction time when choosing gems in correct responses was `r printnum(mean(r_rt_gems$median_rt))`and for choosing stones was `r printnum(mean(r_rt_nongems$median_rt))`. A paired t-test reveals that there is significant difference between median reaction times (`r apa_print(wilcox.test(r_rt_gem_diff$median_rt_gems,r_rt_gem_diff$median_rt_nongems,paired = TRUE))$statistic`). 

For participants that did not mention the rule,  the median reaction time when choosing gems in correct responses was `r printnum(mean(nr_rt_gems$median_rt))`and for choosing stones was `r printnum(mean(nr_rt_nongems$median_rt))`. A paired t-test reveals that there is no significant difference between median reaction times (`r apa_print(wilcox.test(nr_rt_gem_diff$median_rt_gems,nr_rt_gem_diff$median_rt_nongems,paired = TRUE))$statistic`). 


### Hypothesis 12 (RT: STONES VERSUS STONES ONLY)

```{r hypothesis_12, include = FALSE, cache=F}
rt_other <- phase_2_df %>%
  group_by(subj_id) %>%
  filter(!is_catch_trial, is_consistent,chose_other) %>%
  summarise(median_rt = median(rt, na.rm=TRUE))

r_rt_other <- phase_2_df %>%
  group_by(subj_id) %>%
  filter(!is_catch_trial, mentioned_rule, is_consistent,chose_other) %>%
  summarise(median_rt = median(rt, na.rm=TRUE))

nr_rt_other <- phase_2_df %>%
  group_by(subj_id) %>%
  filter(!is_catch_trial, is_consistent,chose_other, !mentioned_rule) %>%
  summarise(median_rt = median(rt, na.rm=TRUE))


rt_stones <- phase_2_df %>%
  group_by(subj_id) %>%
  filter(!is_catch_trial, is_consistent,chose_onlystones) %>%
  summarise(median_rt = median(rt, na.rm=TRUE))

r_rt_stones <- phase_2_df %>%
  group_by(subj_id) %>%
  filter(!is_catch_trial, is_consistent,chose_onlystones, mentioned_rule) %>%
  summarise(median_rt = median(rt, na.rm=TRUE))

nr_rt_stones <- phase_2_df %>%
  group_by(subj_id) %>%
  filter(!is_catch_trial, is_consistent,chose_onlystones, !mentioned_rule) %>%
  summarise(median_rt = median(rt, na.rm=TRUE))

rt_joined <- rt_gems %>%
  rename(median_rt_other = median_rt) %>%
  inner_join(rt_stones %>%
               rename(median_rt_stones = median_rt),
             by = "subj_id")

r_rt_joined <- r_rt_gems %>%
  rename(median_rt_other = median_rt) %>%
  inner_join(r_rt_stones %>%
               rename(median_rt_stones = median_rt),
             by = "subj_id")

nr_rt_joined <- nr_rt_gems %>%
  rename(median_rt_other = median_rt) %>%
  inner_join(nr_rt_stones %>%
               rename(median_rt_stones = median_rt),
             by = "subj_id")



```

In part 2, the mean median reaction time when choosing the other position in correct responses was `r printnum(mean(rt_other$median_rt))`and for only stones was `r printnum(mean(rt_stones$median_rt))`. A paired t-test reveals that there is a significant difference between median reaction times (`r apa_print(wilcox.test(rt_joined$median_rt_other, rt_joined$median_rt_stones, alternative = "two.sided", paired = TRUE))$statistic`). 

For participants who mentioned the rule, the median reaction time when choosing the other position in correct responses was `r printnum(mean(r_rt_other$median_rt))`and for only stones was `r printnum(mean(r_rt_stones$median_rt))`. A paired t-test reveals that there is a significant difference between median reaction times (`r apa_print(wilcox.test(r_rt_joined$median_rt_other, r_rt_joined$median_rt_stones, alternative = "two.sided", paired = TRUE))$statistic`). 

For participants who did not mention the rule, the median reaction time when choosing the other position in correct responses was `r printnum(mean(nr_rt_other$median_rt))`and for only stones was `r printnum(mean(nr_rt_stones$median_rt))`. A paired t-test reveals that there is a significant difference between median reaction times (`r apa_print(wilcox.test(nr_rt_joined$median_rt_other, nr_rt_joined$median_rt_stones, alternative = "two.sided", paired = TRUE))$statistic`). 



```{r rtgraph, cache=F, fig.width=8}


compute_plot_data <- function(df) {
  df %>%
    filter(!is_consistent, !is_catch_trial) %>%   # <--- use !is_consistent
    mutate(
      choice_type = case_when(
        chose_gems       ~ "Chose Gems",
        chose_other      ~ "Chose Other",
        chose_onlystones ~ "Chose Only Stones",
        TRUE             ~ NA_character_
      )
    ) %>%
    filter(!is.na(choice_type)) %>%
    group_by(subj_id, choice_type) %>%
    summarise(mean_conf_per_subj = mean(confidence, na.rm = TRUE), .groups = "drop") %>%
    group_by(choice_type) %>%
    summarise(
      mean_confidence = mean(mean_conf_per_subj, na.rm = TRUE),
      n = n(),
      sem = sd(mean_conf_per_subj, na.rm = TRUE) / sqrt(n),
      .groups = "drop"
    )
}

plot_data <- phase_2_df %>%
  filter(is_consistent, !is_catch_trial) %>%
  mutate(
    choice_type = case_when(
      chose_gems       ~ "Chose Gems",
      chose_other      ~ "Chose Other",
      chose_onlystones ~ "Chose Only Stones",
      TRUE             ~ NA_character_ 
    )
  ) %>%
  filter(!is.na(choice_type)) %>%
  group_by(choice_type) %>%
  summarise(
    median_rt = median(rt, na.rm = TRUE),
    n = n(),
    sem = sd(rt, na.rm = TRUE) / sqrt(n),
    .groups = 'drop'
  )

rt_plot <- ggplot(plot_data, aes(x = choice_type, y = median_rt, fill = choice_type)) +
  geom_bar(stat = "identity", show.legend = FALSE) +
  geom_errorbar(
    aes(ymin = median_rt - sem, ymax = median_rt + sem),
    width = 0.2
  ) +
  labs(
    title = "Median RT by Choice",
    x = "Participant's Choice",
    y = "Median rt"
  ) +
  theme_minimal()

print(rt_plot)


```

```{r rtdistribution, cache=F, fig.width=8}
library(dplyr)
library(ggplot2)

plot_data <- phase_2_df %>%
  filter(is_consistent, !is_catch_trial) %>%
  mutate(
    choice_type = case_when(
      chose_gems       ~ "Chose Gems",
      chose_other      ~ "Chose Other",
      chose_onlystones ~ "Chose Only Stones",
      TRUE             ~ NA_character_
    )
  ) %>%
  filter(!is.na(choice_type)) %>%
  group_by(subj_id, choice_type) %>%
  summarise(median_rt = median(rt, na.rm = TRUE), .groups = "drop")

summary_data <- plot_data %>%
  group_by(choice_type) %>%
  summarise(
    mean_rt = mean(median_rt, na.rm = TRUE),
    sem = sd(median_rt, na.rm = TRUE) / sqrt(n()),
    .groups = "drop"
  )

choice_colors <- c(
  "Chose Gems" = "forestgreen",
  "Chose Other" = "royalblue",
  "Chose Only Stones" = "firebrick"
)

rt_plot <- ggplot(plot_data, aes(x = choice_type, y = median_rt, fill = choice_type)) +
  geom_boxplot(width = 0.6, alpha = 0.7, color = "gray30", outlier.shape = NA) +
  geom_errorbar(
    data = summary_data,
    mapping = aes(x = choice_type, ymin = mean_rt - sem, ymax = mean_rt + sem),
    width = 0.15,
    size = 0.6,
    inherit.aes = FALSE
  ) +
  geom_point(
    data = summary_data,
    mapping = aes(x = choice_type, y = mean_rt),
    shape = 21, size = 3.5, fill = "white", color = "black", stroke = 1,
    inherit.aes = FALSE
  ) +
  scale_fill_manual(values = choice_colors) +
  labs(
    title = "Mean Median Reaction Times by Choice Type",
    x = "Participant's Choice",
    y = "Mean Median Reaction Time (RT)"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold"),
    legend.position = "none"
  )

print(rt_plot)


```


## Exploratory analysis 

### Hypothesis 8.5 (CONFIDENCE: STONES VERSUS STONES ONLY)
```{r hypothesis_8.5, include = FALSE, cache=F}
otherstone_conf <- phase_2_df %>% 
  group_by(subj_id) %>%
  filter(!is_catch_trial, is_consistent, chose_other) %>%
  summarise(mean_conf=mean(confidence),
            median = median(confidence))

r_otherstone_conf <- phase_2_df %>% 
  group_by(subj_id) %>%
  filter(!is_catch_trial, is_consistent, chose_other, mentioned_rule) %>%
  summarise(mean_conf=mean(confidence),
            median = median(confidence))

nr_otherstone_conf <- phase_2_df %>% 
  group_by(subj_id) %>%
  filter(!is_catch_trial, is_consistent, chose_other, !mentioned_rule) %>%
  summarise(mean_conf=mean(confidence),
            median = median(confidence))


onlystone_conf <- phase_2_df %>% 
  group_by(subj_id) %>%
  filter(!is_catch_trial, is_consistent, chose_onlystones) %>%
  summarise(mean_conf=mean(confidence),
            median = median(confidence))

r_onlystone_conf <- phase_2_df %>% 
  group_by(subj_id) %>%
  filter(!is_catch_trial, is_consistent, chose_onlystones, mentioned_rule) %>%
  summarise(mean_conf=mean(confidence),
            median = median(confidence))

nr_onlystone_conf <- phase_2_df %>% 
  group_by(subj_id) %>%
  filter(!is_catch_trial, is_consistent, chose_onlystones, !mentioned_rule) %>%
  summarise(mean_conf=mean(confidence),
            median = median(confidence))

conf_df <- otherstone_conf %>%
  rename(mean_conf_other = mean_conf, median_other = median) %>%
  inner_join(
    onlystone_conf %>%
      rename(mean_conf_onlystone = mean_conf, median_onlystone = median),
    by = "subj_id"
  )

r_conf_df <- r_otherstone_conf %>%
  rename(mean_conf_other = mean_conf, median_other = median) %>%
  inner_join(
    r_onlystone_conf %>%
      rename(mean_conf_onlystone = mean_conf, median_onlystone = median),
    by = "subj_id"
  )

nr_conf_df <- nr_otherstone_conf %>%
  rename(mean_conf_other = mean_conf, median_other = median) %>%
  inner_join(
    nr_onlystone_conf %>%
      rename(mean_conf_onlystone = mean_conf, median_onlystone = median),
    by = "subj_id"
  )


```

In part 2, mean confidence when touching the stones only position on trials in which participants were correct was `r onlystone_conf$mean_conf %>%mean()%>%printnum()` (median=`r onlystone_conf$mean_conf%>%median()%>%printnum() `). In trials which they chose other stone position, it was `r otherstone_conf$mean_conf %>%mean()%>%printnum()` (median=`r  otherstone_conf$mean_conf%>%median()%>%printnum() `). This difference was not significant (`r apa_print(t.test(conf_df$mean_conf_other, conf_df$mean_conf_onlystone, paired = TRUE))$statistic`).

For participants who mentioned the rule, mean confidence when touching the stones only position on trials in which participants were correct was `r r_onlystone_conf$mean_conf %>%mean()%>%printnum()` (median=`r r_onlystone_conf$mean_conf%>%median()%>%printnum() `). In trials which they chose other stone position, it was `r r_otherstone_conf$mean_conf %>%mean()%>%printnum()` (median=`r  r_otherstone_conf$mean_conf%>%median()%>%printnum() `). This difference was not significant (`r apa_print(t.test(r_conf_df$mean_conf_other, r_conf_df$mean_conf_onlystone, paired = TRUE))$statistic`).

For participants who did not mention the rule, mean confidence when touching the stones only position on trials in which participants were correct was `r nr_onlystone_conf$mean_conf %>%mean()%>%printnum()` (median=`r nr_onlystone_conf$mean_conf%>%median()%>%printnum() `). In trials which they chose other stone position, it was `r nr_otherstone_conf$mean_conf %>%mean()%>%printnum()` (median=`r  nr_otherstone_conf$mean_conf%>%median()%>%printnum() `). This difference was not significant (`r apa_print(t.test(nr_conf_df$mean_conf_other, nr_conf_df$mean_conf_onlystone, paired = TRUE))$statistic`).

### estimates of own performance


```{r correlation, include = FALSE, cache=F}
## Correlation between estimate and actual performance 
# part 1

total_chose_gems_part1 <- phase_1_df %>%
  group_by(subj_id) %>%
  summarise(total_gems = sum(chose_gems, na.rm = TRUE)) 

estimate_chose_gems_part1 <- phase_1_df %>% 
  group_by(subj_id) %>%
  summarise(mean_estimate = mean(number_correct_part1, na.rm = TRUE))

correlation_data <- inner_join(total_chose_gems_part1, 
                               estimate_chose_gems_part1, 
                               by = "subj_id")

cor_part1 <- cor.test(x = correlation_data$total_gems, 
         y = correlation_data$mean_estimate, 
         method = "pearson")

# part 2

total_chose_gems_part2 <- phase_2_df %>%
  group_by(subj_id) %>%
  summarise(consistent = sum(is_consistent, na.rm = TRUE)) 

estimate_chose_gems_part2 <- phase_2_df %>% 
  group_by(subj_id) %>%
  summarise(mean_estimate = mean(number_correct_part2, na.rm = TRUE))

correlation_data <- inner_join(total_chose_gems_part2, 
                               estimate_chose_gems_part2, 
                               by = "subj_id")

cor_part2 <- cor.test(x = correlation_data$consistent, 
         y = correlation_data$mean_estimate, 
         method = "pearson")

# correlation bbetween probability of choosing gems in part 2 and part 1 estimate 

total_chose_gems_part2 <- phase_2_df %>% 
  group_by(subj_id) %>% 
  summarise(chose_gems = sum(chose_gems, na.rm = T))

cor_part2_part1_data <- inner_join(total_chose_gems_part2,
                                   estimate_chose_gems_part1,
                                   by = "subj_id"
                            )

cor_part2_part1 <- cor.test(x = cor_part2_part1_data$chose_gems,
                            y = cor_part2_part1_data$mean_estimate,
                            method = "pearson")

```


People can estimate performance in part 1; Pearson correlation `r sprintf("r = %.2f, 95%% CI [%.2f, %.2f]", cor_part1$estimate, cor_part1$conf.int[1], cor_part1$conf.int[2])` p `r printp(cor_part1$p.value)`. In Part 2, they are also able to estimate their performance in part 2; Pearson correlation `r sprintf("r = %.2f, 95%% CI [%.2f, %.2f]", cor_part2$estimate,cor_part2$conf.int[1], cor_part2$conf.int[2])` p `r printp(cor_part2$p.value)`. 




```{r estimate,  cache=F, fig.width=8}

library(ggplot2)
library(patchwork)

cor_label <- function(x, y) {
  ct <- cor.test(x, y, method = "pearson")
  n <- sum(complete.cases(x, y))
  paste0(
    "r = ", round(ct$estimate, 2), "\n",
    "p = ", signif(ct$p.value, 3), "\n",
    "n = ", n
  )
}


total_chose_gems_part1 <- phase_1_df %>%
  group_by(subj_id) %>%
  summarise(total_gems = sum(chose_gems, na.rm = TRUE))

estimate_chose_gems_part1 <- phase_1_df %>%
  group_by(subj_id) %>%
  summarise(mean_estimate = mean(number_correct_part1, na.rm = TRUE))

correlation_data_part1 <- inner_join(total_chose_gems_part1,
                                     estimate_chose_gems_part1,
                                     by = "subj_id")
label_part1 <- cor_label(
  correlation_data_part1$total_gems,
  correlation_data_part1$mean_estimate
)

plot_part1 <- ggplot(correlation_data_part1, aes(x = total_gems, y = mean_estimate)) +
  geom_point(alpha = 0.7) +
  geom_smooth(method = "lm", col = "blue", se = FALSE) +
  geom_abline(slope = 1, intercept = 0) +
  annotate(
    "text",
    x = Inf, y = Inf,
    label = label_part1,
    hjust = 1.1, vjust = 1.1,
    size = 4
  ) +
  labs(
    title = "Part 1: Correct Trials vs. Estimate",
    x = "Number of Correct Trials",
    y = "Estimate (Part 1)"
  ) +
  theme_minimal() +
  coord_cartesian(xlim = c(0, 50), ylim = c(0, 50))


total_chose_gems_part2 <- phase_2_df %>%
  group_by(subj_id) %>%
  summarise(consistent = sum(is_consistent, na.rm = TRUE))

estimate_chose_gems_part2 <- phase_2_df %>%
  group_by(subj_id) %>%
  summarise(mean_estimate = mean(number_correct_part2, na.rm = TRUE))

correlation_data_part2 <- inner_join(total_chose_gems_part2,
                                     estimate_chose_gems_part2,
                                     by = "subj_id")
label_part2 <- cor_label(
  correlation_data_part2$consistent,
  correlation_data_part2$mean_estimate
)

plot_part2 <- ggplot(correlation_data_part2, aes(x = consistent, y = mean_estimate)) +
  geom_point(alpha = 0.7) +
  geom_smooth(method = "lm", col = "red", se = FALSE) +
  geom_abline(slope = 1, intercept = 0) +
  annotate(
    "text",
    x = Inf, y = Inf,
    label = label_part2,
    hjust = 1.1, vjust = 1.1,
    size = 4
  ) +
  labs(
    title = "Part 2: Consistency vs. Estimate",
    x = "Total Consistent Choices",
    y = "Estimate (Part 2)"
  ) +
  theme_minimal() +
  coord_cartesian(xlim = c(0, 50), ylim = c(0, 50))


combined_plot <- plot_part1 + plot_part2

print(combined_plot)

label_part2_part1 <- cor_label(
  cor_part2_part1_data$chose_gems,
  cor_part2_part1_data$mean_estimate
)

plot_part2_part1 <- ggplot(
  cor_part2_part1_data,
  aes(x = chose_gems, y = mean_estimate)
) +
  geom_point(alpha = 0.7) +
  geom_smooth(method = "lm", col = "purple", se = FALSE) +
  geom_abline(slope = 1, intercept = 0) +
  annotate(
    "text",
    x = Inf, y = Inf,
    label = label_part2_part1,
    hjust = 1.1, vjust = 1.1,
    size = 4
  ) +
  labs(
    title = "Part 2 Gems Chosen vs. Part 1 Estimate",
    x = "Total Gems Chosen (Part 2)",
    y = "Mean Estimate (Part 1)"
  ) +
  theme_minimal() +
  coord_cartesian(xlim = c(0, 50), ylim = c(0, 50))




print(plot_part2_part1)

```


### Effect of numbers on choice

```{r higherorlower, include = FALSE, cache=F}

higher_df <- phase_2_df %>%
  group_by(subj_id) %>%
  filter(!is_catch_trial, !is_consistent) %>%
  mutate(
    higher = case_when(
      response_position == "middle" & correct_choice_position == "left"  ~ middle_number > right_number,
      response_position == "middle" & correct_choice_position == "right" ~ middle_number > left_number,
      
      response_position == "right" & correct_choice_position == "left"   ~ right_number > middle_number,
      response_position == "right" & correct_choice_position == "middle" ~ right_number > left_number,
      
      response_position == "left" & correct_choice_position == "middle" ~ left_number > right_number,
      response_position == "left" & correct_choice_position == "right"  ~ left_number > middle_number,
      
      # Default case (e.g., if a condition wasn't met, or if you had a TRUE ~ FALSE)
      # This will result in NA if no condition is met, which is fine.
      TRUE ~ NA
    )
  ) %>%
  filter(
    case_when(
      response_position == "middle" & correct_choice_position == "left"  ~ middle_number != right_number,
      response_position == "middle" & correct_choice_position == "right" ~ middle_number != left_number,
      
      response_position == "right" & correct_choice_position == "left"   ~ right_number != middle_number,
      response_position == "right" & correct_choice_position == "middle" ~ right_number != left_number,
      
      response_position == "left" & correct_choice_position == "middle" ~ left_number != right_number,
      response_position == "left" & correct_choice_position == "right"  ~ left_number != middle_number,
      
      TRUE ~ FALSE
    )
  )



participant_summary <- higher_df %>%
  group_by(subj_id) %>%
  summarise(
    prop_higher = mean(higher, na.rm = TRUE)
  )


```

When participants were inaccurate in part 2, the mean probability for the participant to choose the option with a greater number was `r higher_df %>% ungroup() %>% summarise(mean=mean(higher)) %>% pull(mean) %>% printnum()`. This was significantly less than chance (`r apa_print(t.test(participant_summary$prop_higher,mu = 0.5,alternative = "two.sided"))$statistic`). 


```{r age,  cache=F, fig.width=8}
library(dplyr)
library(ggplot2)


participant_accuracy_1 <- merged_data %>%
  filter(phase == 'part1', !is_catch_trial) %>%
  group_by(`Participant id`, Age) %>%
  summarise(participant_acc = mean(chose_gems), .groups = 'drop')

participant_accuracy_2 <- merged_data %>%
  filter(phase == 'part2', !is_catch_trial) %>%
  group_by(`Participant id`, Age) %>%
  summarise(participant_acc = mean(is_consistent), .groups = 'drop')

# Avoidance
age_avoid_1 <- merged_data %>%
  filter(phase == 'part1', !is_catch_trial) %>%
  group_by(`Participant id`, Age) %>%
  summarise(participant_avoid = mean(chose_onlystones), .groups = 'drop')

age_avoid_2 <- merged_data %>%
  filter(phase == 'part2', !is_catch_trial) %>%
  group_by(`Participant id`, Age) %>%
  summarise(participant_avoid = mean(chose_onlystones), .groups = 'drop')

age_breaks <- c(17, 22, 27, 32, 37, 42, 47, 52, 57, 62, 67, 72, 77, Inf)
age_labels <- c("18-22", "23-27", "28-32", "33-37", "38-42", "43-47",
                "48-52", "53-57", "58-62", "63-67", "68-72", "73-77", "78+")


combined_participant_data <- bind_rows(
  participant_accuracy_1 %>% mutate(Condition = "Part 1"),
  participant_accuracy_2 %>% mutate(Condition = "Part 2")
)

combined_age_summary <- combined_participant_data %>%
  mutate(
    Age = as.numeric(as.character(Age)),
    AgeBin = cut(Age, breaks = age_breaks, labels = age_labels, right = TRUE, ordered_result = TRUE)
  ) %>%
  filter(!is.na(AgeBin)) %>%
  group_by(AgeBin, Condition) %>% 
  summarise(
    mean_accuracy = mean(participant_acc),
    se = sd(participant_acc) / sqrt(n()),
    .groups = 'drop'
  )

ggplot(combined_age_summary, aes(x = AgeBin, 
                                 y = mean_accuracy, 
                                 group = Condition, 
                                 color = Condition)) +
  geom_line(linewidth = 1.2, position = position_dodge(width = 0.3)) +
  geom_point(size = 3, position = position_dodge(width = 0.3)) +
  geom_errorbar(
    aes(ymin = mean_accuracy - se, ymax = mean_accuracy + se),
    width = 0.25, linewidth = 0.8, position = position_dodge(width = 0.3)
  ) +
  
  scale_color_manual(values = c("Part 1" = "blue", "Part 2" = "red")) +
  
  labs(
    title = "Mean Participant Accuracy by Age Bin",
    x = "Age Bin",
    y = "Mean Accuracy",
    color = "Phase" 
  ) +
  coord_cartesian(ylim = c(0, 1)) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))


combined_avoid_data <- bind_rows(
  age_avoid_1 %>% mutate(Condition = "Part 1"),
  age_avoid_2 %>% mutate(Condition = "Part 2")
)

combined_age_summary_avoid <- combined_avoid_data %>%
  mutate(
    Age = as.numeric(as.character(Age)),
    AgeBin = cut(Age, breaks = age_breaks, labels = age_labels, right = TRUE, ordered_result = TRUE)
  ) %>%
  filter(!is.na(AgeBin)) %>%
  group_by(AgeBin, Condition) %>%
  summarise(
    mean_avoid = mean(participant_avoid), 
    se = sd(participant_avoid) / sqrt(n()),
    .groups = 'drop'
  )

ggplot(combined_age_summary_avoid, aes(x = AgeBin, 
                                       # FIX: Plot the new mean_avoid column
                                       y = mean_avoid, 
                                       group = Condition, 
                                       color = Condition)) +
  geom_line(linewidth = 1.2, position = position_dodge(width = 0.3)) +
  geom_point(size = 3, position = position_dodge(width = 0.3)) +
  geom_errorbar(
    # FIX: Use mean_avoid for error bars
    aes(ymin = mean_avoid - se, ymax = mean_avoid + se), 
    width = 0.25, linewidth = 0.8, position = position_dodge(width = 0.3)
  ) +
  
  # --- THE FIX ---
  # Names here MUST match the names from combined_avoid_data
  scale_color_manual(values = c("Part 1" = "blue", "Part 2" = "red")) +
  
  labs(
    title = "Mean Participant Avoidance by Age Bin",
    x = "Age Bin",
    y = "Mean Avoidance", # FIX: Correct y-axis label
    color = "Phase"
  ) +
  coord_cartesian(ylim = c(0, 1)) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))


# 1. Create a data frame with ONE row per participant
#    We convert Age to numeric, filter out NAs, and then
#    use distinct() to keep only one unique row for each participant.
participant_age_data <- merged_data %>%
  mutate(Age_numeric = as.numeric(as.character(Age))) %>%
  filter(!is.na(Age_numeric)) %>%
  distinct(`Participant id`, Age_numeric)

conditions <- phase_1_df %>%
  ungroup() %>% 
  distinct(subj_id, stoneposition) %>% 
  
  count(stoneposition)

print(conditions)
  

ggplot(participant_age_data, aes(x = Age_numeric)) +
  geom_histogram(binwidth = 5, fill = "cornflowerblue", color = "white", alpha = 0.9, boundary = 0) +
  labs(
    title = "Distribution of Participant Age",
    x = "Age",
    y = "Frequency (Count of Participants)" # Updated label
  ) +
  theme_minimal()


```
```{r agebias, cache=F, fig.width=8}
library(dplyr)
library(ggplot2)


correlated_data <- inner_join(
  phase_1_acc,
  phase_2_acc,
  by = "subj_id",
  suffix = c("_p1", "_p2") 
)


correlation_test <- cor.test(
  ~ acc_p1 + acc_p2, 
  data = correlated_data
)

print(correlation_test)


ggplot(correlated_data, aes(x = acc_p1, y = acc_p2)) +
  geom_point(alpha = 0.6, size = 2) +
  geom_smooth(method = "lm", color = "blue", fill = "blue", alpha = 0.1) +
  labs(
    title = "Correlation between Phase 1 and Phase 2 Accuracy",
    x = "Phase 1 Accuracy",
    y = "Phase 2 Accuracy"
  ) +
  theme_minimal()

```
```{r organize_hover_data, cache=F, fig.width = 8}

sep <- ";"

mouse_long <- raw_df %>%
  filter(!(subj_id %in% participants_to_exclude), trial_type == "testmouse") %>%
  select(subj_id, trial_num, mouse_x, mouse_y, mouse_t) %>%
  mutate(
    mouse_x = if_else(is.na(mouse_x) | mouse_x == "", NA_character_, as.character(mouse_x)),
    mouse_y = if_else(is.na(mouse_y) | mouse_y == "", NA_character_, as.character(mouse_y)),
    mouse_t = if_else(is.na(mouse_t) | mouse_t == "", NA_character_, as.character(mouse_t))
  ) %>%
  mutate(
    x_tokens = str_split(mouse_x, pattern = fixed(sep)),
    y_tokens = str_split(mouse_y, pattern = fixed(sep)),
    t_tokens = str_split(mouse_t, pattern = fixed(sep))
  ) %>%
  filter(!map_lgl(x_tokens, ~ all(is.na(.x)))) %>%
  mutate(
    n_tokens = pmax(map_int(x_tokens, length),
                    map_int(y_tokens, length),
                    map_int(t_tokens, length))
  ) %>%
  rowwise() %>%
  mutate(idx = list(seq_len(n_tokens))) %>%
  ungroup() %>%
  unnest(cols = c(idx)) %>%
  mutate(frame_index = idx) %>%
  mutate(
    x_token = map2_chr(x_tokens, frame_index, ~ if(length(.x) >= .y) .x[[.y]] else NA_character_),
    y_token = map2_chr(y_tokens, frame_index, ~ if(length(.x) >= .y) .x[[.y]] else NA_character_),
    t_token = map2_chr(t_tokens, frame_index, ~ if(length(.x) >= .y) .x[[.y]] else NA_character_)
  ) %>%
  select(subj_id, trial_num, frame_index, x_token, y_token, t_token)

mouse_long <- mouse_long %>%
  mutate(
    mouse_x = as.numeric(str_trim(x_token)),
    mouse_y = as.numeric(str_trim(y_token)),
    mouse_t = as.numeric(str_trim(t_token))
  ) %>%
  select(-x_token, -y_token, -t_token) %>%
  filter(!(is.na(mouse_x) & is.na(mouse_y)))

mouse_movements <- mouse_long %>%
  arrange(subj_id, trial_num, frame_index) %>%
  group_by(subj_id, trial_num) %>%
  mutate(
    x_prev = lag(mouse_x),
    y_prev = lag(mouse_y),
    t_prev = lag(mouse_t),
    dx = as.numeric(mouse_x) - as.numeric(x_prev),
    dy = as.numeric(mouse_y) - as.numeric(y_prev),
    dist = sqrt((dx)^2 + (dy)^2),
    dt = as.numeric(mouse_t) - as.numeric(t_prev),
    speed = ifelse(!is.na(dist) & !is.na(dt) & dt > 0, dist / dt, NA_real_)
  ) %>%
  ungroup()

summary(mouse_movements$dist)
quantile(mouse_movements$dist, probs = c(0, 0.5, 0.9, 0.99, 1), na.rm = TRUE)

mouse_movements %>%
  arrange(desc(dist)) %>%
  slice_head(n = 20) %>%
  select(subj_id, trial_num, frame_index, mouse_x, mouse_y, x_prev, y_prev, dx, dy, dist) %>%
  print(n = 20)

subject_maxdist <- mouse_movements %>%
  group_by(subj_id) %>%
  summarise(
    n_points   = n(),
    n_moves    = sum(!is.na(dist)),
    max_dist   = max(dist, na.rm = TRUE),
    median_dist= median(dist, na.rm = TRUE),
    p99_dist   = quantile(dist, 0.99, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  arrange(desc(max_dist))

print(head(subject_maxdist, 20))

p_hist <- ggplot(subject_maxdist, aes(x = median_dist)) +
  geom_histogram(bins = 60, boundary = 0, closed = "left", fill = "steelblue", color = "white", na.rm = TRUE) +
  labs(title = "Distribution of participants' median frame-to-frame jump",
       x = "median dist (pixels/frame)",
       y = "Number of participants") +
  theme_minimal(base_size = 13)

print(p_hist)

tablet_suspects <- subject_maxdist %>%
  filter(median_dist > 200) %>%
  pull(subj_id)


hover_part1 <- raw_df %>%
  filter(
    !(subj_id %in% participants_to_exclude),
    trial_type == "testmouse",
    phase == 'part1',
    !(subj_id %in% tablet_suspects)
  ) %>%
  select(
    subj_id, 
    trial_num,
    stoneposition,
    left_number, 
    middle_number, 
    right_number,
    is_catch_trial,
    chose_gems,
    chose_other,
    chose_onlystones,
    frames_hover_left,
    frames_hover_middle,
    frames_hover_right,
    frames_hover_none,
    hover_time_left,
    hover_time_middle,
    hover_time_right,
    hover_time_none, 
    mouse_x,
    mouse_y,
    mouse_t,
    mouse_universal_x,
    mouse_universal_y,
    hover_location,
    total_frames,
    correct_choice_position,
    response_position
  ) %>%
  mutate(

    across(starts_with("frames_hover_"), as.numeric),
    across(starts_with("hover_time_"), as.numeric),

    gem_frame_hover = case_when(
      correct_choice_position == "left"   ~ frames_hover_left,
      correct_choice_position == "middle" ~ frames_hover_middle,
      correct_choice_position == "right"  ~ frames_hover_right,
      TRUE ~ NA_real_
    ),
    gem_time_hover = case_when(
      correct_choice_position == "left"   ~ hover_time_left,
      correct_choice_position == "middle" ~ hover_time_middle,
      correct_choice_position == "right"  ~ hover_time_right,
      TRUE ~ NA_real_
    ),
    stone_frame_hover = case_when(
      stoneposition == "left" ~ frames_hover_left,
      stoneposition == "middle" ~ frames_hover_middle,
      stoneposition == "right" ~ frames_hover_right,
      TRUE ~ NA_real_
    ),
    stone_time_hover = case_when(
      stoneposition == "right" ~ hover_time_right,
      stoneposition == "middle" ~ hover_time_middle,
      stoneposition == "left" ~ hover_time_left,
      TRUE ~ NA_real_
    ),
    other_frame_hover = case_when(
      stoneposition == "left" & correct_choice_position == "middle" ~ frames_hover_right,
      stoneposition == "left" & correct_choice_position == "right" ~ frames_hover_middle,
      stoneposition == "middle" & correct_choice_position == "right" ~ frames_hover_left,
      stoneposition == "middle" & correct_choice_position == "left" ~ frames_hover_right,
      stoneposition == "right" & correct_choice_position == "middle" ~ frames_hover_left,
      stoneposition == "right" & correct_choice_position == "left" ~ frames_hover_middle,
      TRUE ~ NA_real_
    ),
    other_time_hover = case_when(
      stoneposition == "left" & correct_choice_position == "middle" ~ hover_time_right,
      stoneposition == "left" & correct_choice_position == "right" ~ hover_time_middle,
      stoneposition == "middle" & correct_choice_position == "right" ~ hover_time_left,
      stoneposition == "middle" & correct_choice_position == "left" ~ hover_time_right,
      stoneposition == "right" & correct_choice_position == "middle" ~ hover_time_left,
      stoneposition == "right" & correct_choice_position == "left" ~ hover_time_middle,
      TRUE ~ NA_real_
    )
    
  )

hover_part2 <- raw_df %>%
  filter(
    !(subj_id %in% participants_to_exclude),
    trial_type == "testmouse",
    phase == 'part2',
    !(subj_id %in% tablet_suspects)
  ) %>%
  select(
    subj_id, 
    trial_num,
    stoneposition,
    left_number, 
    middle_number, 
    right_number,
    is_catch_trial,
    chose_gems,
    chose_other,
    chose_onlystones,
    frames_hover_left,
    frames_hover_middle,
    frames_hover_right,
    frames_hover_none,
    hover_time_left,
    hover_time_middle,
    hover_time_right,
    hover_time_none, 
    mouse_x,
    mouse_y,
    mouse_t,
    mouse_universal_x,
    mouse_universal_y,
    hover_location,
    total_frames,
    correct_choice_position,
    is_consistent,
    response_position
  ) 

part1_correct_position <- phase_1_df %>%
  mutate(
    correct_choice_position_part1 = correct_choice_position
  ) %>%
  select(subj_id, trial_num, correct_choice_position_part1) 
  
hover_part2 <- hover_part2 %>% 
  left_join(part1_correct_position, by = c('subj_id', 'trial_num'))

hover_part2 <- hover_part2 %>% 
  mutate(

    across(starts_with("frames_hover_"), as.numeric),
    across(starts_with("hover_time_"), as.numeric),

    is_consistent_frame_hover = case_when(
      correct_choice_position == "left"   ~ frames_hover_left,
      correct_choice_position == "middle" ~ frames_hover_middle,
      correct_choice_position == "right"  ~ frames_hover_right,
      TRUE ~ NA_real_
    ),
    is_consistent_time_hover = case_when(
      correct_choice_position == "left"   ~ hover_time_left,
      correct_choice_position == "middle" ~ hover_time_middle,
      correct_choice_position == "right"  ~ hover_time_right,
      TRUE ~ NA_real_
    ),
    stone_frame_hover = case_when(
      stoneposition == "left" ~ frames_hover_left,
      stoneposition == "middle" ~ frames_hover_middle,
      stoneposition == "right" ~ frames_hover_right,
      TRUE ~ NA_real_
    ),
    stone_time_hover = case_when(
      stoneposition == "right" ~ hover_time_right,
      stoneposition == "middle" ~ hover_time_middle,
      stoneposition == "left" ~ hover_time_left,
      TRUE ~ NA_real_
    ),
    gem_frame_hover = case_when(
      correct_choice_position_part1 == 'left' ~ frames_hover_left,
      correct_choice_position_part1 == 'middle' ~ frames_hover_middle,
      correct_choice_position_part1 == 'right' ~ frames_hover_right
    ),
    gem_time_hover = case_when(
      correct_choice_position_part1 == 'left' ~ hover_time_left,
      correct_choice_position_part1 == "middle" ~ hover_time_middle,
      correct_choice_position_part1 == 'right' ~ hover_time_right
    ),
    other_frame_hover = case_when(
      stoneposition == "left" & correct_choice_position_part1 == "middle" ~ frames_hover_right,
      stoneposition == "left" & correct_choice_position_part1 == "right" ~ frames_hover_middle,
      stoneposition == "middle" & correct_choice_position_part1 == "right" ~ frames_hover_left,
      stoneposition == "middle" & correct_choice_position_part1 == "left" ~ frames_hover_right,
      stoneposition == "right" & correct_choice_position_part1 == "middle" ~ frames_hover_left,
      stoneposition == "right" & correct_choice_position_part1 == "left" ~ frames_hover_middle,
      TRUE ~ NA_real_
    ),
    other_time_hover = case_when(
      stoneposition == "left" & correct_choice_position_part1 == "middle" ~ hover_time_right,
      stoneposition == "left" & correct_choice_position_part1== "right" ~ hover_time_middle,
      stoneposition == "middle" & correct_choice_position_part1 == "right" ~ hover_time_left,
      stoneposition == "middle" & correct_choice_position_part1== "left" ~ hover_time_right,
      stoneposition == "right" & correct_choice_position_part1== "middle" ~ hover_time_left,
      stoneposition == "right" & correct_choice_position_part1== "left" ~ hover_time_middle,
      TRUE ~ NA_real_
    ),
    
    
  )
  




```

```{r gems, cache=F}
library(lme4)

gems_data <- phase_2_df %>%
  filter(!is_catch_trial) %>%
  inner_join(
    phase_1_df %>% 
      filter(!is_catch_trial) %>%
      mutate(chose_gems_p1 = chose_gems) %>%
      select(chose_gems_p1, subj_id, trial_num),
    by = c("subj_id", "trial_num")
  ) %>%
  filter(!chose_gems_p1, !is_consistent)

model <- glmer(
  chose_gems ~ 1 + (1 | subj_id),
  data = gems_data,
  family = binomial,
  control = glmerControl(
    optimizer = "bobyqa",)
)

summary(model)

plogis(fixef(model)[1])

```

We ran a mixed-effects logistic regression model, for the probability that participants chose gems in part 2 given that they did not choose gems in part 1 and incorrectly reconstruct in part 2. Participants choose the option with gems with a probability of `r printnum(round(plogis(fixef(model)[1]),2))`, significantly higher than chance (β = `r printnum(round(fixef(model)[1], 2))`,  
SE = `r printnum(round(summary(model)$coefficients[1, "Std. Error"], 2))`, 
*z* = `r printnum(round(summary(model)$coefficients[1, "z value"], 2))`, 
*p* `r printp(summary(model)$coefficients[1, "Pr(>|z|)"])`)


```{r chose_onlystones, cache=F}

chose_onlystones_df_exp1 <- read.csv(
  "/Users/sawakeishiro/Desktop/code/reconstruction/analysis/exp1/chose_onlystones_df_exp1.csv"
)


chose_onlystones_df_exp2 <- phase_2_df %>%
  mutate(
    chose_gems_p2 = chose_gems,
    chose_onlystones_p2 = chose_onlystones
  ) %>%
  inner_join(
    phase_1_df %>%
      mutate(chose_onlystones_p1 = chose_onlystones) %>%
      select(subj_id, trial_num, chose_onlystones_p1)
  ) %>%
  select(chose_onlystones_p1, chose_onlystones_p2, chose_gems_p2, subj_id, trial_num)

chose_onlystones_df_combined <- bind_rows(chose_onlystones_df_exp2, chose_onlystones_df_exp1) %>%
  filter(chose_onlystones_p1, !chose_onlystones_p2)



model <- glmer(
  chose_gems_p2 ~ 1 + (1 | subj_id),
  data = chose_onlystones_df_combined,
  family = binomial,
  control = glmerControl(
    optimizer = "bobyqa",)
)

summary(model)

plogis(fixef(model)[1])
  

```

Running a mixed-effects logistic regression model, combining data from both Exp. 1 and 2, we find that participants are more likely to choose the option with gems given that they chose the stones only option in part 1 and were incorrect in part 2. The probability of choosing the option with gems is `r printnum(round(plogis(fixef(model)[1]), 2))`. This effect is marginally significant  (β = `r printnum(round(fixef(model)[1], 2))`,  
SE = `r printnum(round(summary(model)$coefficients[1, "Std. Error"], 2))`, 
*z* = `r printnum(round(summary(model)$coefficients[1, "z value"], 2))`, 
*p* `r printp(summary(model)$coefficients[1, "Pr(>|z|)"])`)



```{r standardise_data, cache=F}

keys <- c("subj_id", "trial_num")

names1 <- names(phase_1_df)
names2 <- names(phase_2_df)

common_names <- intersect(names1, names2)
common_names
common_nonkey <- setdiff(common_names, keys)
common_nonkey

two_parts_df <- phase_1_df %>%
  inner_join(phase_2_df %>%
               mutate(
                 chose_gems_p2 = chose_gems,
                 chose_onlystones_p2 = chose_onlystones, 
                 chose_other_p2 = chose_other,
                 response_position_p2 = response_position,
                 correct_choice_position_p2 = correct_choice_position,
                 rt_p2 = rt
               ) %>%
               select(-c(left_word, 
                         left_number,
                         middle_word,
                         middle_number,
                         right_word,
                         right_number,
                         chose_gems,
                         chose_onlystones,
                         chose_other,
                         response_position,
                         correct_choice_position,
                         is_catch_trial,
                         is_show_trial,
                         stoneposition,
                         emeraldposition,
                         rt, 
                         mentioned_rule)),
             by = c("subj_id", "trial_num"))

write.csv(
  two_parts_df,
  "two_parts_df_exp2.csv",
  row.names = FALSE
)


```


\newpage

# References

::: {#refs custom-style="Bibliography"}
:::
