---
title             : "The title"
shorttitle        : "Title"

author: 
  - name          : "First Author"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "Postal address"
    email         : "my@email.com"
    role: # Contributorship roles (e.g., CRediT, https://credit.niso.org/)
      - "Conceptualization"
      - "Writing - Original Draft Preparation"
      - "Writing - Review & Editing"
  - name          : "Ernst-August Doelle"
    affiliation   : "1,2"
    role:
      - "Writing - Review & Editing"
      - "Supervision"

affiliation:
  - id            : "1"
    institution   : "Wilhelm-Wundt-University"
  - id            : "2"
    institution   : "Konstanz Business School"

authornote: |
  Add complete departmental affiliations for each author here. Each new line herein must be indented, like this line.

  Enter author note here.

abstract: |
  One or two sentences providing a **basic introduction** to the field,  comprehensible to a scientist in any discipline.
  Two to three sentences of **more detailed background**, comprehensible  to scientists in related disciplines.
  One sentence clearly stating the **general problem** being addressed by  this particular study.
  One sentence summarizing the main result (with the words "**here we show**" or their equivalent).
  Two or three sentences explaining what the **main result** reveals in direct comparison to what was thought to be the case previously, or how the  main result adds to previous knowledge.
  One or two sentences to put the results into a more **general context**.
  Two or three sentences to provide a **broader perspective**, readily comprehensible to a scientist in any discipline.
  
  <!-- https://tinyurl.com/ybremelq -->
  
keywords          : "keywords"
wordcount         : "X"

bibliography      : "r-references.bib"

floatsintext      : no
linenumbers       : yes
draft             : no
mask              : no

figurelist        : no
tablelist         : no
footnotelist      : no

classoption       : "man"
output            : papaja::apa6_word
---

```{r setup, include = FALSE}
library("papaja")
r_refs("r-references.bib")

library('groundhog')
groundhog.library(
  c(
    'papaja',
    'tidyverse'
  ), "2025-06-01"
)

r_refs("r-references.bib")

knitr::opts_chunk$set(warning=F,echo=F,message=F,cache=T)
```

```{r analysis-preferences}
# Seed for random number generation
set.seed(42)
knitr::opts_chunk$set(cache.extra = knitr::rand_seed)
```



# Methods
We report how we determined our sample size, all data exclusions (if any), all manipulations, and all measures in the study. <!-- 21-word solution (Simmons, Nelson & Simonsohn, 2012; retrieved from http://ssrn.com/abstract=2160588) -->

## Participants

## Material

## Procedure

## Preregistered data analysis 
We used `r cite_r("r-references.bib")` for all our analyses.
```{r loadformat, include = FALSE, cache=F}
library(tidyverse)
library(stringr)
library(effsize)

file_list <- '/Users/sawakeishiro/Desktop/code/reconstruction/data/Experiment3/experiment3_data.txt'

raw_df <- map_dfr(file_list, function(file) {

  lines <- readLines(file)
  starts <- which(grepl('^"?\\s*success', lines))
  starts <- c(starts, length(lines) + 1)
  chunks <- map2(starts[-length(starts)], starts[-1] - 1, ~ lines[.x:.y])
  map_dfr(chunks, ~ read_csv(
      paste(.x, collapse = "\n"),
      show_col_types = FALSE,
      col_types = cols(.default = "c") 
    ))  %>%
    mutate(
      subj_id = PROLIFIC_PID,
      is_catch_trial = as.logical(is_catch_trial),
      is_show_trial = as.logical(is_show_trial),
      response = as.numeric(response),
      chose_gems = as.logical(chose_gems),
      chose_onlystones = as.logical(chose_onlystones),
      chose_other = as.logical(chose_other),
      left_number = as.numeric(left_number),
      middle_number = as.numeric(middle_number),
      right_number = as.numeric(right_number),
      cumulative_score = as.numeric(cumulative_score),
      is_consistent = as.logical(is_consistent),
      is_consistent_computer = as.logical(is_consistent_computer),
      trial_num = as.numeric(trial_num),
      rt = as.numeric(rt),
      choice_rt = as.numeric(choice_rt),
      reveal_rt = as.numeric(reveal_rt),
      confidence_rt = as.numeric(confidence_rt),
      comprehension_failures_part1 = as.numeric(comprehension_failures_part1),
      comprehension_failures_part2_1 = as.numeric(comprehension_failures_part2_1),
      comprehension_failures_part2_2 = as.numeric(comprehension_failures_part2_2),
      comprehension_failures_part2_3 = as.numeric(comprehension_failures_part2_3),
      comprehension_failures_part2_4 = as.numeric(comprehension_failures_part2_4),
      confidence = as.numeric(confidence),
      rule_check = as.character(rule_check),
      worker_comments = as.character(worker_comments),
      number_correct_part1 = as.numeric(number_correct_part1_slider),
      number_correct_part1_computer = as.numeric(number_correct_part1_computer_slider),
      number_correct_part2 = as.numeric(number_correct_part2_slider),
      computer_chose_gems = as.logical(computer_chose_gems),
      computer_chose_onlystones = as.logical(computer_chose_onlystones),
      computer_chose_other =as.logical(computer_chose_other)
    )
})

comments_sparse_df <- raw_df %>%
  dplyr::select(subj_id, rule_check, worker_comments, number_correct_part1, number_correct_part2) %>%
  dplyr::filter(!is.na(worker_comments) | !is.na(rule_check) | !is.na(number_correct_part1) | !is.na(number_correct_part2))

comments_clean_df <- comments_sparse_df %>%
  group_by(subj_id) %>%
  summarise(
    rule_check = first(na.omit(rule_check)),
    worker_comments = first(na.omit(worker_comments)),
    number_correct_part1 = first(na.omit(number_correct_part1)),
    number_correct_part2 = first(na.omit(number_correct_part2)),
    .groups = 'drop' 
  )

print(head(comments_clean_df))

write.csv(comments_clean_df, 'comments_exp3_clean.csv', row.names = FALSE)

# Process data for Phase 1
phase_1_df <- raw_df %>%
  filter(phase == 'part1') %>%
  dplyr::select(
    subj_id,
    trial_num,
    left_word,
    left_number,
    middle_word, 
    middle_number,
    right_word, 
    right_number, 
    chose_gems,
    chose_onlystones,
    chose_other,
    response_position,
    correct_choice_position,
    response_position,
    computer_chose_gems,
    computer_chose_onlystones,
    computer_chose_other,
    is_catch_trial, 
    is_show_trial,
    stoneposition,
    emeraldposition,
    rt,
    choice_rt,
    reveal_rt,
    comprehension_failures_part1,
    number_correct_part1,
    computer_response_position) %>%
  group_by(subj_id) %>%
  arrange(trial_num) 

write.csv(raw_df, 'raw_df.csv')
phase_2_df <- raw_df %>%
  filter(phase == 'part2') %>%
  dplyr::select(
    subj_id, 
    trial_num,
    is_consistent,
    is_consistent_computer,
    correct_choice_position,
    confidence,
    chose_gems,
    chose_onlystones,
    chose_other,
    rt,
    choice_rt,
    confidence_rt,
    is_catch_trial,
    is_show_trial,
    stoneposition,
    emeraldposition,
    left_word,
    left_number,
    middle_word, 
    middle_number,
    right_word, 
    right_number, 
    comprehension_failures_part2_1,
    comprehension_failures_part2_2,
    comprehension_failures_part2_3,
    comprehension_failures_part2_4,
    number_correct_part2,
    response_position
  ) %>%
  group_by(subj_id) %>%
  arrange(trial_num)  %>%
  filter(!is.na(is_consistent)) 

for_merge_phase_1_df <- phase_1_df %>%
  select(subj_id, trial_num, computer_response_position, part1_response_position = response_position) 

phase_2_df <- phase_2_df %>%
  left_join(for_merge_phase_1_df, by=c("subj_id", "trial_num")) %>%
  mutate(
    computer_response_position_p1 = computer_response_position,
    part1_response_position = part1_response_position
  ) %>%
    select(!computer_response_position, !response_position)

#excluding participants

total_participants_before <- length(unique(phase_1_df$subj_id))
cat("Total number of participants before exclusion:", total_participants_before)

exclude_catch <- phase_1_df %>%
  group_by(subj_id) %>%
  summarise(
    n_failed_catch = sum(!chose_gems[is_catch_trial == TRUE], na.rm = TRUE),)%>%
  filter(n_failed_catch >= 2) %>%
  pull(subj_id)

exclude_part1_rt <- phase_1_df %>%
  group_by(subj_id) %>%
  summarise(
  first_quartile_rt = quantile(rt, probs = 0.25, na.rm = TRUE))%>%
  filter(first_quartile_rt < 100 ) %>%
  pull(subj_id)

exclude_part2_rt <- phase_2_df %>%
  group_by(subj_id) %>%
  summarise(
    first_quartile_rt = quantile(rt, probs = 0.25, na.rm = TRUE)) %>%
  filter( first_quartile_rt < 100 ) %>%
  pull(subj_id)

exclude_part2_confidence_rt <- phase_2_df %>%
  group_by(subj_id) %>%
  summarise(
    first_quantile_rt = quantile(confidence_rt, probs = 0.25, na.rm = TRUE)
  ) %>%
  filter(first_quantile_rt < 100) %>%
  pull(subj_id)
  

exclude_comprehension <- raw_df %>%
  filter(comprehension_failures_part1 >= 2 | comprehension_failures_part2_1 >= 2 |  comprehension_failures_part2_2 >= 2| comprehension_failures_part2_3 >= 2| comprehension_failures_part2_4 >= 2) %>%
  distinct(subj_id) %>%
  pull(subj_id)


participants_to_exclude <- unique(c(exclude_catch,
                                    exclude_part1_rt, 
                                    exclude_part2_rt, 
                                    exclude_part2_confidence_rt,
                                    exclude_comprehension))


cat("Number of participants to exclude:", length(participants_to_exclude), "\n")
if (length(participants_to_exclude) > 0) {
  cat("Excluding participants:", paste(participants_to_exclude, collapse=", "), "\n")
}



phase_1_df <- phase_1_df %>%
  filter(!(subj_id %in% participants_to_exclude))

phase_2_df <- phase_2_df %>%
  filter(!(subj_id %in% participants_to_exclude))

total_participants_after <- length(unique(phase_1_df$subj_id))
cat("Total number of participants after exclusion:", total_participants_after, "\n")


two_parts_df <- phase_1_df %>%
  inner_join(
    phase_2_df,
    by = c("subj_id", "trial_num"),
    suffix = c("_part1", "_part2")
  )

# rank participants based on score:
top_ids <- raw_df %>%
  dplyr::select(
    subj_id, 
    cumulative_score
) %>%
  filter(!subj_id %in% participants_to_exclude) %>%
  group_by(subj_id) %>%
  summarise(max_cumulative_score = max(cumulative_score, na.rm = TRUE)) %>%
  arrange(desc(max_cumulative_score)) %>%
  slice_head(n = 30) %>% # 20*0.3 = 6
  pull(subj_id)

# bonus participants
print(top_ids)

library(readr)

top_ids %>%
  paste0(",0.75") %>%
  write_lines("top_ids.txt")



```

```{r demographic, include = FALSE, cache=F}

library(readr)
library(dplyr)

  
file1_path <- "/Users/sawakeishiro/Desktop/code/reconstruction/data/Experiment3/prolific_demographic_export_693ff48663deaa431f1072a1.csv"
file2_path <- "/Users/sawakeishiro/Desktop/code/reconstruction/data/Experiment3/prolific_demographic_export_6941bd0557301af8bb256c00.csv"
output_dir <- "/Users/sawakeishiro/Desktop/code/reconstruction/data/Experiment3/"
output_file_path <- file.path(output_dir, "prolific_demographic_combined.csv")


tryCatch({
  data1 <- read_csv(file1_path)
  data2 <- read_csv(file2_path)

  combined_data <- bind_rows(data1, data2)

  write_csv(combined_data, output_file_path)

  print(paste("Successfully combined files and saved to:", output_file_path))

}, error = function(e) {
  print(paste("An error occurred:", e$message))
})

library(readr)
library(dplyr)

combined_prolific_path <- "/Users/sawakeishiro/Desktop/code/reconstruction/data/Experiment3/prolific_demographic_combined.csv"
output_dir <- "/Users/sawakeishiro/Desktop/code/reconstruction/data/Experiment3/"
merged_output_path <- file.path(output_dir, "merged_prolific_and_raw_data.csv")


prolific_data <- read_csv(combined_prolific_path)

merged_data <- inner_join(
  prolific_data, 
  raw_df, 
  by = c("Participant id" = "subj_id")
)

write_csv(merged_data, merged_output_path)

print(paste("Successfully merged data and saved to:", merged_output_path))
print(paste("Original prolific rows:", nrow(prolific_data)))
print(paste("Original raw_df rows:", nrow(raw_df)))
print(paste("Total merged rows:", nrow(merged_data)))



```

```{r explicit rule discoverers, include = FALSE, cache=F}
### RULE DISCOVERY 

rule_list <- c("65dca6bb88611ac902cc60fb",
               "5dcdb60be64aaa07c14065cc",
               "5e38777a87243b095c67543b",
               "61282ca83be28773ac7c421f",
               "62b20c0065e9809e66d13fa5",
               "661403131cecbe255b64ca6f",
               "66b8e8f12fa9160e4ca9f25b",
               "66cc906a1d6bea34dd2e57b0",
               "6720fbc84253185973cbe564",
               "67665f7b0329bd3e30515ad2",
               "677a8e5affbe5673b3a79e6b",
               "682b8071ffc2da7edd49f10c",
               "654a86d08714f739c9fcea21"
)

two_parts_df <- two_parts_df %>%
  mutate(mentioned_rule = subj_id %in% rule_list)

phase_1_df <- phase_1_df %>%
  mutate(mentioned_rule = subj_id %in% rule_list)

phase_2_df <- phase_2_df %>%
  mutate(mentioned_rule = subj_id %in% rule_list)

stoneposition<- raw_df %>%
  group_by(stoneposition) %>%
  summarise(n=length(unique(subj_id)))
```

### phase 1 accuracy

```{r phase_1_and_2_accuracy,  cache=F, fig.width=8}

# Bins: 1-5, 6-10, 11-15, 16-20, 21-25, 26-30, 31-35, 36-40, 41-45, 46-50

# --- Define Position Dodge ---
pd <- position_dodge(width = 0.2)
bin_labels <- c("1–5", "6–10", "11–15", "16–20", "21–25", 
                "26–30", "31–35", "36–40", "41–45", "46–50")

# Breaks are set at 0.5, 5.5, 10.5, etc. to correctly group the integer trial numbers
breaks <- c(0.5, 5.5, 10.5, 15.5, 20.5, 25.5, 30.5, 35.5, 40.5, 45.5, 50.5)

# This sequence creates vertical lines between the bins (at 1.5, 2.5, etc.)
vline_x <- seq(1.5, length(bin_labels) - 0.5, by = 1)

# --- Create binned dataframe ---
# Uses raw_df and rule_list which are available from previous chunks
df_binned <- raw_df %>%
  filter(!is_catch_trial, phase %in% c("part1", "part2")) %>%
  mutate(
    mentioned_rule = subj_id %in% rule_list,
    bin = cut(trial_num, breaks = breaks, labels = bin_labels,
              include.lowest = TRUE, right = TRUE),
    bin = factor(bin, levels = bin_labels)
  )

# --- Summarise by subj, phase, bin, and mentioned_rule ---
by_subj_bias <- df_binned %>%
  group_by(subj_id, phase, bin, mentioned_rule) %>%
  summarise(bias = mean(chose_onlystones, na.rm = TRUE), .groups = "drop")

# --- Subgroup summary (mentioned vs not) ---
group_summary_sub <- by_subj_bias %>%
  group_by(phase, bin, mentioned_rule) %>%
  summarise(
    m  = mean(bias, na.rm = TRUE),
    se = sd(bias,  na.rm = TRUE) / sqrt(n()),
    .groups = "drop"
  )

# --- Rewritten Plot ---
ggplot(group_summary_sub, 
       aes(x = bin, y = m, 
           color = mentioned_rule,  # Swapped from 'phase'
           shape = phase,           # Swapped from 'mentioned_rule'
           linetype = phase,        # Swapped from 'mentioned_rule'
           group = interaction(mentioned_rule, phase))) +
  
  # vertical and horizontal lines (from your original)
  geom_vline(xintercept = vline_x, linetype = "dashed", linewidth = 0.3, color = "grey") +
  geom_hline(yintercept = 1/3, linewidth = 0.3) +
  
  # Geoms now inherit aesthetics and use position_dodge
  geom_line(linewidth = 0.7, position = pd) +
  geom_point(position = pd, size = 2) +
  geom_errorbar(aes(ymin = m - se, ymax = m + se), 
                position = pd, width = 0.2, linetype = 'solid') +
  
  # Updated scales to match new aesthetics
  
  scale_shape_discrete(name = "Part") +
  scale_linetype_discrete(name = "Part") +
  
  scale_color_discrete(
    name = "Mentioned position",
    labels = c("TRUE" = "Mentioned", "FALSE" = "NOT mentioned")
  ) +
 

  # Updated labs
  labs(x = "Trial Bin", y = "P(stones only)",
       title = "Binned bias by rule discovery (group mean ± SE)") +
  theme_minimal()



phase_1_acc <- phase_1_df %>%
  group_by(subj_id) %>%
  filter(!is_catch_trial) %>%
  summarise(acc=mean(chose_gems))

r_phase_1_acc <- phase_1_df %>%
  group_by(subj_id) %>%
  filter(!is_catch_trial, mentioned_rule) %>%
  summarise(acc=mean(chose_gems))

nr_phase_1_acc <- phase_1_df %>%
  group_by(subj_id) %>%
  filter(!is_catch_trial, !mentioned_rule) %>%
  summarise(acc=mean(chose_gems))

phase_1_acc_slopes <- phase_1_df %>%
  group_by(subj_id) %>%
  filter(!is_catch_trial) %>%
  summarise(r=cor(chose_gems,trial_num))

r_phase_1_acc_slopes <- phase_1_df %>%
  group_by(subj_id) %>%
  filter(!is_catch_trial, mentioned_rule) %>%
  summarise(r=cor(chose_gems,trial_num))

nr_phase_1_acc_slopes <- phase_1_df %>%
  group_by(subj_id) %>%
  filter(!is_catch_trial, !mentioned_rule) %>%
  summarise(r=cor(chose_gems,trial_num))

phase_2_acc <- phase_2_df %>%
  group_by(subj_id) %>%
  filter(!is_catch_trial) %>%
  summarise(acc=mean(is_consistent, na.rm =TRUE))

r_phase_2_acc <- phase_2_df %>%
  group_by(subj_id) %>%
  filter(!is_catch_trial, mentioned_rule) %>%
  summarise(acc=mean(is_consistent, na.rm =TRUE))

nr_phase_2_acc <- phase_2_df %>%
  group_by(subj_id) %>%
  filter(!is_catch_trial, !mentioned_rule) %>%
  summarise(acc=mean(is_consistent, na.rm =TRUE))

phase_2_acc_slopes <- phase_2_df %>%
  group_by(subj_id) %>%
  filter(!is_catch_trial) %>%
  summarise(r=cor(is_consistent,trial_num, use = "complete.obs"))

r_phase_2_acc_slopes <- phase_2_df %>%
  group_by(subj_id) %>%
  filter(!is_catch_trial, mentioned_rule) %>%
  summarise(r=cor(is_consistent,trial_num, use = "complete.obs"))

nr_phase_2_acc_slopes <- phase_2_df %>%
  group_by(subj_id) %>%
  filter(!is_catch_trial, !mentioned_rule) %>%
  summarise(r=cor(is_consistent,trial_num, use = "complete.obs"))

```


In part 1, mean proportion correct (in non-catch trials) was `r phase_1_acc %>% pull(acc) %>%mean()%>%printnum()` (median=`r phase_1_acc %>% pull(acc) %>%median()%>%printnum()`) and was not significantly higher than chance (1/3; `r apa_print(t.test(phase_1_acc %>% pull(acc),mu=1/3))$statistic`). Participants did not significantly improve as a function of trial number (a t-test on subject-specific learning slopes against 0: `r apa_print(t.test(phase_1_acc_slopes %>% pull(r)))$statistic`). 

The mean accuracy for participants who mentioned the rule was `r r_phase_1_acc %>% pull(acc) %>%mean()%>%printnum()` (median=`r r_phase_1_acc %>% pull(acc) %>%median()%>%printnum()`) and was significantly higher than chance (1/3; `r apa_print(t.test(r_phase_1_acc %>% pull(acc),mu=1/3))$statistic`). Participants did not significantly improve as a function of trial number (a t-test on subject-specific learning slopes against 0: `r apa_print(t.test(r_phase_1_acc_slopes %>% pull(r)))$statistic`). 

The mean accuracy for participants who did not mention the rule was r `r nr_phase_1_acc %>% pull(acc) %>%mean()%>%printnum()` (median=`r nr_phase_1_acc %>% pull(acc) %>%median()%>%printnum()`) and was not significantly higher than chance (1/3; `r apa_print(t.test(nr_phase_1_acc %>% pull(acc),mu=1/3))$statistic`). Participants did not improve as a function of trial number (a t-test on subject-specific learning slopes against 0: `r apa_print(t.test(nr_phase_1_acc_slopes %>% pull(r)))$statistic`). 



### Hypothesis 1 (RULE LEARNING AVOIDANCE)

```{r phase1avoidance, include = FALSE, cache=F}
phase_1_pstones <- phase_1_df %>%
  group_by(subj_id) %>%
  filter(!is_catch_trial) %>%
  summarise(pstones=mean(chose_onlystones,na.rm = TRUE))

r_phase_1_pstones <- phase_1_df %>%
  group_by(subj_id) %>%
  filter(!is_catch_trial, mentioned_rule) %>%
  summarise(pstones=mean(chose_onlystones,na.rm = TRUE))

nr_phase_1_pstones <- phase_1_df %>%
  group_by(subj_id) %>%
  filter(!is_catch_trial, !mentioned_rule) %>%
  summarise(pstones=mean(chose_onlystones,na.rm = TRUE))


```

In part 1, participants chose the side that always had stones (in non-catch trials)  on `r phase_1_pstones %>% pull(pstones) %>%mean()%>%printnum()` (median=`r phase_1_pstones %>% pull(pstones) %>%median()%>%printnum()`) of all trials, not significantly below chance (1/3; `r apa_print(t.test(phase_1_pstones %>% pull(pstones),mu=1/3))$statistic`).

Participants who mentioned the rule chose the side that always had stones on `r r_phase_1_pstones %>% pull(pstones) %>%mean()%>%printnum()` (median=`r r_phase_1_pstones %>% pull(pstones) %>%median()%>%printnum()`) of all trials, significantly below chance (1/3; `r apa_print(t.test(r_phase_1_pstones %>% pull(pstones),mu=1/3))$statistic`). This is in contrast to participants who did not mention the rule who chose the side that always had stones on `r nr_phase_1_pstones %>% pull(pstones) %>%mean()%>%printnum()` (median=`r nr_phase_1_pstones %>% pull(pstones) %>%median()%>%printnum()`) of all trials, performing at chance (1/3; `r apa_print(t.test(nr_phase_1_pstones %>% pull(pstones),mu=1/3))$statistic`). The difference in means between the two groups are significant (`r apa_print(t.test(r_phase_1_pstones %>% pull(pstones),nr_phase_1_pstones %>% pull(pstones)))$statistic`)

### Hypothesis 2 (RULE LEARNING: SLOPES)

```{r phase1slopes, include = FALSE, cache=F}
library(broom)
pstones_slopes <- two_parts_df %>%
  group_by(subj_id) %>%
  filter(!is_catch_trial_part1) %>% 
  do(tidy(glm(
    chose_onlystones_part1 ~ trial_num, 
    family = binomial, 
    data = .
  ))) %>%
  filter(term == "trial_num") %>%
  select(subj_id, slope = estimate)

r_pstones_slopes <- two_parts_df %>%
  group_by(subj_id) %>%
  filter(!is_catch_trial_part1, mentioned_rule) %>% 
  do(tidy(glm(
    chose_onlystones_part1 ~ trial_num, 
    family = binomial, 
    data = .
  ))) %>%
  filter(term == "trial_num") %>%
  select(subj_id, slope = estimate)

nr_pstones_slopes <- two_parts_df %>%
  group_by(subj_id) %>%
  filter(!is_catch_trial_part1, !mentioned_rule) %>% 
  do(tidy(glm(
    chose_onlystones_part1 ~ trial_num, 
    family = binomial, 
    data = .
  ))) %>%
  filter(term == "trial_num") %>%
  select(subj_id, slope = estimate)


```

Participants did not improve as a function of trial number (a t-test on subject-specific learning slopes from a logistic regression against 0: `r apa_print(t.test(pstones_slopes %>% pull(slope)))$statistic`).This was true for participants who explicitly learnt the rule (`r apa_print(t.test(r_pstones_slopes %>% pull(slope)))$statistic`). For participants who did not learn the rule, the results were marginally significant, suggesting some learning as a function of trial number (`r apa_print(t.test(nr_pstones_slopes %>% pull(slope)))$statistic`). 

```{r softmax_model, cache=FALSE}

library(tidyverse)
library(mlogit)
library(broom)
library(purrr)

part1_trialtype <- phase_1_df %>%
  mutate(
    part1_type = case_when(
      chose_gems ~ 'im_correct',
      computer_chose_gems ~ 'com_correct',
      chose_gems == FALSE & computer_chose_gems == FALSE ~ 'no_correct'
    ),
    part1_com_chose = case_when(
      computer_chose_gems ~ 'gems',
      computer_chose_onlystones ~ 'stones',
      computer_chose_other ~ 'other'
    ),
    part1_chose_gems = chose_gems | computer_chose_gems,
    i_chose_gems = chose_gems,
    part1_correct_choice_position = correct_choice_position
  ) %>%
  select(subj_id, trial_num, part1_type, part1_com_chose, part1_chose_gems, 
         computer_chose_gems, computer_chose_other, computer_chose_onlystones, 
         i_chose_gems, part1_correct_choice_position)

combined_trialtype <- phase_2_df %>%
  left_join(part1_trialtype, by = c("subj_id", "trial_num")) %>%
  mutate(
    part2_type = case_when(
      is_consistent ~ 'my_choice',
      is_consistent_computer ~ 'com_choice',
      !is_consistent & !is_consistent_computer ~ 'no_choice'
    ),
    rewarded = part1_chose_gems,
    trial_type = case_when(
      part1_chose_gems & is_consistent ~ "im_correct_my_choice",
      part1_chose_gems & is_consistent_computer ~ "im_correct_com_choice",
      part1_chose_gems & part2_type == "no_choice" ~ "im_correct_no_choice",
      computer_chose_gems & is_consistent ~ "com_correct_my_choice",
      computer_chose_gems & is_consistent_computer ~ "com_correct_com_choice",
      computer_chose_gems & part2_type == "no_choice" ~ "com_correct_no_choice",
      part1_type == "no_correct" & is_consistent ~ "no_correct_my_choice",
      part1_type == "no_correct" & is_consistent_computer ~ "no_correct_com_choice",
      part1_type == "no_correct" & part2_type == "no_choice" ~ "no_correct_no_choice"
    )
  )


combined_trialtype <- combined_trialtype %>%
  rowwise() %>%
  mutate(
    part1_subject_pos = as.character(part1_response_position),
    part1_computer_pos = case_when(
      computer_chose_gems          ~ part1_correct_choice_position,
      computer_chose_onlystones    ~ stoneposition,
      computer_chose_other         ~ {
        others <- setdiff(c("left","middle","right"),
                          c(part1_correct_choice_position, stoneposition))
        if (length(others) > 0) others[1] else NA_character_
      },
      TRUE ~ NA_character_
    ),
    part1_none_pos = {
      used <- c(part1_subject_pos, part1_computer_pos)
      leftover <- setdiff(c("left","middle","right"), used)
      if (length(leftover) > 0) leftover[1] else NA_character_
    }
  ) %>%
  ungroup()

spatial_long <- combined_trialtype %>%
  filter(!is_catch_trial) %>%
  mutate(key = 1) %>%
  crossing(alt = c("left", "middle", "right")) %>%
  select(-key) %>%
  rowwise() %>%
  mutate(
    part2_response_pos = case_when(
      part2_type == "my_choice"  ~ part1_subject_pos,
      part2_type == "com_choice" ~ part1_computer_pos,
      part2_type == "no_choice"  ~ part1_none_pos,
      TRUE ~ NA_character_
    ),
    
    choice = as.integer(part2_response_pos == alt),
    
    value = case_when(
      alt == "left"   ~ left_number,
      alt == "middle" ~ middle_number,
      alt == "right"  ~ right_number
    ),
    
    chosen_by_me = as.integer(part1_subject_pos == alt),
    
    chosen_by_computer = as.integer(part1_computer_pos == alt),

    gems = as.integer(part1_correct_choice_position == alt),

    rewarded = as.integer(part1_chose_gems == 1 & part1_correct_choice_position == alt),

    i_revealed_gems = as.integer(i_chose_gems == 1 & part1_correct_choice_position == alt),

    part1_my_choice  = as.integer(part1_subject_pos == alt),
    
    part1_com_choice = as.integer(part1_computer_pos == alt),

    onlystones = as.integer(stoneposition == alt),
    
    my_reward = as.integer(part1_my_choice == 1 & (part1_correct_choice_position == alt)),
    
    com_reward = as.integer(part1_com_choice == 1 & (part1_correct_choice_position == alt)),
    
    stone_trial = trial_num * onlystones,
    
    chid = paste0(subj_id, "_", trial_num)
  ) %>%
  ungroup() %>%
  mutate(
    value = as.numeric(value),
    choice = as.integer(choice),
    chosen_by_me = as.integer(chosen_by_me),
    gems = as.integer(gems),
    rewarded = as.integer(rewarded),
    i_revealed_gems = as.integer(i_revealed_gems),
    trial_num = as.numeric(trial_num)
  ) %>%
  select(subj_id, trial_num, chid, alt, choice, value,
         chosen_by_me, chosen_by_computer, gems, rewarded, i_revealed_gems,
         part1_my_choice, part1_com_choice, part1_subject_pos, part1_computer_pos,
         stone_trial, stoneposition)


fit_subject_mlogit <- function(df_sub) {
  subj <- unique(df_sub$subj_id)
  stone_val <- unique(df_sub$stoneposition)[1]
  n_chid <- df_sub %>% distinct(chid) %>% nrow()

  req_cols <- c("alt", "choice", "chid", "chosen_by_me", "gems", "rewarded", "i_revealed_gems")

  if (!all(req_cols %in% names(df_sub))) {
    message("Missing required cols for subj ", subj, ": ", paste(setdiff(req_cols, names(df_sub)), collapse = ", "))
    return(NULL)
  }

  df_sub <- df_sub %>% mutate(alt = factor(alt, levels = c("left","middle","right")))

  mdf <- tryCatch({
    mlogit.data(df_sub, choice = "choice", shape = "long",
                chid.var = "chid", alt.var = "alt", id.var = "subj_id")
  }, error = function(e) NULL)

  if (is.null(mdf)) return(NULL)

  fmla <- choice ~ chosen_by_me + chosen_by_computer + gems + rewarded + i_revealed_gems | 0

  model <- tryCatch({
    mlogit(fmla, data = mdf)
  }, error = function(e) {
    message("mlogit failed for subj ", subj, ": ", e$message)
    return(NULL)
  })

  if (is.null(model)) return(NULL)

  coefs <- tryCatch({
    est <- coef(model)
    vc  <- vcov(model)
    se <- sqrt(diag(vc))[names(est)]

    tibble(
      subj_id = subj,
      term = names(est),
      estimate = as.numeric(est),
      std.error = as.numeric(se),
      t_stat = estimate / std.error,
      p_val = 2 * (1 - pnorm(abs(t_stat))),
      n_chid = n_chid,
      stoneposition = stone_val
    )
  }, error = function(e) {
    tdf <- broom::tidy(model)
    if (!is.null(tdf) && all(c("term","estimate","std.error") %in% names(tdf))) {
      return(tdf %>%
               transmute(subj_id = subj,
                         term = term,
                         estimate = estimate,
                         std.error = std.error,
                         t_stat = estimate / std.error,
                         p_val = 2 * (1 - pnorm(abs(t_stat))),
                         stoneposition = stone_val,
                         n_chid = n_chid))
    } else {
      return(tibble(
        subj_id = subj,
        term = "ERROR_extract_coefs",
        estimate = NA_real_,
        std.error = NA_real_,
        t_stat = NA_real_,
        p_val = NA_real_,
        stoneposition = stone_val,
        n_chid = n_chid,
        notes = paste("coef extraction failed:", conditionMessage(e))
      ))
    }
  })

  return(coefs)
}

per_subj_results <- spatial_long %>%
  group_by(subj_id) %>%
  group_split() %>%
  set_names(map_chr(., ~ unique(.$subj_id))) %>%
  map_dfr(~ fit_subject_mlogit(.x), .id = "subj_id_try")

all_subj <- unique(spatial_long$subj_id)
fitted_subj <- unique(per_subj_results$subj_id)
failed_subjects <- setdiff(all_subj, fitted_subj)

if (length(failed_subjects)>0) {
  message("Failed subjects: ", paste(failed_subjects, collapse=", "))
}

if ("stoneposition" %in% names(per_subj_results)) {
  per_subj_results <- per_subj_results %>% mutate(stoneposition = as.character(stoneposition))
}

per_subj_betas <- per_subj_results %>%
  select(subj_id, term, estimate, std.error, t_stat, p_val, n_chid, stoneposition)

group_level_betas <- per_subj_betas %>%
  filter(!grepl("^ERROR", term)) %>%
  group_by(term) %>%
  summarise(
    mean_beta = mean(estimate, na.rm = TRUE),
    median_beta = median(estimate, na.rm = TRUE),
    sd_beta   = sd(estimate, na.rm = TRUE),
    n         = sum(!is.na(estimate)),
    ttest     = list(if (n>1) t.test(estimate, mu = 0) else NULL),
    .groups = "drop"
  ) %>%
  mutate(
    t_stat  = if_else(n>1, map_dbl(ttest, ~ .x$statistic), NA_real_),
    p_value = if_else(n>1, map_dbl(ttest, ~ .x$p.value), NA_real_),
    conf_low  = if_else(n>1, map_dbl(ttest, ~ .x$conf.int[1]), NA_real_),
    conf_high = if_else(n>1, map_dbl(ttest, ~ .x$conf.int[2]), NA_real_),
    significance = case_when(
      p_value < 0.001 ~ "***",
      p_value < 0.01  ~ "**",
      p_value < 0.05  ~ "*",
      TRUE            ~ "ns"
    )
  ) %>%
  select(-ttest)

```

### Hypothesis 3 (RECONSTRUCTION ACCURACY)

We will test the null hypothesis that participants’ accuracy in reconstructing their choices from part 1 is at chance (1/3). This will be tested by taking the mean beta coefficient for each participant for the variable chosen by me and performing a one-sample t-test.

In part 2, mean proportion correct (in non-catch trials) was `r phase_2_acc %>% pull(acc) %>%mean()%>%printnum()` (median=`r phase_2_acc %>% pull(acc) %>%median()%>%printnum()`).Mean accuracy for participants who mentioned rule was `r r_phase_2_acc %>% pull(acc) %>%mean()%>%printnum()` (median=`r r_phase_2_acc %>% pull(acc) %>%median()%>%printnum()`) and mean accuracy for participants who did not mention the rule was `r nr_phase_2_acc %>% pull(acc) %>%mean()%>%printnum()` (median=`r nr_phase_2_acc %>% pull(acc) %>%median()%>%printnum()`).

The mean beta coefficient for *chosen by me* was `r group_level_betas %>% filter(term == "chosen_by_me") %>% pull(mean_beta)` (median = `r group_level_betas %>% filter(term == "chosen_by_me") %>% pull(median_beta)`). The results from the one-sample t-test are significant `r apa_print(per_subj_betas %>% filter(term == "chosen_by_me") %>% pull(estimate) %>% t.test(mu = 0))$statistic`

### Hypothesis 4 (RECONSTRUCTION BIAS): 

We will test the null hypothesis that participants were equally likely to choose a box regardless of whether they (either participant or computer) obtained a reward or not. This will be tested by taking the mean beta coefficient for each participant for the variable gems and performing a one-sample t-test.


The mean beta coefficient for *gems* was `r group_level_betas %>% filter(term == "gems") %>% pull(mean_beta)` (median = `r group_level_betas %>% filter(term == "gems") %>% pull(median_beta)`). The results from the one-sample t-test are not significant `r apa_print(per_subj_betas %>% filter(term == "gems") %>% pull(estimate) %>% t.test(mu = 0))$statistic`. 

### Hypothesis 5 (EFFECT OF REWARD)

We will test the null hypothesis that participants were equally likely to choose the box that contained gems regardless of whether the gems were awarded or not. This will be tested by taking the mean beta coefficient for each participant for the variable rewarded and performing a one-sample t-test.

The mean beta coefficient for *rewarded* was `r group_level_betas %>% filter(term == "rewarded") %>% pull(mean_beta)` (median = `r group_level_betas %>% filter(term == "rewarded") %>% pull(median_beta)`). The results from the one-sample t-test are not significant `r apa_print(per_subj_betas %>% filter(term == "rewarded") %>% pull(estimate) %>% t.test(mu = 0))$statistic`. 

### Hypothesis 6 (EFFECT OF ACCURACY)

We will test the null hypothesis that participants’ accuracy in part 2 is unaffected by their accuracy in part 1 controlling for reward. This will be tested by taking the mean beta coefficient for each participant for the variable “I revealed gems” and performing a one-sample t-test.

The mean beta coefficient for *I revealed gems* was `r group_level_betas %>% filter(term == "i_revealed_gems") %>% pull(mean_beta)` (median = `r group_level_betas %>% filter(term == "i_revealed_gems") %>% pull(median_beta)`). The results from the one-sample t-test are significant `r apa_print(per_subj_betas %>% filter(term == "i_revealed_gems") %>% pull(estimate) %>% t.test(mu = 0))$statistic`. 

### Hypothesis 7: (POSITION BIAS IN RECONSTRUCTION ERRORS)

We will test the null hypothesis that participants’remembered decisions are not biased for or against the stones-only location. For this, we will focus on trials in which the original decision revealed gems, but they made an error in reconstruction. We will compare the probability of choosing the stones-only position against 50%. Since different participants will have different number of trials for this analysis, this will be tested using a mixed-effects logistic regression (chose_stones_only∼1+(1|subj_id), run with lme4’s glmer, assuming a binomial variable and using a bobyqa optimizer).

```{r hypothesis_6, include = TRUE, cache=F}

part1_choices <- phase_1_df %>%
  mutate(
    part1_choice = case_when(
      chose_gems ~ "gem",
      chose_onlystones ~ "stone",
      chose_other ~ "other"
    ),
    stone_number = case_when(
      stoneposition == 'right' ~ right_number,
      stoneposition == "middle" ~ middle_number,
      stoneposition == "left" ~ left_number
    ),
    other_number = case_when(
      correct_choice_position == "right" & stoneposition == "middle" ~ left_number,
      correct_choice_position == "right" & stoneposition == "left" ~ middle_number,
      correct_choice_position == "middle" & stoneposition == "right" ~ left_number,
      correct_choice_position == "middle" & stoneposition == "left" ~ right_number, 
      correct_choice_position == "left" & stoneposition == "right" ~ middle_number,
      correct_choice_position == "left" & stoneposition == "middle" ~ right_number
    ),
    stone_higher_than_other = case_when(
      stone_number > other_number ~ TRUE,
      stone_number <= other_number ~ FALSE,
    )
  ) %>%
  select(subj_id, trial_num, part1_choice, is_catch_trial, mentioned_rule, stone_number, other_number, stone_higher_than_other)

part2_choices <- phase_2_df %>%
  mutate(
    part2_choice = case_when(
      chose_gems ~ "gem",
      chose_onlystones ~ "stone",
      chose_other ~ "other"
    )
  ) %>%
  select(subj_id, trial_num, part2_choice, is_consistent, is_catch_trial, mentioned_rule)

choices_merged <- part2_choices %>%
  left_join(part1_choices, by = c("subj_id", "trial_num", "is_catch_trial", "mentioned_rule")) %>%
  mutate(
    chose_onlystones_part2 = (part2_choice == 'stone')
  )

never_stone_ids <- choices_merged %>%
  group_by(subj_id) %>%
  summarise(ever_stone = any(part2_choice == "stone", na.rm = TRUE)) %>%
  filter(!ever_stone) %>%
  pull(subj_id)

choices_merged <- choices_merged %>%
  filter(!subj_id %in% never_stone_ids)

is_not_consistent_chosegems_part1 <- choices_merged %>%
  filter(!is_catch_trial,
         part1_choice == 'gem',
         !is_consistent)

r_is_not_consistent_chosegems_part1 <- choices_merged %>%
  filter(!is_catch_trial,
         part1_choice == 'gem',
         !is_consistent,
         mentioned_rule)

nr_is_not_consistent_chosegems_part1 <- choices_merged %>%
  filter(!is_catch_trial,
         part1_choice == 'gem',
         !is_consistent,
         !mentioned_rule)



library(lme4)


h6_model <- glmer(chose_onlystones_part2  ~ 1 + (1 | subj_id),
                  data = is_not_consistent_chosegems_part1,
                  family = binomial,
                  control = glmerControl(optimizer = "bobyqa"))

r_h6_model <- glmer(chose_onlystones_part2 ~ 1 + (1 | subj_id),
                  data = r_is_not_consistent_chosegems_part1,
                  family = binomial,
                  control = glmerControl(optimizer = "bobyqa"))

nr_h6_model <- glmer(chose_onlystones_part2 ~ 1 + (1 | subj_id),
                  data = nr_is_not_consistent_chosegems_part1,
                  family = binomial,
                  control = glmerControl(optimizer = "bobyqa"))

# controlling for trial number
h6_model_trial <-glmer(chose_onlystones_part2 ~ trial_num + (1 + trial_num | subj_id),
                  data = is_not_consistent_chosegems_part1,
                  family = binomial,
                  control = glmerControl(optimizer = "bobyqa"))

r_h6_model_trial <- glmer(chose_onlystones_part2 ~ trial_num + (1 + trial_num | subj_id),
                  data = r_is_not_consistent_chosegems_part1,
                  family = binomial,
                  control = glmerControl(optimizer = "bobyqa"))

nr_h6_model_trial <- glmer(chose_onlystones_part2 ~ trial_num + (1 + trial_num | subj_id),
                  data = nr_is_not_consistent_chosegems_part1,
                  family = binomial,
                  control = glmerControl(optimizer = "bobyqa"))

```

In Part 2 trials in which the original decision revealed gems but participants made an error in reconstruction, the estimated probability of selecting the stones-only position was `r printnum(round(unname(1 / (1 + exp(-fixef(h6_model)))), 2))` (p = `r printp(summary(h6_model)$coefficients[,"Pr(>|z|)"][1])`).

For participants who mentioned the rule the estimated probability of selecting the stones-only position was `r printnum(round(unname(1 / (1 + exp(-fixef(r_h6_model)))), 2))` (p = `r printp(summary(r_h6_model)$coefficients[,"Pr(>|z|)"][1])`).

For participants who did not mention the rule, the estimated probability of selecting the stones-only position was `r printnum(round(unname(1 / (1 + exp(-fixef(nr_h6_model)))), 2))` (p = `r printp(summary(nr_h6_model)$coefficients[,"Pr(>|z|)"][1])`).

### Hypothesis 8: (CONFIDENCE: CORRECT VERSUS INCORRECT)

We will test the null hypothesis that participants’ confidence for part 2 is equal for correct and incorrect decisions. This will be tested using a paired t-test. 

```{r h8, cache=F}


conf_mean_correct <- phase_2_df %>%
  group_by(subj_id) %>%
  filter(!is_catch_trial, is_consistent) %>%
  summarise(mean_conf = mean(confidence, na.rm = TRUE),
            median_conf = median(confidence, na.rm=TRUE))

r_conf_mean_correct <- phase_2_df %>%
  group_by(subj_id) %>%
  filter(!is_catch_trial, is_consistent, mentioned_rule) %>%
  summarise(mean_conf = mean(confidence, na.rm = TRUE),
            median_conf = median(confidence, na.rm=TRUE))

nr_conf_mean_correct <- phase_2_df %>%
  group_by(subj_id) %>%
  filter(!is_catch_trial, is_consistent, !mentioned_rule) %>%
  summarise(mean_conf = mean(confidence, na.rm = TRUE),
            median_conf = median(confidence, na.rm=TRUE))


conf_mean_incorrect <- phase_2_df %>%
  group_by(subj_id) %>%
  filter(!is_catch_trial, !is_consistent) %>%
  summarise(mean_conf = mean(confidence, na.rm=TRUE),
            median_conf = median(confidence, na.rm=TRUE))

r_conf_mean_incorrect <- phase_2_df %>%
  group_by(subj_id) %>%
  filter(!is_catch_trial, !is_consistent, mentioned_rule) %>%
  summarise(mean_conf = mean(confidence, na.rm=TRUE),
            median_conf = median(confidence, na.rm=TRUE))

nr_conf_mean_incorrect <- phase_2_df %>%
  group_by(subj_id) %>%
  filter(!is_catch_trial, !is_consistent,!mentioned_rule) %>%
  summarise(mean_conf = mean(confidence, na.rm=TRUE),
            median_conf = median(confidence, na.rm=TRUE))

# rt 

rt_mean_correct <- phase_2_df %>%
  group_by(subj_id) %>%
  filter(!is_catch_trial, is_consistent) %>%
  summarise(mean_rt = mean(choice_rt, na.rm = TRUE),
            median_rt = median(choice_rt, na.rm=TRUE))

r_rt_mean_correct <- phase_2_df %>%
  group_by(subj_id) %>%
  filter(!is_catch_trial, is_consistent, mentioned_rule) %>%
  summarise(mean_rt = mean(choice_rt, na.rm = TRUE),
            median_rt = median(choice_rt, na.rm=TRUE))

nr_rt_mean_correct <- phase_2_df %>%
  group_by(subj_id) %>%
  filter(!is_catch_trial, is_consistent, !mentioned_rule) %>%
  summarise(mean_rt = mean(choice_rt, na.rm = TRUE),
            median_rt = median(choice_rt, na.rm=TRUE))

rt_mean_incorrect <- phase_2_df %>%
  group_by(subj_id) %>%
  filter(!is_catch_trial, !is_consistent) %>%
  summarise(mean_rt = mean(choice_rt, na.rm=TRUE),
            median_rt = median(choice_rt, na.rm=TRUE))

r_rt_mean_incorrect <- phase_2_df %>%
  group_by(subj_id) %>%
  filter(!is_catch_trial, !is_consistent, mentioned_rule) %>%
  summarise(mean_rt = mean(choice_rt, na.rm=TRUE),
            median_rt = median(choice_rt, na.rm=TRUE))

nr_rt_mean_incorrect <- phase_2_df %>%
  group_by(subj_id) %>%
  filter(!is_catch_trial, !is_consistent, !mentioned_rule) %>%
  summarise(mean_rt = mean(choice_rt, na.rm=TRUE),
            median_rt = median(choice_rt, na.rm=TRUE))


conf_paired <- conf_mean_correct %>%
  rename(mean_conf_correct = mean_conf) %>%
  inner_join(
    conf_mean_incorrect %>%
      rename(mean_conf_incorrect = mean_conf),
    by = "subj_id"
  )

r_conf_paired <- r_conf_mean_correct %>%
  rename(mean_conf_correct = mean_conf) %>%
  inner_join(
    conf_mean_incorrect %>%
      rename(mean_conf_incorrect = mean_conf),
    by = "subj_id"
  )


nr_conf_paired <- nr_conf_mean_correct %>%
  rename(mean_conf_correct = mean_conf) %>%
  inner_join(
    conf_mean_incorrect %>%
      rename(mean_conf_incorrect = mean_conf),
    by = "subj_id"
  )




rt_paired <- rt_mean_correct %>%
  rename(median_rt_correct = median_rt) %>%
  inner_join(
    rt_mean_incorrect %>%
      rename(median_rt_incorrect = median_rt),
    by = "subj_id"
  )

r_rt_paired <- r_rt_mean_correct %>%
  rename(median_rt_correct = median_rt) %>%
  inner_join(
    rt_mean_incorrect %>%
      rename(median_rt_incorrect = median_rt),
    by = "subj_id"
  )

nr_rt_paired <- nr_rt_mean_correct %>%
  rename(median_rt_correct = median_rt) %>%
  inner_join(
    rt_mean_incorrect %>%
      rename(median_rt_incorrect = median_rt),
    by = "subj_id"
  )




```

In part 2, mean confidence for correct decisions was `r conf_mean_correct$mean_conf%>%mean()%>%printnum()` (median=`r conf_mean_correct$mean_conf%>%median()%>%printnum()`). The mean confidence for incorrect decisions was `r conf_mean_incorrect$mean_conf%>%mean()%>%printnum()` (median=`r conf_mean_incorrect$mean_conf%>%median()%>%printnum()`). The difference in confidence was significant `r apa_print(t.test(conf_paired$mean_conf_correct,conf_paired$mean_conf_incorrect,paired = TRUE))$statistic`

For participants who mentioned the rule, mean confidence for correct decisions was `r r_conf_mean_correct$mean_conf%>%mean()%>%printnum()` (median=`r r_conf_mean_correct$mean_conf%>%median()%>%printnum()`). The mean confidence for incorrect decisions was `r r_conf_mean_incorrect$mean_conf%>%mean()%>%printnum()` (median=`r r_conf_mean_incorrect$mean_conf%>%median()%>%printnum()`). The difference in confidence was significant `r apa_print(t.test(r_conf_paired$mean_conf_correct,r_conf_paired$mean_conf_incorrect,paired = TRUE))$statistic`

For participants who did not mention the rule, mean confidence for correct decisions was `r nr_conf_mean_correct$mean_conf%>%mean()%>%printnum()` (median=`r nr_conf_mean_correct$mean_conf%>%median()%>%printnum()`). The mean confidence for incorrect decisions was `r nr_conf_mean_incorrect$mean_conf%>%mean()%>%printnum()` (median=`r nr_conf_mean_incorrect$mean_conf%>%median()%>%printnum()`). The difference in confidence was significant `r apa_print(t.test(nr_conf_paired$mean_conf_correct,nr_conf_paired$mean_conf_incorrect,paired = TRUE))$statistic`

### Hypothesis 8: (RT: CORRECT VERSUS INCORRECT)

In part 2, mean median RT for correct decisions was `r rt_mean_correct$median_rt%>%mean()%>%printnum()` (median=`r rt_mean_correct$median_rt%>%median()%>%printnum()`). The mean median RT for incorrect decisions was `r rt_mean_incorrect$median_rt%>%mean()%>%printnum()` (median=`r rt_mean_incorrect$median_rt%>%median()%>%printnum()`). The difference in median RT was significant `r apa_print(t.test(rt_paired$median_rt_correct,rt_paired$median_rt_incorrect,paired = TRUE))$statistic`. 

For participants who mentioned the rule, mean median RT for correct decisions was `r r_rt_mean_correct$median_rt%>%mean()%>%printnum()` (median=`r r_rt_mean_correct$median_rt%>%median()%>%printnum()`). The mean median RT for incorrect decisions was `r r_rt_mean_incorrect$median_rt%>%mean()%>%printnum()` (median=`r r_rt_mean_incorrect$median_rt%>%median()%>%printnum()`). The difference in median RT was significant `r apa_print(t.test(r_rt_paired$median_rt_correct,r_rt_paired$median_rt_incorrect,paired = TRUE))$statistic`

For participants who did not mention the rule, mean median RT for correct decisions was `r nr_rt_mean_correct$median_rt%>%mean()%>%printnum()` (median=`r nr_rt_mean_correct$median_rt%>%median()%>%printnum()`). The mean median RT for incorrect decisions was `r nr_rt_mean_incorrect$median_rt%>%mean()%>%printnum()` (median=`r nr_rt_mean_incorrect$median_rt%>%median()%>%printnum()`). The difference in median RT was significant `r apa_print(t.test(nr_rt_paired$median_rt_correct,nr_rt_paired$median_rt_incorrect,paired = TRUE))$statistic`

### Hypothesis 9: (CONFIDENCE: CHOOSING GEMS IN PART 2)

We will test the null hypothesis that participants’ confidence  for part 2 (correct trials only) is equal for choosing an option that contained gems in part 1 and choosing an option that did not contain gems in part 1. This will be tested using a paired t-test.

```{r h9, cache=F}

#conf
gem_conf <- phase_2_df %>%
  group_by(subj_id) %>%
  filter(!is_catch_trial, is_consistent, chose_gems) %>%
  summarise(mean_conf=mean(confidence),
            median = median(confidence))

r_gem_conf <- phase_2_df %>%
  group_by(subj_id) %>%
  filter(!is_catch_trial, is_consistent, chose_gems, mentioned_rule) %>%
  summarise(mean_conf=mean(confidence),
            median = median(confidence))

nr_gem_conf <- phase_2_df %>%
  group_by(subj_id) %>%
  filter(!is_catch_trial, is_consistent, chose_gems, !mentioned_rule) %>%
  summarise(mean_conf=mean(confidence),
            median = median(confidence))

stone_conf <- phase_2_df %>% 
  group_by(subj_id) %>%
  filter(!is_catch_trial, is_consistent, !chose_gems) %>%
  summarise(mean_conf=mean(confidence),
            median = median(confidence))

r_stone_conf <- phase_2_df %>% 
  group_by(subj_id) %>%
  filter(!is_catch_trial, is_consistent, !chose_gems, mentioned_rule) %>%
  summarise(mean_conf=mean(confidence),
            median = median(confidence))

nr_stone_conf <- phase_2_df %>% 
  group_by(subj_id) %>%
  filter(!is_catch_trial, is_consistent, !chose_gems, !mentioned_rule) %>%
  summarise(mean_conf=mean(confidence),
            median = median(confidence))

conf_df <- gem_conf %>%
  rename(mean_conf_gem = mean_conf, median_gem = median) %>%
  inner_join(stone_conf %>% rename(mean_conf_stone = mean_conf, median_stone = median),
             by = "subj_id")

r_conf_df <- r_gem_conf %>%
  rename(mean_conf_gem = mean_conf, median_gem = median) %>%
  inner_join(r_stone_conf %>% rename(mean_conf_stone = mean_conf, median_stone = median),
             by = "subj_id")

nr_conf_df <- nr_gem_conf %>%
  rename(mean_conf_gem = mean_conf, median_gem = median) %>%
  inner_join(nr_stone_conf %>% rename(mean_conf_stone = mean_conf, median_stone = median),
             by = "subj_id")

#rt
gem_rt <- phase_2_df %>%
  group_by(subj_id) %>%
  filter(!is_catch_trial, is_consistent, chose_gems) %>%
  summarise(mean_rt=mean(choice_rt),
            median_rt = median(choice_rt))

r_gem_rt <- phase_2_df %>%
  group_by(subj_id) %>%
  filter(!is_catch_trial, is_consistent, chose_gems, mentioned_rule) %>%
  summarise(mean_rt=mean(choice_rt),
            median_rt = median(choice_rt))

nr_gem_rt <- phase_2_df %>%
  group_by(subj_id) %>%
  filter(!is_catch_trial, is_consistent, chose_gems, !mentioned_rule) %>%
  summarise(mean_rt=mean(choice_rt),
            median_rt = median(choice_rt))

stone_rt <- phase_2_df %>% 
  group_by(subj_id) %>%
  filter(!is_catch_trial, is_consistent, !chose_gems) %>%
  summarise(mean_rt =mean(choice_rt),
            median_rt = median(choice_rt))

r_stone_rt <- phase_2_df %>% 
  group_by(subj_id) %>%
  filter(!is_catch_trial, is_consistent, !chose_gems, mentioned_rule) %>%
  summarise(mean_rt =mean(choice_rt),
            median_rt = median(choice_rt))

nr_stone_rt <- phase_2_df %>% 
  group_by(subj_id) %>%
  filter(!is_catch_trial, is_consistent, !chose_gems, !mentioned_rule) %>%
  summarise(mean_rt =mean(choice_rt),
            median_rt = median(choice_rt))

rt_df <- gem_rt %>%
  rename(mean_rt_gem = mean_rt, median_rt_gem = median_rt) %>%
  inner_join(stone_rt %>% rename(mean_rt_stone = mean_rt, median_rt_stone = median_rt),
             by = "subj_id")

r_rt_df <- r_gem_rt %>%
  rename(mean_rt_gem = mean_rt, median_rt_gem = median_rt) %>%
  inner_join(r_stone_rt %>% rename(mean_rt_stone = mean_rt, median_rt_stone = median_rt),
             by = "subj_id")

nr_rt_df <- nr_gem_rt %>%
  rename(mean_rt_gem = mean_rt, median_rt_gem = median_rt) %>%
  inner_join(nr_stone_rt %>% rename(mean_rt_stone = mean_rt, median_rt_stone = median_rt),
             by = "subj_id")

```

In part 2, mean confidence when touching gems on trials in which participants were correct was `r gem_conf$mean_conf %>%mean()%>%printnum()` (median=`r gem_conf$mean_conf%>%median()%>%printnum() `). In trials which they chose stones, it was `r stone_conf$mean_conf %>%mean()%>%printnum()` (median=`r  stone_conf$mean_conf%>%median()%>%printnum() `). This difference was not significant (`r apa_print(t.test(conf_df$mean_conf_gem, conf_df$mean_conf_stone, paired=TRUE))$statistic`). 

For participants who mentioned the rule, mean confidence when touching gems on trials in which participants were correct was `r r_gem_conf$mean_conf %>%mean()%>%printnum()` (median=`r r_gem_conf$mean_conf%>%median()%>%printnum() `). In trials which they chose stones, it was `r r_stone_conf$mean_conf %>%mean()%>%printnum()` (median=`r  r_stone_conf$mean_conf%>%median()%>%printnum() `). This difference was not significant (`r apa_print(t.test(r_conf_df$mean_conf_gem, r_conf_df$mean_conf_stone, paired=TRUE))$statistic`). 

For participants who did not mention the rule, mean confidence when touching gems on trials in which participants were correct was `r nr_gem_conf$mean_conf %>%mean()%>%printnum()` (median=`r nr_gem_conf$median%>%median()%>%printnum() `). In trials which they chose stones, it was `r nr_stone_conf$mean_conf %>%mean()%>%printnum()` (median=`r  nr_stone_conf$median%>%median()%>%printnum() `). This difference was not significant (`r apa_print(t.test(nr_conf_df$mean_conf_gem, nr_conf_df$mean_conf_stone, paired=TRUE))$statistic`). 

### Hypothesis 9: (RT: CHOOSING GEMS IN PART 2)

In part 2, mean median RT when touching gems on trials in which participants were correct was `r gem_rt$median_rt %>%mean()%>%printnum()` (median=`r gem_rt$median_rt%>%median()%>%printnum() `). In trials which they chose stones, it was `r stone_rt$median_rt %>%mean()%>%printnum()` (median=`r  stone_rt$median_rt %>%median()%>%printnum() `). This difference was marginally significant (`r apa_print(t.test(rt_df$median_rt_gem, rt_df$median_rt_stone, paired=TRUE))$statistic`). 

For participants who mentioned the rule, mean median RT when touching gems on trials in which participants were correct was `r r_gem_rt$median_rt %>%mean()%>%printnum()` (median=`r r_gem_rt$median_rt%>%median()%>%printnum() `). In trials which they chose stones, it was `r r_stone_rt$median_rt %>%mean()%>%printnum()` (median=`r  r_stone_rt$median_rt %>%median()%>%printnum() `). This difference was significant (`r apa_print(t.test(r_rt_df$median_rt_gem, r_rt_df$median_rt_stone, paired=TRUE))$statistic`). 

For participants who did not mention the rule, mean median RT when touching gems on trials in which participants were correct was `r nr_gem_rt$median_rt %>%mean()%>%printnum()` (median=`r nr_gem_rt$median_rt%>%median()%>%printnum() `). In trials which they chose stones, it was `r nr_stone_rt$median_rt %>%mean()%>%printnum()` (median=`r  nr_stone_rt$median_rt %>%median()%>%printnum() `). This difference was not significant (`r apa_print(t.test(nr_rt_df$median_rt_gem, nr_rt_df$median_rt_stone, paired=TRUE))$statistic`). 

### Hypothesis 10 (CONFIDENCE: REWARDED VERSUS UNREWARDED): 

#### WE ABANDON THIS HYPOTHESIS AS IT IS IDENTICAL TO HYPOTHESIS 9

We will test the null hypothesis that participants’confidence for part 2 (correct trials only) is equal for choosing an option that was rewarded and choosing an option that was not rewarded. This will be tested using a paired t-test.

```{r h10, cache=F}



```

### Hypothesis 11 (CONFIDENCE: I REVEALED VERSUS COMPUTER REVEALED): 

We will test the null hypothesis that participants’ confidence for part 2 (correct trials only) is equal for choosing an option where they revealed gems in part 1 and choosing an option the computer revealed gems in part 1. This will be tested using a paired t-test.


```{r h11, cache=F}

part1_choices <- phase_1_df %>%
  mutate(rewarded = case_when(
    chose_gems|computer_chose_gems ~ TRUE,
    TRUE ~ FALSE
  ))%>%
  select(subj_id, trial_num, rewarded )

phase_2_merged_df <- phase_2_df %>%
    left_join(part1_choices, by = c("subj_id", "trial_num"))

#conf
conf_ttest_data <- phase_2_merged_df %>%
  filter(!is_catch_trial, is_consistent, rewarded) %>%
  group_by(subj_id, chose_gems) %>% 
  summarise(mean_conf = mean(confidence, na.rm = TRUE), .groups = "drop") %>%
  pivot_wider(
    names_from = chose_gems,
    values_from = mean_conf,
    names_prefix = "chose_gems_"
  ) %>%
  drop_na()

r_conf_ttest_data <- phase_2_merged_df %>%
  filter(!is_catch_trial, is_consistent, rewarded, mentioned_rule) %>%
  group_by(subj_id, chose_gems) %>% 
  summarise(mean_conf = mean(confidence, na.rm = TRUE), .groups = "drop") %>%
  pivot_wider(
    names_from = chose_gems,
    values_from = mean_conf,
    names_prefix = "chose_gems_"
  ) %>%
  drop_na()

nr_conf_ttest_data <- phase_2_merged_df %>%
  filter(!is_catch_trial, is_consistent, rewarded, !mentioned_rule) %>%
  group_by(subj_id, chose_gems) %>% 
  summarise(mean_conf = mean(confidence, na.rm = TRUE), .groups = "drop") %>%
  pivot_wider(
    names_from = chose_gems,
    values_from = mean_conf,
    names_prefix = "chose_gems_"
  ) %>%
  drop_na()

conf_test_result <- t.test(
  conf_ttest_data$chose_gems_TRUE, 
  conf_ttest_data$chose_gems_FALSE, 
  paired = TRUE
)

r_conf_test_result <- t.test(
  r_conf_ttest_data$chose_gems_TRUE, 
  r_conf_ttest_data$chose_gems_FALSE, 
  paired = TRUE
)

nr_conf_test_result <- t.test(
  nr_conf_ttest_data$chose_gems_TRUE, 
  nr_conf_ttest_data$chose_gems_FALSE, 
  paired = TRUE
)

#rt
rt_ttest_data <- phase_2_merged_df %>%
  filter(!is_catch_trial, is_consistent, rewarded) %>%
  group_by(subj_id, chose_gems) %>% 
  summarise(median_rt = median(choice_rt, na.rm = TRUE), .groups = "drop") %>%
  pivot_wider(
    names_from = chose_gems,
    values_from = median_rt,
    names_prefix = "chose_gems_"
  ) %>%
  drop_na()

r_rt_ttest_data <- phase_2_merged_df %>%
  filter(!is_catch_trial, is_consistent, rewarded, mentioned_rule) %>%
  group_by(subj_id, chose_gems) %>% 
  summarise(median_rt = median(choice_rt, na.rm = TRUE), .groups = "drop") %>%
  pivot_wider(
    names_from = chose_gems,
    values_from = median_rt,
    names_prefix = "chose_gems_"
  ) %>%
  drop_na()

nr_rt_ttest_data <- phase_2_merged_df %>%
  filter(!is_catch_trial, is_consistent, rewarded, !mentioned_rule) %>%
  group_by(subj_id, chose_gems) %>% 
  summarise(median_rt = median(choice_rt, na.rm = TRUE), .groups = "drop") %>%
  pivot_wider(
    names_from = chose_gems,
    values_from = median_rt,
    names_prefix = "chose_gems_"
  ) %>%
  drop_na()

rt_test_result <- t.test(
  rt_ttest_data$chose_gems_TRUE,
  rt_ttest_data$chose_gems_FALSE,
  paired=TRUE
)

r_rt_test_result <- t.test(
  r_rt_ttest_data$chose_gems_TRUE,
  r_rt_ttest_data$chose_gems_FALSE,
  paired=TRUE
)

nr_rt_test_result <- t.test(
  nr_rt_ttest_data$chose_gems_TRUE,
  nr_rt_ttest_data$chose_gems_FALSE,
  paired=TRUE
)




```

The mean confidence for trials in which participants revealed gems and were correct in part 1 is `r printnum(mean(conf_ttest_data$chose_gems_TRUE))` (median = `r printnum(median(conf_ttest_data$chose_gems_TRUE))`). Mean confidence for trials in which the computer revealed gems and participants were correct in part 1 is `r printnum(mean(conf_ttest_data$chose_gems_FALSE))` (median = `r printnum(median(conf_ttest_data$chose_gems_FALSE))`). The difference between the two means are not significant `r apa_print(conf_test_result)$statistic`. 

For participants who mentioned the rule, the mean confidence for trials in which participants revealed gems and were correct in part 1 is `r printnum(mean(r_conf_ttest_data$chose_gems_TRUE))` (median = `r printnum(median(r_conf_ttest_data$chose_gems_TRUE))`). Mean confidence for trials in which the computer revealed gems and participants were correct in part 1 is `r printnum(mean(r_conf_ttest_data$chose_gems_FALSE))` (median = `r printnum(median(r_conf_ttest_data$chose_gems_FALSE))`). The difference between the two means are not significant `r apa_print(r_conf_test_result)$statistic`. 

For participants who did not mention the rule, the mean confidence for trials in which participants revealed gems and were correct in part 1 is `r printnum(mean(nr_conf_ttest_data$chose_gems_TRUE))` (median = `r printnum(median(nr_conf_ttest_data$chose_gems_TRUE))`). Mean confidence for trials in which the computer revealed gems and participants were correct in part 1 is `r printnum(mean(nr_conf_ttest_data$chose_gems_FALSE))` (median = `r printnum(median(nr_conf_ttest_data$chose_gems_FALSE))`). The difference between the two means are not significant `r apa_print(nr_conf_test_result)$statistic`. 

### Hypothesis 11 (RT: I REVEALED VERSUS COMPUTER REVEALED): 

The mean median RT for trials in which participants revealed gems and were correct in part 1 is `r printnum(mean(rt_ttest_data$chose_gems_TRUE))` (median = `r printnum(median(rt_ttest_data$chose_gems_TRUE))`). Mean median RT for trials in which the computer revealed gems and participants were correct in part 1 is `r printnum(mean(rt_ttest_data$chose_gems_FALSE))` (median = `r printnum(median(rt_ttest_data$chose_gems_FALSE))`). The difference between the two means is not significant `r apa_print(rt_test_result)$statistic`. 

For participants who mentioned the rule, the mean median RT for trials in which participants revealed gems and were correct in part 1 is `r printnum(mean(r_rt_ttest_data$chose_gems_TRUE))` (median = `r printnum(median(r_rt_ttest_data$chose_gems_TRUE))`). Mean median RT for trials in which the computer revealed gems and participants were correct in part 1 is `r printnum(mean(r_rt_ttest_data$chose_gems_FALSE))` (median = `r printnum(median(r_rt_ttest_data$chose_gems_FALSE))`). The difference between the two means is marginally significant `r apa_print(r_rt_test_result)$statistic`.

For participants who did not mention the rule, the mean median RT for trials in which participants revealed gems and were correct in part 1 is `r printnum(mean(nr_rt_ttest_data$chose_gems_TRUE))` (median = `r printnum(median(nr_rt_ttest_data$chose_gems_TRUE))`). Mean median RT for trials in which the computer revealed gems and participants were correct in part 1 is `r printnum(mean(nr_rt_ttest_data$chose_gems_FALSE))` (median = `r printnum(median(nr_rt_ttest_data$chose_gems_FALSE))`). The difference between the two means is not significant `r apa_print(nr_rt_test_result)$statistic`.

## PLOTS

```{r proportion_figure, cache=F, fig.width=8}
hum_part1_choices <- phase_1_df %>%
  select(subj_id, trial_num, part1_chose_gems = chose_gems)

part1_trialtype <- phase_1_df %>%
  mutate(
    part1_type = case_when(
      chose_gems ~ 'im_correct',
      computer_chose_gems ~ 'com_correct',
      chose_gems == FALSE & computer_chose_gems == FALSE ~ 'no_correct'
    ),
  ) %>%
  select(subj_id, trial_num, part1_type)

combined_trialtype <- phase_2_df %>%
  left_join(part1_trialtype, by = c("subj_id", "trial_num")) %>%
  mutate(
    part2_type = case_when(
      is_consistent ~ 'my_choice',
      is_consistent_computer ~ 'com_choice',
      !is_consistent & !is_consistent_computer ~ 'no_choice'
    )
  ) 

im_correct_combined_trialtype <- combined_trialtype %>%
  group_by(subj_id) %>%
  filter(!is_catch_trial, part1_type == "im_correct") %>%
  summarise(
    mean_mychoice = mean(part2_type=='my_choice'),
    mean_comchoice = mean(part2_type=='com_choice'),
    mean_nochoice = mean(part2_type=='no_choice')
  ) %>%
  pivot_longer(
    cols = c(mean_mychoice, mean_comchoice, mean_nochoice),
    names_to = "choice_type",
    values_to = "value"
  ) %>% 
  group_by(choice_type) %>%
  summarise(
    mean_value = mean(value),
    n = n(),
    sem = sd(value, na.rm = TRUE) / sqrt(n),
    .groups = "drop"
            )


com_correct_combined_trialtype <- combined_trialtype %>%
  group_by(subj_id) %>%
  filter(!is_catch_trial, part1_type == "com_correct") %>%
  summarise(
    mean_mychoice = mean(part2_type=='my_choice'),
    mean_comchoice = mean(part2_type=='com_choice'),
    mean_nochoice = mean(part2_type=='no_choice')
  ) %>%
  pivot_longer(
    cols = c(mean_mychoice, mean_comchoice, mean_nochoice),
    names_to = "choice_type",
    values_to = "value"
  ) %>% 
  group_by(choice_type) %>%
  summarise(
    mean_value = mean(value),
    n = n(),
    sem = sd(value, na.rm = TRUE) / sqrt(n),
    .groups = "drop"
            )

no_correct_combined_trialtype <- combined_trialtype %>%
  group_by(subj_id) %>%
  filter(!is_catch_trial, part1_type == "no_correct") %>%
  summarise(
    mean_mychoice = mean(part2_type=='my_choice'),
    mean_comchoice = mean(part2_type=='com_choice'),
    mean_nochoice = mean(part2_type=='no_choice')
  ) %>%
  pivot_longer(
    cols = c(mean_mychoice, mean_comchoice, mean_nochoice),
    names_to = "choice_type",
    values_to = "value"
  ) %>% 
  group_by(choice_type) %>%
  summarise(
    mean_value = mean(value),
    n = n(),
    sem = sd(value, na.rm = TRUE) / sqrt(n),
    .groups = "drop"
            )


summarise_choice <- function(data, cond_label) {
  data %>%
    group_by(subj_id) %>%
    filter(!is_catch_trial, part1_type == cond_label) %>%
    summarise(
      mean_mychoice  = mean(part2_type == 'my_choice'),
      mean_comchoice = mean(part2_type == 'com_choice'),
      mean_nochoice  = mean(part2_type == 'no_choice')
    ) %>%
    pivot_longer(
      cols = c(mean_mychoice, mean_comchoice, mean_nochoice),
      names_to = "choice_type",
      values_to = "value"
    ) %>%
    group_by(choice_type) %>%
    summarise(
      condition  = cond_label,
      mean_value = mean(value),
      n          = n(),
      sem        = sd(value, na.rm = TRUE) / sqrt(n),
      .groups = "drop"
    )
}

im_df  <- summarise_choice(combined_trialtype, "im_correct")
com_df <- summarise_choice(combined_trialtype, "com_correct")
no_df  <- summarise_choice(combined_trialtype, "no_correct")

all_df <- bind_rows(im_df, com_df, no_df)

p <- ggplot(all_df,
            aes(x = choice_type, y = mean_value, fill = choice_type)) +
  geom_col(position = position_dodge(width = 0.9), alpha = 0.8) +
  geom_errorbar(aes(ymin = mean_value - sem,
                    ymax = mean_value + sem),
                width = 0.2, linewidth = 0.7, alpha = 0.5) +
  facet_grid(~ condition) +  
  labs(
    title = "Proportion of Choice Types in Part 2 by Part 1 Choice Types",
    x = "Chosen in Part 1 by:",
    y = "Proportion Selected in Part 2"
  ) +
  scale_x_discrete(
    labels = c(
    mean_mychoice  = "Participant",
    mean_comchoice = "Computer",
    mean_nochoice  = "Neither"
  )
  ) + 
  facet_grid(
    ~ condition,
    labeller = labeller(
      condition = c(
        im_correct  = "Participant Correct Trials",
        com_correct = "Computer Correct Trials",
        no_correct  = "Neither Correct Trials"
      )
    )
  ) +
   scale_fill_manual(
  name = "Part 1 choice:",
  values = c(
    mean_mychoice  = "#ff7f0e",   
    mean_comchoice = "#FA938E",   
    mean_nochoice  =  "#84B0FE"  
  ),
  labels = c(
    mean_mychoice  = "Participant Choice",
    mean_comchoice = "Computer Choice",
    mean_nochoice  = "Neither"
  )
)+ 
  
  theme_minimal() + 
  theme(
  axis.text.x = element_text(angle = 45, vjust = 0.8, hjust = 0.5)
) 

p

mychoice_by_condition <- combined_trialtype %>%
  filter(!is_catch_trial) %>%
  group_by(subj_id, part1_type) %>%
  summarise(
    mean_mychoice = mean(part2_type == "my_choice"),
    .groups = "drop"
  ) %>%
  filter(part1_type %in% c("com_correct", "no_correct")) %>%
  pivot_wider(
    names_from = part1_type,
    values_from = mean_mychoice
  )

t_test_result <- t.test(
  mychoice_by_condition$com_correct,
  mychoice_by_condition$no_correct,
  paired = TRUE
)




```
```{r confidence_figure, cache=F, fig.width=8}
summarise_confidence <- function(data, cond_label) {
  data %>%
    filter(!is_catch_trial, part1_type == cond_label) %>%  
    group_by(subj_id) %>%
    summarise(
      conf_my   = mean(confidence[part2_type == "my_choice"],   na.rm = TRUE),
      conf_com  = mean(confidence[part2_type == "com_choice"],  na.rm = TRUE),
      conf_none = mean(confidence[part2_type == "no_choice"],   na.rm = TRUE),
      .groups = "drop"
    ) %>%
    pivot_longer(
      cols = starts_with("conf_"),
      names_to = "choice_type",
      values_to = "value"
    ) %>%
    mutate(
      choice_type = case_when(
        choice_type == "conf_my"   ~ "mean_mychoice",
        choice_type == "conf_com"  ~ "mean_comchoice",
        choice_type == "conf_none" ~ "mean_nochoice",
        TRUE ~ choice_type
      ),
      condition = cond_label
    ) %>%
    group_by(condition, choice_type) %>%
    summarise(
      mean_value = mean(value, na.rm = TRUE),
      n = sum(!is.na(value)),                             
      sem = sd(value, na.rm = TRUE) / sqrt(n),
      .groups = "drop"
    )
}

im_df  <- summarise_confidence(combined_trialtype, "im_correct")
com_df <- summarise_confidence(combined_trialtype, "com_correct")
no_df  <- summarise_confidence(combined_trialtype, "no_correct")

all_df <- bind_rows(im_df, com_df, no_df)

p <- ggplot(all_df,
            aes(x = choice_type, y = mean_value, fill = choice_type)) +
  geom_col(position = position_dodge(width = 0.9),
           alpha = 0.8) +
  geom_errorbar(aes(ymin = mean_value - sem, ymax = mean_value + sem),
                width = 0.2, linewidth = 0.7, position = position_dodge(width = 0.9), alpha = 0.5) +
  facet_grid(~ condition,
             labeller = labeller(
               condition = c(
                 im_correct  = "Participant Correct Trials",
                 com_correct = "Computer Correct Trials",
                 no_correct  = "Neither Correct Trials"
               )
             )) +
  scale_x_discrete(
    labels = c(
      mean_mychoice  = "Participant",
      mean_comchoice = "Computer",
      mean_nochoice  = "Neither"
    )
  ) +
   scale_fill_manual(
  name = "Part 1 choice:",
  values = c(
    mean_mychoice  = "#ff7f0e",   
    mean_comchoice = "#FA938E",   
    mean_nochoice  =  "#84B0FE"  
  ),
  labels = c(
    mean_mychoice  = "Participant Choice",
    mean_comchoice = "Computer Choice",
    mean_nochoice  = "Neither"
  )
)+
  labs(
    title = "Mean Confidence by Part 2 Choice by Part 1 Choice",
    x = "Chosen in Part 1 by:",
    y = "Mean Confidence"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, vjust = 0.8, hjust = 0.5)
  )

p


```


```{r choice_rt_figure, cache=F, fig.width=8}

summarise_choice_rt <- function(data, cond_label) {
  data %>%
    filter(!is_catch_trial, part1_type == cond_label) %>%  
    group_by(subj_id) %>%
    summarise(
      rt_my   = median(choice_rt[part2_type == "my_choice"],   na.rm = TRUE),
      rt_com  = median(choice_rt[part2_type == "com_choice"],  na.rm = TRUE),
      rt_none = median(choice_rt[part2_type == "no_choice"],   na.rm = TRUE),
      .groups = "drop"
    ) %>%
    pivot_longer(
      cols = starts_with("rt_"),
      names_to = "choice_type",
      values_to = "value"
    ) %>%
    mutate(
      choice_type = case_when(
        choice_type == "rt_my"   ~ "mean_mychoice",
        choice_type == "rt_com"  ~ "mean_comchoice",
        choice_type == "rt_none" ~ "mean_nochoice",
        TRUE ~ choice_type
      ),
      condition = cond_label
    ) %>%
    group_by(condition, choice_type) %>%
    summarise(
      mean_value = mean(value, na.rm = TRUE),
      n = sum(!is.na(value)),                             
      sem = sd(value, na.rm = TRUE) / sqrt(n),
      .groups = "drop"
    )
}

im_df  <- summarise_choice_rt(combined_trialtype, "im_correct")
com_df <- summarise_choice_rt(combined_trialtype, "com_correct")
no_df  <- summarise_choice_rt(combined_trialtype, "no_correct")

all_df <- bind_rows(im_df, com_df, no_df)

p <- ggplot(all_df,
            aes(x = choice_type, y = mean_value, fill = choice_type)) +
  geom_col(position = position_dodge(width = 0.9),
           alpha = 0.8) +
  geom_errorbar(aes(ymin = mean_value - sem, ymax = mean_value + sem),
                width = 0.2, linewidth = 0.7, position = position_dodge(width = 0.9), alpha = 0.5) +
  facet_grid(~ condition,
             labeller = labeller(
               condition = c(
                 im_correct  = "Participant Correct Trials",
                 com_correct = "Computer Correct Trials",
                 no_correct  = "Neither Correct Trials"
               )
             )) +
  scale_x_discrete(
    labels = c(
      mean_mychoice  = "Participant",
      mean_comchoice = "Computer",
      mean_nochoice  = "Neither"
    )
  ) +
  scale_fill_manual(
  name = "Part 1 choice:",
  values = c(
    mean_mychoice  = "#ff7f0e",   
    mean_comchoice = "#FA938E",   
    mean_nochoice  =  "#84B0FE"  
  ),
  labels = c(
    mean_mychoice  = "Participant Choice",
    mean_comchoice = "Computer Choice",
    mean_nochoice  = "Neither"
  )
) +
  labs(
    title = "Median RT by Part 2 Choice by Part 1 Choice",
    x = "Chosen in Part 1 by:",
    y = "Median RT (ms)"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, vjust = 0.8, hjust = 0.5)
  )

p

```



```{r reconstruction_accuracy,  cache=F, fig.width=8}

part1_choices <- phase_1_df %>%
  select(subj_id, trial_num, part1_chose_gems = chose_gems)

phase_2_merged_df <- phase_2_df %>%
  left_join(part1_choices, by = c("subj_id", "trial_num"))

data_for_test <- phase_2_merged_df %>%
  filter(!is_catch_trial, !is.na(part1_chose_gems))

long_paired_data_all <- data_for_test %>%
  group_by(subj_id, part1_chose_gems) %>%
  summarise(mean_accuracy = mean(is_consistent, na.rm = TRUE), .groups = "drop") %>%
  mutate(
    phase_1_choice = factor(part1_chose_gems, labels = c("Did Not Choose Gems", "Chose Gems")),
    group = "all"
  )

r_long_paired_data <- data_for_test %>%
  filter(mentioned_rule) %>%
  group_by(subj_id, part1_chose_gems) %>%
  summarise(mean_accuracy = mean(is_consistent, na.rm = TRUE), .groups = "drop") %>%
  mutate(
    phase_1_choice = factor(part1_chose_gems, labels = c("Did Not Choose Gems", "Chose Gems")),
    group = "r_mentioned"
  )

nr_long_paired_data <- data_for_test %>%
  filter(!mentioned_rule) %>%
  group_by(subj_id, part1_chose_gems) %>%
  summarise(mean_accuracy = mean(is_consistent, na.rm = TRUE), .groups = "drop") %>%
  mutate(
    phase_1_choice = factor(part1_chose_gems, labels = c("Did Not Choose Gems", "Chose Gems")),
    group = "nr_not_mentioned"
  )

plot_data_summary_all <- long_paired_data_all %>%
  group_by(group, phase_1_choice) %>%
  summarise(
    grand_mean_accuracy = mean(mean_accuracy, na.rm = TRUE),
    n = n(),
    sem = sd(mean_accuracy, na.rm = TRUE) / sqrt(n),
    .groups = "drop"
  )

r_plot_data_summary <- r_long_paired_data %>%
  group_by(group, phase_1_choice) %>%
  summarise(
    grand_mean_accuracy = mean(mean_accuracy, na.rm = TRUE),
    n = n(),
    sem = sd(mean_accuracy, na.rm = TRUE) / sqrt(n),
    .groups = "drop"
  )

nr_plot_data_summary <- nr_long_paired_data %>%
  group_by(group, phase_1_choice) %>%
  summarise(
    grand_mean_accuracy = mean(mean_accuracy, na.rm = TRUE),
    n = n(),
    sem = sd(mean_accuracy, na.rm = TRUE) / sqrt(n),
    .groups = "drop"
  )

plot_data_summary <- bind_rows(plot_data_summary_all,
                               r_plot_data_summary,
                               nr_plot_data_summary) %>%
  mutate(group = factor(group, levels = c("all", "r_mentioned", "nr_not_mentioned")))

long_subject_all <- long_paired_data_all
long_subject_r   <- r_long_paired_data
long_subject_nr  <- nr_long_paired_data

y_limits <- c(0, 1)

build_group_plot <- function(summary_df, subject_df, plot_title, bar_color) {
  ggplot(summary_df, aes(x = phase_1_choice, y = grand_mean_accuracy)) +
    geom_line(data = subject_df,
              aes(x = phase_1_choice, y = mean_accuracy, group = subj_id),
              color = "gray70", alpha = 0.4, inherit.aes = FALSE) +
    geom_point(data = subject_df,
               aes(x = phase_1_choice, y = mean_accuracy, group = subj_id),
               color = "gray70", alpha = 0.4, inherit.aes = FALSE) +
    # bars
    geom_bar(stat = "identity", width = 0.6, fill = bar_color) +
    # error bars
    geom_errorbar(aes(ymin = grand_mean_accuracy - sem, ymax = grand_mean_accuracy + sem),
                  width = 0.2, color = "black") +
    labs(title = plot_title,
         x = "Choice in Phase 1",
         y = "Mean Phase 2 Accuracy") +
    scale_y_continuous(limits = y_limits, expand = c(0, 0)) +
    theme_minimal() +
    theme(legend.position = "none",
          plot.title = element_text(size = 11, face = "bold"))
}

# --- create each plot ---
plot_all <- build_group_plot(plot_data_summary_all, long_subject_all,
                             "All participants", "#4C72B0")

plot_r <- build_group_plot(r_plot_data_summary, long_subject_r,
                           "Position mentioned", "#55A868")

plot_nr <- build_group_plot(nr_plot_data_summary, long_subject_nr,
                            "Position NOT mentioned", "#C44E52")

print(plot_all)
print(plot_r)
print(plot_nr)

plot_data_paired <- data_for_test %>%
  group_by(subj_id, mentioned_rule, part1_chose_gems) %>%
  summarise(recon_acc = mean(is_consistent, na.rm = TRUE), .groups = "drop")

group_summary_new <- plot_data_paired %>%
  group_by(mentioned_rule, part1_chose_gems) %>%
  summarise(
    m = mean(recon_acc, na.rm = TRUE),
    se = sd(recon_acc, na.rm = TRUE) / sqrt(n()),
    n = n(),
    .groups = "drop"
  ) %>%
  mutate(p1_label = ifelse(part1_chose_gems, "Part1: correct", "Part1: incorrect"),
         p1_label = factor(p1_label, levels = c("Part1: incorrect","Part1: correct")))

plot_pairs_new <- plot_data_paired %>%
  mutate(p1_label = ifelse(part1_chose_gems, "Part1: correct", "Part1: incorrect")) %>%
  mutate(p1_label = factor(p1_label, levels = c("Part1: incorrect","Part1: correct")))

library(scales)

p <- ggplot() +
  geom_line(data = plot_pairs_new, aes(x = p1_label, y = recon_acc, group = subj_id),
            color = "grey70", alpha = 0.35, size = 0.4) +
  geom_point(data = plot_pairs_new, aes(x = p1_label, y = recon_acc),
             color = "grey40", alpha = 0.6, position = position_dodge(width = 0.4), size = 1.7) +
  geom_point(data = group_summary_new, aes(x = p1_label, y = m, color = mentioned_rule),
             position = position_dodge(width = 0.4), size = 3) +
  geom_errorbar(data = group_summary_new, aes(x = p1_label, ymin = m - se, ymax = m + se, color = mentioned_rule),
                width = 0.12, position = position_dodge(width = 0.4), size = 0.7) +
  facet_wrap(~ mentioned_rule, nrow = 1,
             labeller = as_labeller(c("FALSE" = "Did NOT mention position", "TRUE" = "Mentioned position"))) +
  scale_y_continuous(limits = c(0.25,1), breaks = seq(0,1,0.25), labels = percent_format(accuracy = 1)) +
  labs(x = NULL,
       y = "Reconstruction accuracy (Part 2)",
       title = "Reconstruction accuracy in Part 2 by Part 1 guess accuracy",
       color = "Mentioned rule") +
  theme_minimal(base_size = 12) +
  theme(legend.position = "none",
        strip.text = element_text(face = "bold"),
        plot.title = element_text(size = 12, face = "bold"),
        plot.subtitle = element_text(size = 10))


print(p)



```




```{r probability,  cache=F, fig.width=8}


h6_viz_data <- is_not_consistent_chosegems_part1 %>%
  # Create trial bins
  mutate(
    trial_bin = cut(
      trial_num,
      breaks = c(0, 10, 20, 30, 40, 50),
      labels = c("1-10", "11-20", "21-30", "31-40", "41-50")
    )
  ) %>%
  
  group_by(trial_bin, mentioned_rule, subj_id) %>%
  summarise(
    n = n(),
    prob_stones = mean(chose_onlystones_part2, na.rm = TRUE),
    se = sqrt(prob_stones * (1 - prob_stones) / n)
  ) %>%
  mutate(
    ci_lower = prob_stones - 1.96 * se,
    ci_upper = prob_stones + 1.96 * se
  ) %>%
  ungroup() %>%
  mutate(
    Group = ifelse(mentioned_rule, "Position mentioned", "Position NOT mentioned")
  ) 


library(dplyr)
library(ggplot2)

h6_subject_bin <- is_not_consistent_chosegems_part1 %>%
  mutate(
    trial_bin = cut(
      trial_num,
      breaks = c(0, 10, 20, 30, 40, 50),
      labels = c("1-10", "11-20", "21-30", "31-40", "41-50"),
      right = TRUE
    )
  ) %>%
  group_by(subj_id, trial_bin, mentioned_rule) %>%
  summarise(
    prob_stones = mean(chose_onlystones_part2, na.rm = TRUE),
    n_trials = n(),
    .groups = "drop"
  ) %>%
  mutate(
    Group = ifelse(mentioned_rule, "Position mentioned", "Position NOT mentioned"),
    trial_bin = factor(trial_bin, levels = c("1-10","11-20","21-30","31-40","41-50"))
  )

summary_df <- h6_subject_bin %>%
  group_by(trial_bin, Group) %>%
  summarise(
    mean_prob = mean(prob_stones, na.rm = TRUE),
    se = sd(prob_stones, na.rm = TRUE) / sqrt(n()),
    n_subj = n(),
    .groups = "drop"
  )


pd <- position_dodge(width = 0.75)

ggplot(h6_subject_bin, aes(x = trial_bin, y = prob_stones, fill = Group, group = Group)) +
  geom_jitter(aes(color = Group), 
              position = position_jitterdodge(jitter.width = 0.12, dodge.width = 0.75), 
              size = 1.6, alpha = 0.6, show.legend = FALSE) +
  geom_errorbar(data = summary_df,
                aes(x = trial_bin, y = mean_prob, ymin = mean_prob - se, ymax = mean_prob + se, group = Group),
                width = 0.18,
                position = pd,
                color = "black",
                linewidth = 0.8) +
  geom_line(data = summary_df,
            aes(x = trial_bin, y = mean_prob, color = Group, group = Group),
            position = pd,
            linewidth = 0.9,
            alpha = 0.8) +
  geom_point(data = summary_df,
             aes(x = trial_bin, y = mean_prob, group = Group),
             position = pd,
             shape = 23, size = 3, fill = "white", color = "black") +
  geom_hline(yintercept = 0.5, linetype = "dashed", color = "black", linewidth = 0.7) +
  labs(
    title = "Exp. 2: Subject-level probability of choosing stones by trial bin",
    x = "Trial bin",
    y = "Probability of choosing stones (per subject)",
    fill = "Group"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    legend.position = "right",
    panel.grid.minor = element_blank()
  )


h6_subject_bin_stonehigher <- is_not_consistent_chosegems_part1 %>%
  mutate(
    trial_bin = cut(
      trial_num,
      breaks = c(0, 10, 20, 30, 40, 50),
      labels = c("1-10", "11-20", "21-30", "31-40", "41-50"),
      right = TRUE
    )
  ) %>%
  group_by(subj_id, trial_bin, mentioned_rule) %>%
  filter(stone_higher_than_other) %>%
  summarise(
    prob_stones = mean(chose_onlystones_part2, na.rm = TRUE),
    n_trials = n(),
    .groups = "drop"
  ) %>%
  mutate(
    Group = ifelse(mentioned_rule, "Position mentioned", "Position NOT mentioned"),
    trial_bin = factor(trial_bin, levels = c("1-10","11-20","21-30","31-40","41-50"))
  )

summary_df_stonehigher <- h6_subject_bin_stonehigher %>%
  group_by(trial_bin, Group) %>%
  summarise(
    mean_prob = mean(prob_stones, na.rm = TRUE),
    se = sd(prob_stones, na.rm = TRUE) / sqrt(n()),
    n_subj = n(),
    .groups = "drop"
  )

ggplot(h6_subject_bin_stonehigher, aes(x = trial_bin, y = prob_stones, fill = Group, group = Group)) +
  geom_jitter(aes(color = Group), 
              position = position_jitterdodge(jitter.width = 0.12, dodge.width = 0.75), 
              size = 1.6, alpha = 0.6, show.legend = FALSE) +
  geom_errorbar(data = summary_df_stonehigher,
                aes(x = trial_bin, y = mean_prob, ymin = mean_prob - se, ymax = mean_prob + se, group = Group),
                width = 0.18,
                position = pd,
                color = "black",
                linewidth = 0.8) +
  geom_line(data = summary_df_stonehigher,
            aes(x = trial_bin, y = mean_prob, color = Group, group = Group),
            position = pd,
            linewidth = 0.9,
            alpha = 0.8) +
  geom_point(data = summary_df_stonehigher,
             aes(x = trial_bin, y = mean_prob, group = Group),
             position = pd,
             shape = 23, size = 3, fill = "white", color = "black") +
  geom_hline(yintercept = 0.5, linetype = "dashed", color = "black", linewidth = 0.7) +
  labs(
    title = "Exp. 2: Subject-level probability of choosing stones by trial bin (stone higher)",
    x = "Trial bin",
    y = "Probability of choosing stones (per subject)",
    fill = "Group"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    legend.position = "right",
    panel.grid.minor = element_blank()
  )

library(ggplot2)
library(dplyr)

h6_viz_data <- is_not_consistent_chosegems_part1 %>%

  group_by(subj_id, mentioned_rule) %>%

  summarise(
    n = n(),
    prob_stones = mean(chose_onlystones_part2, na.rm = TRUE),
    se = sqrt(prob_stones * (1 - prob_stones) / n) # Std. Error of a proportion
  ) %>%
  
  mutate(
    ci_lower = prob_stones - 1.96 * se,
    ci_upper = prob_stones + 1.96 * se
  ) %>%
  

  ungroup() %>%
  mutate(
    Group = ifelse(mentioned_rule, "Position mentioned", "Position NOT mentioned")
  )

h6_alltrial <- is_not_consistent_chosegems_part1 %>%
  group_by(subj_id) %>%
  summarise(
    mean_prob_stones = mean(chose_onlystones_part2, na.rm = TRUE),
    mentioned_rule  = unique(mentioned_rule)
  ) 


ggplot(h6_alltrial, aes(x = mentioned_rule, y = mean_prob_stones, fill = mentioned_rule)) +
  geom_boxplot(alpha = 0.5, width = 0.6, outlier.shape = NA, color = "gray40") +

  stat_summary(fun = mean, geom = "point", 
               shape = 23, size = 4, fill = "white", color = "black") +
  
  stat_summary(fun.data = mean_se, geom = "errorbar", 
               width = 0.2, color = "black", linewidth = 0.8) +
  
  geom_jitter(width = 0.1, alpha = 0.5, color = "gray50") +
  
  labs(
    title = "Probability of Choosing Stones by Rule Mention",
    x = "",
    y = "Probability of Choosing Stones",
    fill = "Group"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    legend.position = "none",
    panel.grid.minor = element_blank()
  )

```

## Exploratory analysis 

### estimates of own performance


```{r correlation, include = FALSE, cache=F}
## Correlation between estimate and actual performance 
# part 1

computer_accuracy <- raw_df %>%
  select(subj_id, trial_num, number_correct_part1_computer_slider)

part_1_data <- phase_1_df %>%
  inner_join(computer_accuracy, by=c('subj_id','trial_num'))%>%
  filter(!is_catch_trial) %>%
  mutate(
    rewarded = case_when(
      chose_gems | computer_chose_gems ~ TRUE,
      TRUE ~ FALSE
    )
  )


total_chose_gems_part1 <- part_1_data %>%
  group_by(subj_id) %>%
  filter(!is_catch_trial) %>% 
  summarise(total_gems = sum(chose_gems, na.rm = TRUE)) 

total_rewarded_part1 <- part_1_data %>%
  group_by(subj_id) %>%
  filter(!is_catch_trial) %>% 
  summarise(total_reward = sum(rewarded, na.rm=T))

computer_total_gems <- part_1_data %>%
  group_by(subj_id) %>% 
  filter(!is_catch_trial) %>%
  summarise(total_gems = sum(computer_chose_gems, na.rm=T))

estimate_chose_gems_part1 <- phase_1_df %>% 
  group_by(subj_id) %>%
  summarise(mean_estimate = mean(number_correct_part1, na.rm = TRUE))

estimate_computer_chose_gems <- part_1_data %>%
  group_by(subj_id) %>%
  mutate(number_correct_part1_computer_slider = as.numeric(number_correct_part1_computer_slider)) %>%
  summarise(mean_estimate = mean(number_correct_part1_computer_slider), na.rm=T)

correlation_data_chose_gems <- inner_join(total_chose_gems_part1, 
                               estimate_chose_gems_part1, 
                               by = "subj_id")

correlation_data_rewarded <- inner_join(total_rewarded_part1, estimate_chose_gems_part1, by = 'subj_id')

correlation_data_computer <- inner_join(computer_total_gems, estimate_computer_chose_gems, by='subj_id') 

cor_part1_chose_gems <- cor.test(x = correlation_data_chose_gems$total_gems, 
         y = correlation_data_chose_gems$mean_estimate, 
         method = "pearson")

cor_part1_rewarded  <- cor.test(x = correlation_data_rewarded$total_reward,
                                y = correlation_data_rewarded$mean_estimate,
                                method = "pearson")

cor_part1_computer <- cor.test(x = correlation_data_computer$total_gems,
                               y = correlation_data_computer$mean_estimate,
                               method = "pearson")




# part 2

total_chose_gems_part2 <- phase_2_df %>%
  group_by(subj_id) %>%
  filter(!is_catch_trial) %>% 
  summarise(consistent = sum(is_consistent, na.rm = TRUE)) 

estimate_chose_gems_part2 <- phase_2_df %>% 
  group_by(subj_id) %>%
  summarise(mean_estimate = mean(number_correct_part2, na.rm = TRUE))

correlation_data <- inner_join(total_chose_gems_part2, 
                               estimate_chose_gems_part2, 
                               by = "subj_id")

cor_part2 <- cor.test(x = correlation_data$consistent, 
         y = correlation_data$mean_estimate, 
         method = "pearson")


```


People cannot estimate performance in part 1 (estimating the number of trials in which **they** revealed gems); Pearson correlation `r sprintf("r = %.2f, 95%% CI [%.2f, %.2f]", cor_part1_chose_gems$estimate, cor_part1_chose_gems$conf.int[1], cor_part1_chose_gems$conf.int[2])` p = `r printp(cor_part1_chose_gems$p.value)`. However, people can estimate the number of trials in which they were rewarded; Pearson correlation `r sprintf("r = %.2f, 95%% CI [%.2f, %.2f]", cor_part1_rewarded$estimate, cor_part1_rewarded$conf.int[1], cor_part1_rewarded$conf.int[2])` p `r printp(cor_part1_rewarded$p.value)`. When asked to estimate the performance of the computer, participants fail to do so accurately; Pearson correlation `r sprintf("r = %.2f, 95%% CI [%.2f, %.2f]", cor_part1_computer$estimate, cor_part1_computer$conf.int[1], cor_part1_computer$conf.int[2])` p = `r printp(cor_part1_computer$p.value)`.

In Part 2, they are able to estimate their performance in part 2; Pearson correlation `r sprintf("r = %.2f, 95%% CI [%.2f, %.2f]", cor_part2$estimate,cor_part2$conf.int[1], cor_part2$conf.int[2])` p `r printp(cor_part2$p.value)`. However, participants are unable to estimate the performance of the computer accurately 




```{r estimate,  cache=F, fig.width=8}
library(dplyr)
library(ggplot2)
library(broom)

theme_set(theme_minimal(base_size = 13))

plot_correlation_fixed <- function(df, x_col, y_col, x_lab, y_lab, title) {
  df2 <- df %>%
    select({{x_col}}, {{y_col}}) %>%
    rename(x = 1, y = 2) %>%
    filter(is.finite(x), is.finite(y))
  if (nrow(df2) < 2) stop("Not enough observations to compute correlation for: ", title)
  ct <- cor.test(df2$x, df2$y, method = "pearson")
  r_val <- unname(ct$estimate)
  p_val <- ct$p.value
  n_val <- nrow(df2)

  p <- ggplot(df2, aes(x = x, y = y)) +
    geom_point(alpha = 0.7, size = 2) +
    geom_smooth(method = "lm", se = FALSE, color = "steelblue", linewidth = 0.8) +
    labs(title = title, x = x_lab, y = y_lab) +
    annotate("text", x = Inf, y = Inf,
             label = sprintf("r = %.2f\np = %.3g\nn = %d", r_val, p_val, n_val),
             hjust = 1.05, vjust = 1.1, size = 3.6)
  list(plot = p, cor.test = ct, data = df2)
}

total_chose_gems_part1 <- part_1_data %>%
  filter(!is_catch_trial) %>%
  group_by(subj_id) %>%
  summarise(total_gems = sum(chose_gems, na.rm = TRUE), .groups = "drop")

estimate_chose_gems_part1 <- phase_1_df %>%
  group_by(subj_id) %>%
  summarise(mean_estimate = mean(number_correct_part1, na.rm = TRUE), .groups = "drop")

corr_df_1 <- inner_join(total_chose_gems_part1, estimate_chose_gems_part1, by = "subj_id") %>% drop_na()

res_part1_chose <- plot_correlation_fixed(corr_df_1, total_gems, mean_estimate,
                                    x_lab = "Total times participant chose gems (part1)",
                                    y_lab = "Mean estimate (part1)",
                                    title = "Part1: Participant chosen gems vs estimate")
print(res_part1_chose$plot)
print(res_part1_chose$cor.test)

total_rewarded_part1 <- part_1_data %>%
  filter(!is_catch_trial) %>%
  group_by(subj_id) %>%
  summarise(total_reward = sum(rewarded, na.rm = TRUE), .groups = "drop")

corr_df_2 <- inner_join(total_rewarded_part1, estimate_chose_gems_part1, by = "subj_id") %>% drop_na()

res_part1_reward <- plot_correlation_fixed(corr_df_2, total_reward, mean_estimate,
                                     x_lab = "Total rewarded trials (part1)",
                                     y_lab = "Mean estimate (part1)",
                                     title = "Part1: Rewarded trials vs estimate")
print(res_part1_reward$plot)
print(res_part1_reward$cor.test)

computer_total_gems <- part_1_data %>%
  filter(!is_catch_trial) %>%
  group_by(subj_id) %>%
  summarise(total_gems = sum(computer_chose_gems, na.rm = TRUE), .groups = "drop")

computer_estimate_per_subj <- part_1_data %>%
  group_by(subj_id) %>%
  summarise(mean_computer_estimate = mean(as.numeric(number_correct_part1_computer_slider), na.rm = TRUE), .groups = "drop")

corr_df_3 <- inner_join(computer_total_gems, computer_estimate_per_subj, by = "subj_id") %>% drop_na()

res_part1_computer <- plot_correlation_fixed(corr_df_3, total_gems, mean_computer_estimate,
                                      x_lab = "Computer: total gem reveals (part1)",
                                      y_lab = "Computer mean estimate (slider)",
                                      title = "Part1: Computer gem reveals vs computer estimate")
print(res_part1_computer$plot)
print(res_part1_computer$cor.test)


total_chose_gems_part2 <- phase_2_df %>%
  filter(!is_catch_trial) %>%
  group_by(subj_id) %>%
  summarise(consistent = sum(is_consistent, na.rm = TRUE), .groups = "drop")

estimate_chose_gems_part2 <- phase_2_df %>%
  group_by(subj_id) %>%
  summarise(mean_estimate = mean(number_correct_part2, na.rm = TRUE), .groups = "drop")

corr_df_4 <- inner_join(total_chose_gems_part2, estimate_chose_gems_part2, by = "subj_id") %>% drop_na()

res_part2 <- plot_correlation_fixed(corr_df_4, consistent, mean_estimate,
                             x_lab = "Total consistent choices (part2)",
                             y_lab = "Mean estimate (part2)",
                             title = "Part2: Consistency vs estimate")
print(res_part2$plot)
print(res_part2$cor.test)

```


### Effect of numbers on choice

```{r higherorlower, include = FALSE, cache=F}

higher_df <- phase_2_df %>%
  group_by(subj_id) %>%
  filter(!is_catch_trial, !is_consistent) %>%
  mutate(
    higher = case_when(
      response_position == "middle" & correct_choice_position == "left"  ~ middle_number > right_number,
      response_position == "middle" & correct_choice_position == "right" ~ middle_number > left_number,
      
      response_position == "right" & correct_choice_position == "left"   ~ right_number > middle_number,
      response_position == "right" & correct_choice_position == "middle" ~ right_number > left_number,
      
      response_position == "left" & correct_choice_position == "middle" ~ left_number > right_number,
      response_position == "left" & correct_choice_position == "right"  ~ left_number > middle_number,
      
      TRUE ~ NA
    )
  ) %>%

  filter(
    case_when(
      response_position == "middle" & correct_choice_position == "left"  ~ middle_number != right_number,
      response_position == "middle" & correct_choice_position == "right" ~ middle_number != left_number,
      
      response_position == "right" & correct_choice_position == "left"   ~ right_number != middle_number,
      response_position == "right" & correct_choice_position == "middle" ~ right_number != left_number,
      
      response_position == "left" & correct_choice_position == "middle" ~ left_number != right_number,
      response_position == "left" & correct_choice_position == "right"  ~ left_number != middle_number,

      TRUE ~ FALSE
    )
  )



participant_summary <- higher_df %>%
  group_by(subj_id) %>%
  summarise(
    prop_higher = mean(higher, na.rm = TRUE)
  )


```

When participants were inaccurate in part 2, the mean probability for the participant to choose the option with a greater number was `r higher_df %>% ungroup() %>% summarise(mean=mean(higher)) %>% pull(mean) %>% printnum()`. This was significantly greater than chance (`r apa_print(t.test(participant_summary$prop_higher,mu = 0.5,alternative = "two.sided"))$statistic`). 


```{r age,  cache=F, fig.width=8}
library(dplyr)
library(ggplot2)

# Accuracy
participant_accuracy_1 <- merged_data %>%
  filter(phase == 'part1', !is_catch_trial) %>%
  group_by(`Participant id`, Age) %>%
  summarise(participant_acc = mean(chose_gems), .groups = 'drop')

participant_accuracy_2 <- merged_data %>%
  filter(phase == 'part2', !is_catch_trial) %>%
  group_by(`Participant id`, Age) %>%
  summarise(participant_acc = mean(is_consistent), .groups = 'drop')

age_avoid_1 <- merged_data %>%
  filter(phase == 'part1', !is_catch_trial) %>%
  group_by(`Participant id`, Age) %>%
  summarise(participant_avoid = mean(chose_onlystones), .groups = 'drop')

age_avoid_2 <- merged_data %>%
  filter(phase == 'part2', !is_catch_trial) %>%
  group_by(`Participant id`, Age) %>%
  summarise(participant_avoid = mean(chose_onlystones), .groups = 'drop')

age_breaks <- c(17, 22, 27, 32, 37, 42, 47, 52, 57, 62, 67, 72, 77, Inf)
age_labels <- c("18-22", "23-27", "28-32", "33-37", "38-42", "43-47",
                "48-52", "53-57", "58-62", "63-67", "68-72", "73-77", "78+")

combined_participant_data <- bind_rows(
  participant_accuracy_1 %>% mutate(Condition = "Part 1"),
  participant_accuracy_2 %>% mutate(Condition = "Part 2")
)

combined_age_summary <- combined_participant_data %>%
  mutate(
    Age = as.numeric(as.character(Age)),
    AgeBin = cut(Age, breaks = age_breaks, labels = age_labels, right = TRUE, ordered_result = TRUE)
  ) %>%
  filter(!is.na(AgeBin)) %>%
  group_by(AgeBin, Condition) %>% 
  summarise(
    mean_accuracy = mean(participant_acc),
    se = sd(participant_acc) / sqrt(n()),
    .groups = 'drop'
  )

ggplot(combined_age_summary, aes(x = AgeBin, 
                                 y = mean_accuracy, 
                                 group = Condition, 
                                 color = Condition)) +
  geom_line(linewidth = 1.2, position = position_dodge(width = 0.3)) +
  geom_point(size = 3, position = position_dodge(width = 0.3)) +
  geom_errorbar(
    aes(ymin = mean_accuracy - se, ymax = mean_accuracy + se),
    width = 0.25, linewidth = 0.8, position = position_dodge(width = 0.3)
  ) +
  scale_color_manual(values = c("Part 1" = "blue", "Part 2" = "red")) +
  
  labs(
    title = "Mean Participant Accuracy by Age Bin",
    x = "Age Bin",
    y = "Mean Accuracy",
    color = "Phase" 
  ) +
  coord_cartesian(ylim = c(0, 1)) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

combined_avoid_data <- bind_rows(
  age_avoid_1 %>% mutate(Condition = "Part 1"),
  age_avoid_2 %>% mutate(Condition = "Part 2")
)

combined_age_summary_avoid <- combined_avoid_data %>%
  mutate(
    Age = as.numeric(as.character(Age)),
    AgeBin = cut(Age, breaks = age_breaks, labels = age_labels, right = TRUE, ordered_result = TRUE)
  ) %>%
  filter(!is.na(AgeBin)) %>%
  group_by(AgeBin, Condition) %>%
  summarise(
    mean_avoid = mean(participant_avoid), 
    se = sd(participant_avoid) / sqrt(n()),
    .groups = 'drop'
  )

ggplot(combined_age_summary_avoid, aes(x = AgeBin, 
                                       y = mean_avoid, 
                                       group = Condition, 
                                       color = Condition)) +
  geom_line(linewidth = 1.2, position = position_dodge(width = 0.3)) +
  geom_point(size = 3, position = position_dodge(width = 0.3)) +
  geom_errorbar(
    aes(ymin = mean_avoid - se, ymax = mean_avoid + se), 
    width = 0.25, linewidth = 0.8, position = position_dodge(width = 0.3)
  ) +
  scale_color_manual(values = c("Part 1" = "blue", "Part 2" = "red")) +
  
  labs(
    title = "Mean Participant Avoidance by Age Bin",
    x = "Age Bin",
    y = "Mean Avoidance", 
    color = "Phase"
  ) +
  coord_cartesian(ylim = c(0, 1)) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

participant_age_data <- merged_data %>%
  mutate(Age_numeric = as.numeric(as.character(Age))) %>%
  filter(!is.na(Age_numeric)) %>%
  distinct(`Participant id`, Age_numeric)

conditions <- phase_1_df %>%
  ungroup() %>% 
  distinct(subj_id, stoneposition) %>% 
  count(stoneposition)

print(conditions)
  

ggplot(participant_age_data, aes(x = Age_numeric)) +
  geom_histogram(binwidth = 5, fill = "cornflowerblue", color = "white", alpha = 0.9, boundary = 0) +
  labs(
    title = "Distribution of Participant Age",
    x = "Age",
    y = "Frequency (Count of Participants)" # Updated label
  ) +
  theme_minimal()


```
```{r agebias, cache=F, fig.width=8}
library(dplyr)
library(ggplot2)

correlated_data <- inner_join(
  phase_1_acc,
  phase_2_acc,
  by = "subj_id",
  suffix = c("_p1", "_p2") # <-- THE FIX: 'suffix' (singular)
)

correlation_test <- cor.test(
  ~ acc_p1 + acc_p2, 
  data = correlated_data
)

print(correlation_test)

ggplot(correlated_data, aes(x = acc_p1, y = acc_p2)) +
  geom_point(alpha = 0.6, size = 2) +
  geom_smooth(method = "lm", color = "blue", fill = "blue", alpha = 0.1) +
  labs(
    title = "Correlation between Phase 1 and Phase 2 Accuracy",
    x = "Phase 1 Accuracy",
    y = "Phase 2 Accuracy"
  ) +
  theme_minimal()

```
```{r clean_hover_data, cache=F, fig.width = 8}

## Exclude participants who may have used tablets:

sep <- ";"

mouse_long <- raw_df %>%
  filter(!(subj_id %in% participants_to_exclude), trial_type == "testmouse") %>%
  select(subj_id, trial_num, mouse_x, mouse_y, mouse_t) %>%
  mutate(
    mouse_x = if_else(is.na(mouse_x) | mouse_x == "", NA_character_, as.character(mouse_x)),
    mouse_y = if_else(is.na(mouse_y) | mouse_y == "", NA_character_, as.character(mouse_y)),
    mouse_t = if_else(is.na(mouse_t) | mouse_t == "", NA_character_, as.character(mouse_t))
  ) %>%
  mutate(
    x_tokens = str_split(mouse_x, pattern = fixed(sep)),
    y_tokens = str_split(mouse_y, pattern = fixed(sep)),
    t_tokens = str_split(mouse_t, pattern = fixed(sep))
  ) %>%
  # require at least some x tokens
  filter(!map_lgl(x_tokens, ~ all(is.na(.x)))) %>%
  mutate(
    n_tokens = pmax(map_int(x_tokens, length),
                    map_int(y_tokens, length),
                    map_int(t_tokens, length))
  ) %>%
  rowwise() %>%
  mutate(idx = list(seq_len(n_tokens))) %>%
  ungroup() %>%
  unnest(cols = c(idx)) %>%
  mutate(frame_index = idx) %>%
  mutate(
    x_token = map2_chr(x_tokens, frame_index, ~ if(length(.x) >= .y) .x[[.y]] else NA_character_),
    y_token = map2_chr(y_tokens, frame_index, ~ if(length(.x) >= .y) .x[[.y]] else NA_character_),
    t_token = map2_chr(t_tokens, frame_index, ~ if(length(.x) >= .y) .x[[.y]] else NA_character_)
  ) %>%
  select(subj_id, trial_num, frame_index, x_token, y_token, t_token)

mouse_long <- mouse_long %>%
  mutate(
    mouse_x = as.numeric(str_trim(x_token)),
    mouse_y = as.numeric(str_trim(y_token)),
    mouse_t = as.numeric(str_trim(t_token))
  ) %>%
  select(-x_token, -y_token, -t_token) %>%
  filter(!(is.na(mouse_x) & is.na(mouse_y)))

mouse_movements <- mouse_long %>%
  arrange(subj_id, trial_num, frame_index) %>%
  group_by(subj_id, trial_num) %>%
  mutate(
    x_prev = lag(mouse_x),
    y_prev = lag(mouse_y),
    t_prev = lag(mouse_t),
    dx = as.numeric(mouse_x) - as.numeric(x_prev),
    dy = as.numeric(mouse_y) - as.numeric(y_prev),
    dist = sqrt((dx)^2 + (dy)^2),
    dt = as.numeric(mouse_t) - as.numeric(t_prev),
    speed = ifelse(!is.na(dist) & !is.na(dt) & dt > 0, dist / dt, NA_real_)
  ) %>%
  ungroup()

summary(mouse_movements$dist)
quantile(mouse_movements$dist, probs = c(0, 0.5, 0.9, 0.99, 1), na.rm = TRUE)

mouse_movements %>%
  arrange(desc(dist)) %>%
  slice_head(n = 20) %>%
  select(subj_id, trial_num, frame_index, mouse_x, mouse_y, x_prev, y_prev, dx, dy, dist) 

subject_maxdist <- mouse_movements %>%
  group_by(subj_id) %>%
  summarise(
    n_points   = n(),
    n_moves    = sum(!is.na(dist)),
    max_dist   = max(dist, na.rm = TRUE),
    median_dist= median(dist, na.rm = TRUE),
    p99_dist   = quantile(dist, 0.99, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  arrange(desc(max_dist))


p_hist <- ggplot(subject_maxdist, aes(x = median_dist)) +
  geom_histogram(bins = 60, boundary = 0, closed = "left", fill = "steelblue", color = "white", na.rm = TRUE) +
  labs(title = "Distribution of participants' median frame-to-frame jump",
       x = "median dist (pixels/frame)",
       y = "Number of participants") +
  theme_minimal(base_size = 13)

print(p_hist)

tablet_suspects <- subject_maxdist %>%
  filter(median_dist > 150) %>%
  pull(subj_id)

```

```{r hover_data_analysis, cache=F, fig.width = 8}


hover_part1 <- raw_df %>%
  filter(
    !(subj_id %in% participants_to_exclude),
    trial_type == "testmouse",
    phase == 'part1',
    !(subj_id %in% tablet_suspects)
  ) %>%
  select(
    subj_id, 
    trial_num,
    stoneposition,
    left_number, 
    middle_number, 
    right_number,
    is_catch_trial,
    chose_gems,
    chose_other,
    chose_onlystones,
    frames_hover_left,
    frames_hover_middle,
    frames_hover_right,
    frames_hover_none,
    hover_time_left,
    hover_time_middle,
    hover_time_right,
    hover_time_none, 
    mouse_x,
    mouse_y,
    mouse_t,
    mouse_universal_x,
    mouse_universal_y,
    hover_location,
    total_frames,
    correct_choice_position,
    response_position
  ) %>%
  mutate(

    across(starts_with("frames_hover_"), as.numeric),
    across(starts_with("hover_time_"), as.numeric),

    gem_frame_hover = case_when(
      correct_choice_position == "left"   ~ frames_hover_left,
      correct_choice_position == "middle" ~ frames_hover_middle,
      correct_choice_position == "right"  ~ frames_hover_right,
      TRUE ~ NA_real_
    ),
    gem_time_hover = case_when(
      correct_choice_position == "left"   ~ hover_time_left,
      correct_choice_position == "middle" ~ hover_time_middle,
      correct_choice_position == "right"  ~ hover_time_right,
      TRUE ~ NA_real_
    ),
    stone_frame_hover = case_when(
      stoneposition == "left" ~ frames_hover_left,
      stoneposition == "middle" ~ frames_hover_middle,
      stoneposition == "right" ~ frames_hover_right,
      TRUE ~ NA_real_
    ),
    stone_time_hover = case_when(
      stoneposition == "right" ~ hover_time_right,
      stoneposition == "middle" ~ hover_time_middle,
      stoneposition == "left" ~ hover_time_left,
      TRUE ~ NA_real_
    ),
    other_frame_hover = case_when(
      stoneposition == "left" & correct_choice_position == "middle" ~ frames_hover_right,
      stoneposition == "left" & correct_choice_position == "right" ~ frames_hover_middle,
      stoneposition == "middle" & correct_choice_position == "right" ~ frames_hover_left,
      stoneposition == "middle" & correct_choice_position == "left" ~ frames_hover_right,
      stoneposition == "right" & correct_choice_position == "middle" ~ frames_hover_left,
      stoneposition == "right" & correct_choice_position == "left" ~ frames_hover_middle,
      TRUE ~ NA_real_
    ),
    other_time_hover = case_when(
      stoneposition == "left" & correct_choice_position == "middle" ~ hover_time_right,
      stoneposition == "left" & correct_choice_position == "right" ~ hover_time_middle,
      stoneposition == "middle" & correct_choice_position == "right" ~ hover_time_left,
      stoneposition == "middle" & correct_choice_position == "left" ~ hover_time_right,
      stoneposition == "right" & correct_choice_position == "middle" ~ hover_time_left,
      stoneposition == "right" & correct_choice_position == "left" ~ hover_time_middle,
      TRUE ~ NA_real_
    )
    
  )

hover_part2 <- raw_df %>%
  filter(
    !(subj_id %in% participants_to_exclude),
    trial_type == "testmouse",
    phase == 'part2',
    !(subj_id %in% tablet_suspects)
  ) %>%
  select(
    subj_id, 
    trial_num,
    stoneposition,
    left_number, 
    middle_number, 
    right_number,
    is_catch_trial,
    chose_gems,
    chose_other,
    chose_onlystones,
    frames_hover_left,
    frames_hover_middle,
    frames_hover_right,
    frames_hover_none,
    hover_time_left,
    hover_time_middle,
    hover_time_right,
    hover_time_none, 
    mouse_x,
    mouse_y,
    mouse_t,
    mouse_universal_x,
    mouse_universal_y,
    hover_location,
    total_frames,
    correct_choice_position,
    is_consistent,
    response_position
  ) 


part1_positions <- phase_1_df %>%
  mutate(
    gem_position   = correct_choice_position,
    stone_position = stoneposition,
    other_position = map2_chr(gem_position, stone_position, ~ {
      setdiff(c("left", "middle", "right"), c(.x, .y))[1]
    }),
    computer_response_position = computer_response_position,
    correct_choice_position_part1 = correct_choice_position
  ) %>%
  select(subj_id, trial_num, gem_position, stone_position, other_position,
         computer_response_position, correct_choice_position_part1)

  
hover_part2 <- hover_part2 %>% 
  left_join(part1_positions, by = c('subj_id', 'trial_num'))

hover_part2 <- hover_part2 %>% 
  mutate(

    across(starts_with("frames_hover_"), as.numeric),
    across(starts_with("hover_time_"), as.numeric),

    is_consistent_frame_hover = case_when(
      correct_choice_position == "left"   ~ frames_hover_left,
      correct_choice_position == "middle" ~ frames_hover_middle,
      correct_choice_position == "right"  ~ frames_hover_right,
      TRUE ~ NA_real_
    ),
    is_consistent_time_hover = case_when(
      correct_choice_position == "left"   ~ hover_time_left,
      correct_choice_position == "middle" ~ hover_time_middle,
      correct_choice_position == "right"  ~ hover_time_right,
      TRUE ~ NA_real_
    ),
    stone_frame_hover = case_when(
      stoneposition == "left" ~ frames_hover_left,
      stoneposition == "middle" ~ frames_hover_middle,
      stoneposition == "right" ~ frames_hover_right,
      TRUE ~ NA_real_
    ),
    stone_time_hover = case_when(
      stoneposition == "right" ~ hover_time_right,
      stoneposition == "middle" ~ hover_time_middle,
      stoneposition == "left" ~ hover_time_left,
      TRUE ~ NA_real_
    ),
    gem_frame_hover = case_when(
      correct_choice_position_part1 == 'left' ~ frames_hover_left,
      correct_choice_position_part1 == 'middle' ~ frames_hover_middle,
      correct_choice_position_part1 == 'right' ~ frames_hover_right
    ),
    gem_time_hover = case_when(
      correct_choice_position_part1 == 'left' ~ hover_time_left,
      correct_choice_position_part1 == "middle" ~ hover_time_middle,
      correct_choice_position_part1 == 'right' ~ hover_time_right
    ),
    other_frame_hover = case_when(
      stoneposition == "left" & correct_choice_position_part1 == "middle" ~ frames_hover_right,
      stoneposition == "left" & correct_choice_position_part1 == "right" ~ frames_hover_middle,
      stoneposition == "middle" & correct_choice_position_part1 == "right" ~ frames_hover_left,
      stoneposition == "middle" & correct_choice_position_part1 == "left" ~ frames_hover_right,
      stoneposition == "right" & correct_choice_position_part1 == "middle" ~ frames_hover_left,
      stoneposition == "right" & correct_choice_position_part1 == "left" ~ frames_hover_middle,
      TRUE ~ NA_real_
    ),
    other_time_hover = case_when(
      stoneposition == "left" & correct_choice_position_part1 == "middle" ~ hover_time_right,
      stoneposition == "left" & correct_choice_position_part1== "right" ~ hover_time_middle,
      stoneposition == "middle" & correct_choice_position_part1 == "right" ~ hover_time_left,
      stoneposition == "middle" & correct_choice_position_part1== "left" ~ hover_time_right,
      stoneposition == "right" & correct_choice_position_part1== "middle" ~ hover_time_left,
      stoneposition == "right" & correct_choice_position_part1== "left" ~ hover_time_middle,
      TRUE ~ NA_real_
    ),
    
    
  )


hover_part1_joined <- hover_part1 %>%
  left_join(part1_positions, by = c("subj_id", "trial_num"))

hover1_plot_df <- hover_part1_joined %>%
  mutate(
    across(starts_with("hover_time_"), as.numeric),

    consistent_slot = case_when(
      chose_gems   ~ "gem",
      chose_onlystones ~ "stone",
      chose_other ~ "other",
      TRUE                                ~ NA_character_
    )
  ) %>%
  pivot_longer(
    cols = c(gem_time_hover, stone_time_hover, other_time_hover),
    names_to = "hovered_item",
    values_to = "hover_time"
  ) %>%
  mutate(
    hovered_item = recode(hovered_item,
                          gem_time_hover   = "gem",
                          stone_time_hover = "stone",
                          other_time_hover = "other"),
    consistent_slot = factor(consistent_slot,
                             levels = c("gem","stone","other"),
                             labels = c("chose gems", "chose stones", "chose other")),
    hovered_item = factor(hovered_item, levels = c("gem","stone","other"))
  ) %>%
  filter(!is.na(consistent_slot), !is.na(hover_time))


summary_hover1 <- hover1_plot_df %>%
  group_by(consistent_slot, hovered_item) %>%
  summarise(
    mean_hover = mean(hover_time, na.rm = TRUE),
    sd_hover   = sd(hover_time, na.rm = TRUE),
    n          = sum(!is.na(hover_time)),
    se_hover   = sd_hover / sqrt(n),
    .groups = "drop"
  ) %>%
  mutate(se_hover = ifelse(n <= 1, NA_real_, se_hover))

pd <- position_dodge(width = 0.7)

ggplot(summary_hover1, aes(x = consistent_slot, y = mean_hover, fill = hovered_item, group = hovered_item)) +
  geom_col(position = pd, width = 0.6, alpha = 0.95) +
  geom_errorbar(aes(ymin = mean_hover - 1.96 * se_hover, ymax = mean_hover + 1.96 * se_hover),
                position = pd, width = 0.2, color = "black") +
  labs(
    x = "", y = "Mean hover time (ms)", fill = "Hovered item",
    title = "Hover time in part 1 by chosen position and box type"
  ) +
  theme_minimal(base_size = 13) +
  theme(legend.position = "top")

hover_plot_df <- hover_part2 %>%
  mutate(
    consistent_slot = case_when(
      response_position == gem_position   ~ "consistent_is_gem",
      response_position == stone_position ~ "consistent_is_stone",
      response_position == other_position ~ "consistent_is_other",
      TRUE                                ~ NA_character_
    )
  ) %>%
  pivot_longer(
    cols = c(gem_time_hover, stone_time_hover, other_time_hover),
    names_to = "hovered_item",
    values_to = "hover_time"
  ) %>%
  mutate(
    hovered_item = recode(hovered_item,
                          gem_time_hover   = "gem",
                          stone_time_hover = "stone",
                          other_time_hover = "other"),
    consistent_slot = factor(consistent_slot,
                             levels = c("consistent_is_gem","consistent_is_stone","consistent_is_other"),
                             labels = c("is_consistent = gem", "is_consistent = stone", "is_consistent = other")),
    hovered_item = factor(hovered_item, levels = c("gem","stone","other"))
  ) %>%
  filter(!is.na(consistent_slot), !is.na(hover_time))


summary_hover <- hover_plot_df %>%
  group_by(consistent_slot, hovered_item) %>%
  summarise(
    mean_hover = mean(hover_time, na.rm = TRUE),
    sd_hover   = sd(hover_time, na.rm = TRUE),
    n         = sum(!is.na(hover_time)),
    se_hover  = sd_hover / sqrt(n),
    .groups = "drop"
  )

pd <- position_dodge(width = 0.7)

ggplot(summary_hover, aes(x = consistent_slot, y = mean_hover, fill = hovered_item, group = hovered_item)) +
  geom_col(position = pd, width = 0.6, alpha = 0.9) +
  geom_errorbar(aes(ymin = mean_hover - 1.96 * se_hover, ymax = mean_hover + 1.96 * se_hover),
                position = pd, width = 0.2, color = "black") +
  labs(
    x = "", y = "Mean hover time (ms)", fill = "Hovered box type",
    title = "Hover time in part 2 by correct position and box type"
  ) +
  theme_minimal(base_size = 13) +
  theme(legend.position = "top")




```

```{r accuracy_computer_neither, cache = F}



```


# Results

# Discussion


\newpage

# References

::: {#refs custom-style="Bibliography"}
:::
